### 2018-11-26 tspl/instructions.tex
added instructions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Instructions for TSPL Exam
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{article}
\usepackage{a4,amssymb}
% \setlength{\oddsidemargin}{-1.5cm}
% \addtolength{\textwidth}{2cm}
% \addtolength{\textheight}{3cm}

\begin{document}
\pagestyle{empty}
\setcounter{page}{1}

\begin{center}
\large Types and Semantics for Programming Languages
\end{center}

\section*{Instructions}

\begin{enumerate}

\item  The exam lasts two hours.

\item Place your student identity card face-up on the desk in
  front of you.  The invigilator may come to check your identity, and
  in this case you may be asked to allow the invigilator to briefly
  use your computer.  The exam time has been calculated to allow time
  for such interruptions.

\item You may log into your computer as soon as you are ready to do so.

\item To download the exam paper, open a terminal window and 
type the following:
\begin{center}
   \texttt{getpapers}
\end{center}
This will create a subdirectory \texttt{tspl-pe} in your home directory,
containing the following files.
\begin{center}
\begin{tabular}{ll}
\texttt{tspl-pe/papers/exam.pdf}			& the exam \\
\texttt{tspl-pe/papers/instructions.pdf}		& these instructions \\
\texttt{tspl-pe/templates/Exam.lagda}			& exam template to edit \\
\texttt{tspl-pe/original\_templates/Exam.lagda}		& backup of exam template
\end{tabular}
\end{center}
The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates} are
read only, but you may read and write \texttt{tspl-pe/templates}.

\item The textbook \emph{Programming Language Foundations in Agda} is available
by pointing your browser at:
\begin{center}
\texttt{file:///group/examreadonly/full/plfa/}
\end{center}
\item The Agda standard library is available by pointing your browser at:
\begin{center}
\texttt{file:///group/examreadonly/full/agda-stdlib/}
\end{center}
\item The Agda user manual is available by pointing your browser at:
\begin{center}
\texttt{file:///group/examreadonly/full/agda-docs/}
\end{center}
{\it (Note that internet access has been disabled.)}

\begin{center}
  \textbf{Do nothing further until the start of the exam is announced!}
\end{center}

\hfill \textit{Please Turn Over}
\newpage

\item When the start of the exam is announced, open the exam paper
\begin{center}
  \texttt{tspl-pe/papers/exam.pdf}
\end{center}
with the standard PDF viewer.

\item Change to the template directory
\begin{center}
  \texttt{cd tspl-pe/templates}
\end{center}
and edit the file
\begin{center}
  \texttt{Exam.lagda}
\end{center}
to include your answers, using Emacs or anything else suitable.
You are recommended to save your work on a regular basis.

\item Before submitting, make sure your file is processed by Agda
  correctly. Code which prevents the file from compiling should be
  made into comments. If you fail to solve part of a problem, you
  may get more credit if you indicate clearly which part you have
  not solved.

\item \emph{Please ensure before submission that the file}
  \texttt{Exam.lagda} \emph{contains your solutions to the exam.}  Submit
  your file using the command:
  \begin{center}
  \texttt{examsubmit Exam.lagda}
  \end{center}
  If you get an error, please check carefully that your file is called
  \texttt{Exam.lagda} and that you are in the same directory as this
  file. If you continue to have problems, please contact one of the
  invigilators.

Repeated submit commands are allowed, and will overwrite previous
submissions.  The last file submitted will be the one marked.

\item When the invigilators announce the end of the exam, you must
  submit and log out immediately.

\end{enumerate}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-12-02 _site/tspl/instructions.tex
Deploy plfa/plfa.github.io to github.com/plfa/plfa.github.io.git:master

--- 2018-11-26 tspl/instructions.tex
@@
 
-\item  The exam lasts two hours.
+\item
+  The exam lasts two hours.
 
-\item Place your student identity card face-up on the desk in
-  front of you.  The invigilator may come to check your identity, and
-  in this case you may be asked to allow the invigilator to briefly
-  use your computer.  The exam time has been calculated to allow time
-  for such interruptions.
+\item
+  Place your student identity card face-up on the desk in front of you.  The
+  invigilator may come to check your identity, and in this case you may be asked
+  to allow the invigilator to briefly use your computer.  The exam time has been
+  calculated to allow time for such interruptions.
 
-\item You may log into your computer as soon as you are ready to do so.
+\item
+  You may log into your computer as soon as you are ready to do so.
 
-\item To download the exam paper, open a terminal window and 
-type the following:
+\item
+  To download the exam paper, open a terminal window and type:
   \begin{center}
@@
   \end{center}
-This will create a subdirectory \texttt{tspl-pe} in your home directory,
+  This will create a subdirectory \texttt{tspl-mpe} in your home directory,
   containing the following files.
@@
     \begin{tabular}{ll}
-\texttt{tspl-pe/papers/exam.pdf}			& the exam \\
-\texttt{tspl-pe/papers/instructions.pdf}		& these instructions \\
-\texttt{tspl-pe/templates/Exam.lagda}			& exam template to edit \\
-\texttt{tspl-pe/original\_templates/Exam.lagda}		& backup of exam template
+      \texttt{tspl-mpe/papers/exam.pdf}                & the exam \\
+      \texttt{tspl-mpe/papers/instructions.pdf}        & these instructions \\
+      \texttt{tspl-mpe/templates/Exam.lagda}           & exam template to edit \\
+      \texttt{tspl-mpe/original\_templates/Exam.lagda} & backup of exam template
     \end{tabular}
   \end{center}
-The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates} are
-read only, but you may read and write \texttt{tspl-pe/templates}.
+  The directories \texttt{tspl-mpe/papers}, \texttt{tspl-mpe/original\_templates}
+  are read only, but you may read and write \texttt{tspl-mpe/templates}.
 
-\item The textbook \emph{Programming Language Foundations in Agda} is available
-by pointing your browser at:
+\item
+  To setup the Agda environment, open a terminal window and type:
+  and type:
   \begin{center}
-\texttt{file:///group/examreadonly/full/plfa/}
+    \texttt{tspl-setup}
   \end{center}
-\item The Agda standard library is available by pointing your browser at:
-\begin{center}
-\texttt{file:///group/examreadonly/full/agda-stdlib/}
-\end{center}
-\item The Agda user manual is available by pointing your browser at:
-\begin{center}
-\texttt{file:///group/examreadonly/full/agda-docs/}
-\end{center}
+  This will also open a browser, with three tabs which contain the textbook
+  \emph{Programming Language Foundations in Agda}, the documentation for the
+  Agda standard library, and the documentation for Agda are available.
+
 {\it (Note that internet access has been disabled.)}
@@
 \begin{center}
-  \texttt{tspl-pe/papers/exam.pdf}
+  \texttt{tspl-mpe/papers/exam.pdf}
 \end{center}
@@
 \begin{center}
-  \texttt{cd tspl-pe/templates}
+  \texttt{cd tspl-mpe/templates}
 \end{center}

### 2018-12-13 tspl/instructions.tex
updated instructions

@@
     \begin{tabular}{ll}
-\texttt{tspl-pe/papers/exam.pdf}			& the exam \\
-\texttt{tspl-pe/papers/instructions.pdf}		& these instructions \\
-\texttt{tspl-pe/templates/Exam.lagda}			& exam template to edit \\
-\texttt{tspl-pe/original\_templates/Exam.lagda}		& backup of exam template
+      \texttt{tspl-mpe/papers/exam.pdf}                & the exam \\
+      \texttt{tspl-mpe/papers/instructions.pdf}        & these instructions \\
+      \texttt{tspl-mpe/templates/Exam.lagda}           & exam template to edit \\
+      \texttt{tspl-mpe/original\_templates/Exam.lagda} & backup of exam template
     \end{tabular}
   \end{center}
-The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates} are
-read only, but you may read and write \texttt{tspl-pe/templates}.
+  The directories \texttt{tspl-mpe/papers}, \texttt{tspl-mpe/original\_templates}
+  are read only, but you may read and write \texttt{tspl-mpe/templates}.
 
-\item The textbook \emph{Programming Language Foundations in Agda} is available
-by pointing your browser at:
+\item
+  To setup the Agda environment, open a second terminal window and type:
+  and type:
   \begin{center}
-\texttt{file:///group/examreadonly/full/plfa/}
+    \texttt{tspl-setup}
   \end{center}
-\item The Agda standard library is available by pointing your browser at:
-\begin{center}
-\texttt{file:///group/examreadonly/full/agda-stdlib/}
-\end{center}
-\item The Agda user manual is available by pointing your browser at:
-\begin{center}
-\texttt{file:///group/examreadonly/full/agda-docs/}
-\end{center}
+  This will open a browser, with three tabs which contain the textbook
+  \emph{Programming Language Foundations in Agda}, the documentation for the
+  Agda standard library, and the documentation for Agda.  (The browser may
+  also attempt to link to the internet, which brings up a tab labeled
+  ``Problem loading page''; this is expected and not a problem.)
+
 {\it (Note that internet access has been disabled.)}

### 2018-12-13 tspl/instructions.tex
updated instructions

--- 2018-11-26 tspl/instructions.tex
@@
   \end{center}
-The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates} are
-read only, but you may read and write \texttt{tspl-pe/templates}.
+  The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates}
+  are read only, but you may read and write \texttt{tspl-pe/templates}.
 
-\item The textbook \emph{Programming Language Foundations in Agda} is available
-by pointing your browser at:
+\item
+  To setup the Agda environment, open a second terminal window and type:
+  and type:
   \begin{center}
-\texttt{file:///group/examreadonly/full/plfa/}
+    \texttt{tspl-setup}
   \end{center}
-\item The Agda standard library is available by pointing your browser at:
-\begin{center}
-\texttt{file:///group/examreadonly/full/agda-stdlib/}
-\end{center}
-\item The Agda user manual is available by pointing your browser at:
-\begin{center}
-\texttt{file:///group/examreadonly/full/agda-docs/}
-\end{center}
+  This will open a browser, with three tabs which contain the textbook
+  \emph{Programming Language Foundations in Agda}, the documentation for the
+  Agda standard library, and the documentation for Agda.  (The browser may
+  also attempt to link to the internet, which brings up a tab labeled
+  ``Problem loading page''; this is expected and not a problem.)
+
 {\it (Note that internet access has been disabled.)}

### 2019-05-29 courses/tspl/2018/Instructions.tex
Setup courses/ folder.

--- 2018-12-13 tspl/instructions.tex
@@
   \end{center}
-  The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates}
+  The directories \texttt{tspl-pe/papers} and
+  \texttt{tspl-pe/original\_templates}
   are read only, but you may read and write \texttt{tspl-pe/templates}.
@@
   To setup the Agda environment, open a second terminal window and type:
-  and type:
   \begin{center}

### 2019-07-01 sbmf/sbmf/PLFA.tex
Making a subdirectory of sbmf

%\documentclass[runningheads,oribibl]{llncs}
\documentclass[runningheads]{llncs}
\usepackage{natbib}
% \usepackage{stmaryrd}
% \usepackage{proof}
% \usepackage{amsmath}
% \usepackage{amssymb}
% \usepackage{color}
% \usepackage{bbm}
% \usepackage[greek,english]{babel}
% \usepackage{ucs}
% \usepackage[utf8x]{inputenc}
% \usepackage{autofe}
\usepackage{agda}
\usepackage{revsymb}
\usepackage{semantic}
\usepackage{graphicx}
\usepackage{url}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{stix}
\setcitestyle{round,aysep={}}

\begin{document}
\title{Programming Language Foundations in Agda}
\author{Philip Wadler}
\authorrunning{P. Wadler}
\institute{University of Edinburgh\\
  \email{wadler@inf.ed.ac.uk}}
\maketitle
%
\begin{abstract}
One of the leading textbooks for formal methods is
\emph{Software Foundations} (SF), written by Benjamin Pierce in
collaboration with others, and based on Coq. After five years using SF
in the classroom, I have come to the conclusion that Coq is not the
best vehicle for this purpose, as too much of the course needs to
focus on learning tactics for proof derivation, to the cost of
learning programming language theory.  Accordingly, I have written a
new textbook, \emph{Programming Language Foundations in Agda} (PLFA).
PLFA covers much of the same ground as SF, although it is not a
slavish imitation.

What did I learn from writing PLFA? First, that it is possible. One
might expect that without proof tactics that the proofs become too
long, but in fact proofs in PLFA are about the same length as those in
SF. Proofs in Coq require an interactive environment to be understood,
while proofs in Agda can be read on the page.  Second, that
constructive proofs of preservation and progress give immediate rise
to a prototype evaluator. This fact is obvious in retrospect but it is
not exploited in SF (which instead provides a separate normalise
tactic) nor can I find it in the literature.  Third, that using raw
terms with a separate typing relation is far less perspicuous than
using inherently-typed terms. SF uses the former presentation, while
PLFA presents both; the former uses about 1.6 as many lines of Agda
code as the latter, roughly the golden ratio.

The textbook is written as a literate Agda script, and can be found here:
\begin{center}
\url{http://plfa.inf.ed.ac.uk}
\end{center}
\keywords{Agda \and Coq \and lambda calculus \and dependent types.}
\end{abstract}

\section{Introduction}

The most profound connection between logic and computation is a pun.
The doctrine of Propositions as Types asserts that a certain kind of formal
structure may be read in two ways: either as a proposition in logic or
as a type in computing.  Further, a related structure may be read as
either the proof of the proposition or as a programme of the
corresponding type.  Further still, simplification of proofs
corresponds to evaluation of programs.

Accordingly, the title of this paper, and the corresponding textbook,
\emph{Programming Language Foundations in Agda} (hence, PLFA)
also has two readings.  It may be parsed as ``(Programming Language)
Foundations in Agda'' or ``Programming (Language Foundations) in
Agda''---specifications in the proof assistant Agda both describe
programming languages and are themselves programmes.

Since 2013, I have taught a course on Types and Semantics for
Programming Languages to fourth-year undergraduates and masters
students at the University of Edinburgh.  An earlier version of that
course was based on \emph{Types and Programming Languages} by
\citet{Pierce-2002}, but my version was taught from its successor,
\emph{Software Foundations} (hence, SF) by \citet{Pierce-et-al-2010},
which is based on the proof assistance Coq \citep{Huet-et-al-1997}.
I am convinced by the claim of \citet{Pierce-2009}, made in his ICFP
Keynote \emph{Lambda, The Ultimate TA}, that basing a course around a
proof assistant aids learning.

However, after five years of experience, I have come to the conclusion
that Coq is not the best vehicle.  Too much of the course needs to
focus on learning tactics for proof derivation, to the cost of
learning the fundamentals of programming language theory.  Every
concept has to be learned twice: e.g., both the product data type, and
the corresponding tactics for introduction and elimination of
conjunctions.  The rules Coq applies to generate induction hypotheses
can sometimes seem mysterious.  While the \texttt{notation} construct
permits pleasingly flexible syntax, it can be confusing that the same
concept must always be given two names, e.g., both
\texttt{subst~N~x~M} and \texttt{N~[x~:=~M]}.  Names of tactics are
sometimes short and sometimes long; naming conventions in the standard
library can be wildly inconsistent.  \emph{Propositions as types} as a
foundation of proof is present but hidden.

I found myself keen to recast the course in Agda \citep{Bove-et-al-2009}.
In Agda, there is
no longer any need to learn about tactics: there is just
dependently-typed programming, plain and simple. Introduction is
always by a constructor, elimination is always by pattern
matching. Induction is no longer a mysterious separate concept, but
corresponds to the familiar notion of recursion. Mixfix syntax is
flexible while using just one name for each concept, e.g.,
substitution is \texttt{\_[\_:=\_]}. The standard library is not perfect, but
there is a fair attempt at consistency. \emph{Propositions as types} as a
foundation of proof is on proud display.

Alas, there is no textbook for programming language theory in
Agda.  \emph{Verified Functional Programming in Agda} by \citep{Stump-2016}
covers related ground, but focusses more on programming with dependent
types than on the theory of programming languages.

The original goal was to simply adapt \emph{Software Foundations},
maintaining the same text but transposing the code from Coq to Agda.
But it quickly became clear to me that after five years in the
classroom I had my own ideas about how to present the material.  They
say you should never write a book unless you cannot \emph{not} write the
book, and I soon found that this was a book I could not not write.

I am fortunate that my student, Wen Kokke, was keen to help.  She
guided me as a newbie to Agda and provided an infrastructure that is
easy to use and produces pages that are a pleasure to view.
The bulk of the book was written January--June 2018, while on
sabbatical in Rio de Janeiro.

This paper is a personal reflection, summarising what I learned in the
course of writing the textbook. Some of it reiterates advice that is
well-known to some members of the dependently-typed programming
community, but which deserves to be better known.  The paper is
organised as follows.

Section~2 outlines the
topics covered in PLFA, and notes what is omitted.

Section~3 compares Agda and Coq as vehicles for pedagogy.  Before
writing the book, it was not obvious that it was even possible;
conceivably, without tactics some of the proofs might
balloon in size. In fact, it turns out that for the results in PLFA
and SF, the proofs are of roughly comparable size, and (in my opinion)
the proofs in PLFA are more readable and have a pleasing visual
structure.

Section~4 observes that constructive proofs of progress and
preservation combine trivially to produce a constructive evaluator for
terms.  This idea is obvious once you have seen it, yet I cannot
find it described in the literature.  For instance, SF separately
implements a \texttt{normalise} tactic that has nothing to do with
progress and preservation.

Section~5 claims that raw terms should be avoided in favour of
inherently-typed terms.  PLFA develops lambda calculus with both raw
and inherently-typed terms, permitting a comparison.  It turns out the
former is less powerful---it supports substitution only for closed
terms---but significantly longer---about 1.6 times as many lines of code,
roughly the golden ratio.

I will argue that Agda has advantages over Coq for
pedagogic purposes.  My focus is purely on the case of a proof assistant
as an aid to \emph{learning} formal semantics using examples of
\emph{modest} size.  I admit up front that
there are many tasks for which Coq is better suited than Agda.
A proof assistant that supports tactics, such as Coq or Isabelle,
is essential for formalising serious mathematics,
such as the Four-Colour Theorem \citep{Gonthier-2008},
the Odd-Order Theorem \citep{Gonthier-et-al-2013},
or Kepler's Conjecture \citep{Hales-et-al-2017},
or for establishing correctness of software at scale,
as with the CompCert compiler \citep{Leroy-2009,Kastner-et-al-2017}
or the SEL4 operating system \citep{Klein-2009,O'Connor-2016}.

\section{Scope}

PLFA is aimed at students in the last year of an undergraduate
honours programme or the first year of a master or doctorate degree.
It aims to teach the fundamentals of operational semantics of
programming languages, with simply-typed lambda calculus as the
central example.  The textbook is written as a literate script in Agda.
As with SF, the hope is that using
a proof assistant will make the development more concrete
and accessible to students, and give them rapid feedback to find
and correct misaprehensions.

The book is broken into two parts. The first part, Logical Foundations,
develops the needed formalisms.  The second part, Programming Language
Foundations, introduces basic methods of operational semantics.
(SF is divided into books, the first two of which have the same names
as the two parts of PLFA, and cover similar material.)

Each chapter has both a one-word name and a title, the one-word name
being both its module name and its file name.

\subsection*{Part I, Logical Foundations}

\paragraph{Naturals: Natural numbers}
Introduces the inductive definition of natural numbers in terms of
zero and successor, and recursive definitions of addition,
multiplication, and monus. Emphasis is put on how a tiny description
can specify an infinite domain.

\paragraph{Induction: Proof by induction}
Introduces induction to prove properties
such as associativity and commutativity of addition.
Also introduces dependent functions to express universal quantification.
Emphasis is put on the correspondence between induction and recursion.

\paragraph{Relations: Inductive definitions of relations}
Introduces inductive definitions of less than or equal on natural numbers,
and odd and even natural numbers.
Proves properties such as reflexivity, transitivity, and
anti-symmetry, and that the sum of two odd numbers is even.
Emphasis is put on proof by induction over evidence that a relation holds.

\paragraph{Equality: Equality and equational reasoning}
Gives Martin L\"of's and Leibniz's definitions of equality, and proves
them equivalent, and defines the notation for equational reasoning used
throughout the book.

\paragraph{Isomorphism: Isomorphism and embedding}
Introduces isomorphism, which plays an important role in the
subsequent development.  Also introduces dependent records, lambda
terms, and extensionality.

\paragraph{Connectives: Conjunction, disjunction, and implication}
Introduces product, sum, unit, empty, and function types, and their
interpretations as connectives of logic under Propositions as Types.
Emphasis is put on the analogy between these types and product, sum,
unit, zero, and exponential on naturals; e.g., product of numbers is
commutative and product of types is commutative up to isomorphism.

\paragraph{Negation: Negation, with intuitionistic and classical logic}
Introduces logical negation as a function into the empty
type, and explains the difference between classical and intuitionistic
logic.

\paragraph{Quantifiers: Universals and existentials}
Recaps universal quantifiers and their correspondence to dependent
functions, and introduces existential quantifiers and their
correspondence to dependent products.

\paragraph{Decidable: Booleans and decision procedures}
Introduces booleans and decidable types, and why the latter is to be
preferred to the former.

\paragraph{Lists: Lists and higher-order functions}
Gives two different definitions of reverse and proves them equivalent.
Introduces map and fold and their properties, including that fold left
and right are equivalent in a monoid.  Introduces predicates that hold
for all or any member of a list, with membership as a specialisation
of the latter.

\subsection*{Part II, Programming Language Foundations}

\paragraph{Lambda: Introduction to lambda calculus}
Introduces lambda calculus, using a representation with named
variables and a separate typing relation. The language used is PCF
\citep{Plotkin-1977}, with variables, lambda abstraction, application,
zero, successor, case over naturals, and fixpoint. Reduction is
call-by-value and restricted to closed terms.

\paragraph{Properties: Progress and preservation}
Proves key properties of simply-typed
lambda calculus, including progress and preservation.  Progress and
preservation are combined to yield an evaluator.

\paragraph{DeBruijn: Inherently typed de Bruijn representation}
Introduces de Bruijn indices and the inherently-typed representation.
Emphasis is put on the structural similarity between a term and its
corresponding type derivation; in particular, de Bruijn indices
correspond to the judgment that a variable is well-typed under a given
environment.

\paragraph{More: More constructs of simply-typed lambda calculus}
Introduces product, sum, unit, and empty types as well as lists and let bindings
are explained.  Typing and reduction rules are given informally; a few
are then give formally, and the rest are left as exercises for the reader.
The inherently typed representation is used.

\paragraph{Bisimulation: Relating reduction systems}
Shows how to translate the language with ``let'' terms
to the language without, representing a let as an application of an abstraction,
and shows how to relate the source and target languages with a bisimulation.

\paragraph{Inference: Bidirectional type inference}
Introduces bidirectional type inference, and applies it to convert from
a representation with named variables and a separate typing relation
to a representation de Bruijn indices with inherent types. Bidirectional
type inference is shown to be both sound and complete.

\paragraph{Untyped: Untyped calculus with full normalisation}
As a variation on earlier themes, discusses an untyped (but inherently
scoped) lambda calculus.  Reduction is call-by-name over open terms,
with full normalisation (including reduction under lambda terms).  Emphasis
is put on the correspondence between the structure of a term and
evidence that it is in normal form.

\subsection*{Discussion}

PLFA and SF differ in several particulars.  PLFA begins with a computationally
complete language, PCF, while SF begins with a minimal language, simply-typed
lambda calculus with booleans.  PLFA does not include type annotations in terms,
and uses bidirectional type inference, while SF has terms with unique types and
uses type checking.  SF also covers a simple imperative language with Hoare logic,
and for lambda calculus covers subtyping, record types, mutable references, and normalisation
---none of which are treated by PLFA.  PLFA covers an inherently-typed de Bruijn
representation, bidirectional type inference, bisimulation, and an
untyped call-by-name language with full normalisation---none of which
are treated by SF.

SF has a third volume, written by Andrew Appel, on Verified Functional
Algorithms. I'm not sufficiently familiar with that volume to have a view
on whether it would be easy or hard to cover that material in Agda.
And SF recently added a fourth volume on random testing of Coq specifications
using QuickChick.  There is currently no tool equivalent to QuickChick
available for Agda.

There is more material that would be desirable to include in PLFA
which was not due to limits of time.  In future years, PLFA may be
extended to cover additional material, including mutable references,
normalisation, System F, pure type systems, and denotational
semantics.  I'd especially like to include pure type systems as they
provide the readers with a formal model close to the dependent types
used in the book.  My attempts so far to formalise pure type systems
have proved challenging, to say the least.

\section{Proofs in Agda and Coq}

\begin{figure}[t]
  \includegraphics[width=\textwidth]{figures/plfa-progress-1.png}
  \caption{PLFA, Progress (1/2)}
  \label{fig:plfa-progress-1}
\end{figure}

\begin{figure}[p]
  \includegraphics[width=\textwidth]{figures/plfa-progress-2.png}
  \includegraphics[width=\textwidth]{figures/plfa-progress-3a.png}
  \caption{PLFA, Progress (2/2)}
  \label{fig:plfa-progress-2}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=\textwidth]{figures/sf-progress-1.png}
  \caption{SF, Progress (1/2)}
  \label{fig:sf-progress-1}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=\textwidth]{figures/sf-progress-2.png}
  \caption{SF, Progress (2/2)}
  \label{fig:sf-progress-2}
\end{figure}

% \begin{figure}[t]
%   \includegraphics[width=\textwidth]{figures/sf-progress-3.png}
%   \caption{SF, Progress (3/3)}
%   \label{fig:sf-progress-3}
% \end{figure}

The introduction listed several reasons for preferring Agda over Coq.
But Coq tactics enable more compact proofs. Would it be possible for
PLFA to cover the same material as SF, or would the proofs
balloon to unmanageable size?

As an experiment, I first rewrote SF's development of simply-typed
lambda calculus (SF, Chapters Stlc and StlcProp) in Agda.  I was a
newbie to Agda, and translating the entire development, sticking as
closely as possible to the development in SF, took me about two days.
I was pleased to discover that the proofs remained about the same
size.

There was also a pleasing surprise regarding the structure of the
proofs. While most proofs in both SF and PLFA are carried out by
induction over the evidence that a term is well typed, in SF the
central proof, that substitution preserves types, is carried out by
induction on terms for a technical reason
(the context is extended by a variable binding, and hence not
sufficiently ``generic'' to work well with Coq's induction tactic). In
Agda, I had no trouble formulating the same proof over evidence that
the term is well typed, and didn't even notice SF's description of the
issue until I was done.

The rest of the book was relatively easy to complete.  The closest to
an issue with proof size arose when proving that reduction is
deterministic.  There are 18 cases, one case per line.  Ten of the
cases deal with the situation where there are potentially two
different reductions; each case is trivially shown to be
impossible.  Five of the ten cases are redundant, as they just involve
switching the order of the arguments.  I had to copy the cases
suitably permuted. It would be preferable to reinvoke the proof on
switched arguments, but this would not pass Agda's termination checker
since swapping the arguments doesn't yield a recursive call on
structurally smaller arguments.  I suspect tactics could cut down the
proof significantly. I tried to compare with SF's proof that reduction
is deterministic, but failed to find that proof.

SF covers an imperative language with Hoare logic, culminating in
code that takes an imperative programme suitably decorated with preconditions
and postconditions and generates the necessary
verification conditions. The conditions are then verified by a custom
tactic, where any questions of arithmetic are resolved by the
``omega'' tactic invoking a decision procedure.  The entire
exercise would be easy to repeat in Agda, save for the last step: I
suspect Agda's automation would not be up to verifying the generated
conditions, requiring tedious proofs by hand.  However, I had already
decided to omit Hoare logic in order to focus on lambda calculus.

To give a flavour of how the texts compare, I show the
proof of progress for simply-typed lambda calculus from both texts.
Figures~\ref{fig:plfa-progress-1} and \ref{fig:plfa-progress-2}
are taken from PLFA, Chapter Properties,
while Figures~\ref{fig:sf-progress-1} and \ref{fig:sf-progress-2}
are taken from SF, Chapter StlcProp.
Both texts are intended to be read online,
and the figures were taken by grabbing bitmaps of the text as
displayed in a browser.

PLFA puts the formal statements first, followed by informal explanation.
PLFA introduces an auxiliary relation \texttt{Progress} to capture
progress; an exercise (not shown) asks the reader to show it isomorphic
to the usual formulation with a disjunction and an existential.
Layout is used to present the auxiliary relation in inference rule form.
In Agda, any line beginning with two dashes is treated as a comment, making
it easy to use a line of dashes to separate hypotheses from conclusion
in inference rules.  The proof of proposition \texttt{progress} (the different
case making it a distinct name) is layed out carefully. The neat
indented structure emphasises the case analysis, and all right-hand
sides line-up in the same column.  My hope as an author is that students
will read the formal proof first, and use it as a tabular guide
to the informal explanation that follows.

SF puts the informal explanation first, followed by the formal proof.
The text hides the formal proof script under an icon;
the figure shows what appears when the icon is expanded.
As a teacher I was aware that
students might skip it on a first reading, and I would have to hope the
students would return to it and step through it with an interactive
tool in order to make it intelligible.  I expect the students skipped over
many such proofs.  This particular proof forms the basis for a question
of the mock exam and the past exams, so I expect most students will actually
look at this one if not all the others.

\newcommand{\ex}{\texttt{x}}
\newcommand{\why}{\texttt{y}}
\newcommand{\EL}{\texttt{L}}
\newcommand{\EM}{\texttt{M}}
\newcommand{\EN}{\texttt{N}}
\newcommand{\tee}{\texttt{t}}
\newcommand{\tick}{\texttt{\lq}}
\newcommand{\GG}{\Gamma}
\newcommand{\AY}{\texttt{A}}
\newcommand{\BE}{\texttt{B}}

(For those wanting more detail: In PLFA, variables and abstractions and
applications in the object language are written $\tick\;\ex$ and
$\lambdabar\;\ex\;{\Rightarrow}\;\EN$ and $\EL\;{\cdot}\;\EM$.  The
corresponding typing rules are referred to by ${\vdash}\tick\;()$
and ${\vdash}\lambdabar\;{\vdash}\EN$ and ${\vdash}\EL\;{\cdot}\;{\vdash}\EM$, where
${\vdash}\EL$, ${\vdash}\EM$, ${\vdash}\EN$ are the proofs that terms
$\EL$, $\EM$, $\EN$ are well typed, and `()` denotes that there cannot
be evidence that a free variable is well typed in the empty context.
It was decided to overload infix dot for readability, but not other
symbols. In Agda, as in Lisp, almost any sequence of characters is a
name, with spaces essential for separation.)

(In SF, variables and abstractions and applications in
the object language are written $\texttt{tvar}~\ex$ and
$\texttt{tabs}~\ex~\tee$ and $\texttt{tapp}~\tee_1~\tee_2$.
The corresponding typing rules are referred to as
$\texttt{T\_Var}$ and $\texttt{T\_Abs}$ and $\texttt{T\_App}$.)

Both Coq and Agda support interactive proof.  Interaction in Coq is
supported by Proof General, based on Emacs, or by CoqIDE, which
provides an interactive development environment of a sort familiar to
most students.  Interaction in Agda is supported by an Emacs mode.

In Coq, interaction consists of stepping through a proof script, at
each point examining the current goal and the variables currently in
scope, and executing a new command in the script.  Tactics are a whole
sublanguage, which must be learned in addition to the language for
expressing specifications.  There are many tactics one can invoke in
the script at each point; one menu in CoqIDE lists about one hundred
tactics one might invoke, some in alphabetic submenus.  A Coq
script presents the specification proved and the tactics executed.
Interaction is recorded in a script, which the students
may step through at their leisure.  SF contains some prose descriptions
of stepping through scripts, but mainly contains scripts that students
are encouraged to step through on their own.

In Agda, interaction consists of writing code with holes, at each
point examining the current goal and the variables in scope, and
typing code or executing an Emacs command.  The number of commands
available is much smaller than with Coq, the most important ones being
to show the type of the hole and the types of the variables in scope;
to check the code; to do a case analysis on a given variable; or to
guess how to fill in the hole with constructors or variables in scope.
An Agda proof consists of typed code.  The interaction is \emph{not}
recorded.  Students may recreate it by commenting out bits of code and
introducing a hole in their place.   PLFA contains some prose descriptions
of interactively building code, but mainly contains code that students
can read.  They may also introduce holes to interact with the code, but
I expect this will be rarer than with SF.

SF encourages students to interact with all the scripts in the text.
Trying to understand a Coq proof script without running it
interactively is a bit like understanding a chess game by reading
through the moves without benefit of a board, keeping it all in your
head.  In contrast, PLFA provides code that students can read.
Understanding the code often requires working out the types, but
(unlike executing a Coq proof script) this is often easy to do in your
head; when it is not easy, students still have the option of
interaction.

While students are keen to interact to create code, I have found they
are reluctant to interact to understand code created by others. For
this reason, I suspect this may make Agda a more suitable vehicle for
teaching.  Nate Foster suggests this hypothesis is ripe to be tested
empirically, perhaps using techniques similar to those of
\citet{Danas-et-al-2017}.

Neat layout of definitions such as that in
Figure~\ref{fig:plfa-progress-2} in Emacs requires a monospaced font
supporting all the necessary characters.  Securing one has proved
tricky. As of this writing, we use FreeMono, but it lacks a few
characters ($\typecolon$ and $\qed$) which are loaded from fonts with a different
width.  Long arrows are necessarily more than a single character wide.
Instead, we compose reduction —→ from an em dash — and an arrow →.
Similarly for reflexive and transitive closure —↠.

\section{Progress + Preservation = Evaluation}

\begin{figure}
  \includegraphics[width=\textwidth]{figures/plfa-eval.png}
  \caption{PLFA, Evaluation}
  \label{fig:plfa-eval}
\end{figure}

A standard approach to type soundness used by many texts,
including SF and PLFA, is to prove progress and preservation,
as first suggested by \citet{Wright-and-Felleisen-1994}.

\begin{theorem}[Progress] Given term $M$ and type $A$ such that
$\emptyset \vdash M : A$ then either $M$ is a value or
$M \longrightarrow N$ for some term $N$.
\end{theorem}

\begin{theorem}[Preservation] Given terms $M$ and $N$ and type $A$
such that $\emptyset \vdash M : A$ and $M \longrightarrow N$, then
$\emptyset \vdash N : A$.  \end{theorem}

A consequence is that when a term reduces to a value it retains
the same type.  Further, well-typed terms don't get stuck:
that is, unable to reduce further but not yet reduced to a value.
The formulation neatly accommodates the case of non-terminating
reductions that never reach a value.

One useful by-product of the formal specification of a programming
language may be a prototype implementation of that language.  For
instance, given a language specified by a reduction relation, such
as lambda calculus, the prototype might accept a term and apply reductions
to reduce it to a value.  Typically, one might go to some extra work to
create such a prototype.  For instance, SF introduces a \texttt{normalize}
tactic for this purpose.  Some formal methods frameworks, such as
Redex \citep{Felleisen-et-al-2009} and K \citep{Rosu-Serbanuta-2010},
advertise as one of their advantages that they can generate
a prototype from descriptions of the reduction rules.

I was therefore surprised to realise that any constructive proof of
progress and preservation \emph{automatically} gives rise to such a
prototype.  The input is a term together with evidence the term is
well-typed.  (In the inherently-typed case, these are the same thing.)
Progress determines whether we are done, or should take another step;
preservation provides evidence that the new term is well-typed, so we
may iterate. In a language with guaranteed termination, we cannot
iterate forever, but there are a number of well-known techniques to
address that issue; see, e.g., \citet{Bove-and-Capretta-2001},
\citet{Capretta-2005}, or \citet{McBride-2015}.
We use the simplest, similar to McBride's \emph{petrol-driven} (or
\emph{step-indexed}) semantics: provide a maximum number of steps to
execute; if that number proves insufficient, the evaluator returns the
term it reached, and one can resume execution by providing a new
number.

Such an evaluator from PLFA is shown in Figure~\ref{fig:plfa-eval},
where (inspired by cryptocurrencies) the number of steps
to execute is referred to as \emph{gas}. All of the example reduction
sequences in PLFA were computed by the evaluator and then edited to
improve readability; in addition, the text includes examples of
running the evaluator with its unedited output.

It is immediately obvious that progress and preservation make it
trivial to construct a prototype evaluator, and yet I cannot find such
an observation in the literature nor mentioned in an introductory
text.  It does not appear in SF, nor in \citet{Harper-2016}.  A plea
to the Agda mailing list failed to turn up any prior mentions.
The closest related observation I have seen in the published
literature is that evaluators can be extracted from proofs of
normalisation \citep{Berger-1993,Dagand-and-Scherer-2015}.

(Late addition: My plea to the Agda list eventually bore fruit.  Oleg
Kiselyov directed me to unpublished remarks on his web page where he
uses the name \texttt{eval} for a proof of progress and notes ``the
very proof of type soundness can be used to evaluate sample
expressions'' \citep{Kiselyov-2009}.)

\section{Inherent typing is golden}

\begin{figure}[p]
  \includegraphics[width=\textwidth]{figures/raw.png}
  \caption{Raw approach in PLFA}
  \label{fig:raw}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=\textwidth]{figures/inherent.png}
  \caption{Inherent approach in PLFA}
  \label{fig:inherent}
\end{figure}

The second part of PLFA first discusses two different approaches to
modelling simply-typed lambda calculus.  It first presents raw
terms with named variables and a separate typing relation and
then shifts to inherently-typed terms with de Bruijn indices.
Before writing the text, I had thought the two approaches
complementary, with no clear winner.  Now I am convinced that the
inherently-typed approach is superior.

Figure~\ref{fig:raw} presents the raw approach.
It first defines $\texttt{Id}$, $\texttt{Term}$,
$\texttt{Type}$, and $\texttt{Context}$, the abstract syntax
of identifiers, raw terms, types, and contexts.
It then defines two judgements,
$\GG\:{\ni}\:\ex\:{\typecolon}\:\AY$ and
$\GG\:{\vdash}\:\EM\:{\typecolon}\:\AY$,
which hold when under context $\GG$ the variable $\ex$
and the term $\EM$ have type $\AY$, respectively.

Figure~\ref{fig:inherent} presents the inherent approach.
It first defines $\texttt{Type}$ and $\texttt{Context}$, the abstract syntax
of types and contexts, of which the first is as before and the second is
as before with identifiers dropped.  In place of the two judgements,
the types of variables and terms are indexed by a context and a type,
so that $\GG\:{\ni}\:\AY$ and $\GG\:{\vdash}\:\AY$ denote
variables and terms, respectively, that under context $\GG$ hae type $\AY$.
The indexed types closely resemble the previous judgements:
we now represent a variable or a term by the proof that it is well-typed.
In particular, the proof that a variable is well-typed in the raw approach
corresponds to a de Bruijn index in the inherent approach.

The raw approach requires more lines of code than
the inherent approach.  The separate definition of raw
terms is not needed in the inherent
approach; and one judgement in the raw approach needs to check that
$\ex \not\equiv \why$, while the corresponding judgement in the
inherent approach does not.  The difference becomes more pronounced
when including the code for substitution, reductions, and proofs of
progress and preservation.  In particular, where the raw approach
requires one first define substitution and reduction and then prove
they preserve types, the inherent approach establishes substitution at
the same time it defines substitution and reduction.

Stripping
out examples and any proofs that appear in one but not the other (but
could have appeared in both), the full development in PLFA for the raw approach takes 451
lines (216 lines of definitions and 235 lines for the proofs) and the
development for the inherent approach takes 275 lines (with
definitions and proofs interleaved).  We have 451 / 235 = 1.64,
close to the golden ratio.

The inherent approach also has more expressive power.  The raw
approach is restricted to substitution of one variable by a closed
term, while the inherent approach supports simultaneous substitution
of all variables by open terms, using a pleasing formulation due to
\citet{McBride-2005}, inspired by \citet{Goguen-and-McKinna-1997} and
\citet{Altenkirch-and-Reus-1999} and described in
\citet{Allais-et-al-2017}. In fact, I did manage to write a variant of
the raw approach with simultaneous open substitution along the lines
of McBride, but the result was too complex for use in an introductory
text, requiring 695 lines of code---more than the total for the other
two approaches combined.

The text develops both approaches because the raw approach is more
familiar, and because placing the inherent approach first would
lead to a steep learning curve.  By presenting the more long-winded
but less powerful approach first, students can see for themselves the
advantages of de Bruijn indices and inherent types.

There are actually four possible designs, as the choice of named
variables vs de Bruijn indices, and the choice of raw vs
inherently-typed terms may be made independently.  There are synergies
beween the two.  Manipulation of de Bruijn indices can be notoriously
error-prone without inherent-typing to give assurance of correctness.
In inherent typing with named variables, simultaneous substitution by
open terms remains difficult.

The benefits of the inherent approach 
are well known to some. The technique was introduced by
\citet{Altenkirch-and-Reus-1999}, and widely used elsewhere,
notably by \citet{Chapman-2009} and \citet{Allais-et-al-2017}.
I'm grateful to David Darais for bringing it to my attention.

\section{Conclusion}

I look forward to experience teaching from the new text,
and encourage others to use it too.  Please comment!

\paragraph{Acknowledgements}

A special thank you to my coauthor, Wen Kokke.  For inventing ideas on
which PLFA is based, and for hand-holding, many thanks to Conor
McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing me
how much more compact it is to avoid raw terms, thanks to David
Darais.  For inspiring my work by writing SF, thanks to Benjamin
Pierce and his coauthors.  For comments on a draft of this paper, an
extra thank you to James McKinna, Ulf Norell, Andreas Abel, and
Benjamin Pierce. This research was supported by EPSRC Programme Grant
EP/K034413/1.

\bibliographystyle{plainnat}
%\bibliographystyle{splncsnat.bst}
\bibliography{PLFA}

\end{document}

### 2019-07-01 sbmf/sbmf/latex/PLFA.tex
Making a subdirectory of sbmf

--- 2019-07-01 sbmf/sbmf/PLFA.tex
@@
 \usepackage{agda}
-\usepackage{revsymb}
-\usepackage{semantic}
 \usepackage{graphicx}
@@
 \setcitestyle{round,aysep={}}
+
+\DeclareUnicodeCharacter{"2982}{\ensuremath{\typecolon}}  % ⦂
+\DeclareUnicodeCharacter{"220E}{\ensuremath{\QED}}  % ∎
 
@@
 \begin{abstract}
-One of the leading textbooks for formal methods is
-\emph{Software Foundations} (SF), written by Benjamin Pierce in
-collaboration with others, and based on Coq. After five years using SF
-in the classroom, I have come to the conclusion that Coq is not the
-best vehicle for this purpose, as too much of the course needs to
-focus on learning tactics for proof derivation, to the cost of
-learning programming language theory.  Accordingly, I have written a
-new textbook, \emph{Programming Language Foundations in Agda} (PLFA).
-PLFA covers much of the same ground as SF, although it is not a
-slavish imitation.
+The leading textbook for formal methods is \emph{Software Foundations}
+(SF), written by Benjamin Pierce in collaboration with others, and
+based on Coq. After five years using SF in the classroom, I have come
+to the conclusion that Coq is not the best vehicle for this purpose,
+as too much of the course needs to focus on learning tactics for proof
+derivation, to the cost of learning programming language theory.
+Accordingly, I have written a new textbook, \emph{Programming Language
+Foundations in Agda} (PLFA).  PLFA covers much of the same ground as
+SF, although it is not a slavish imitation.
 
@@
 Foundations in Agda'' or ``Programming (Language Foundations) in
-Agda''---specifications in the proof assistant Agda both describe
+Agda''---the specifications in the proof assistant Agda both describe
 programming languages and are themselves programmes.
@@
 writing the book, it was not obvious that it was even possible;
-conceivably, without tactics some of the proofs might
+conceivably, without tactics some of the proofs
 balloon in size. In fact, it turns out that for the results in PLFA
@@
 
-\paragraph{Decidable: Booleans and decision procedures}
-Introduces booleans and decidable types, and why the latter is to be
-preferred to the former.
-
 \paragraph{Lists: Lists and higher-order functions}
@@
 
+\paragraph{Decidable: Booleans and decision procedures}
+Introduces booleans and decidable types, and why the latter is to be
+preferred to the former.
+
 \subsection*{Part II, Programming Language Foundations}
@@
 Introduces lambda calculus, using a representation with named
-variables and a separate typing relation. The language used is PCF
-\citep{Plotkin-1977}, with variables, lambda abstraction, application,
-zero, successor, case over naturals, and fixpoint. Reduction is
-call-by-value and restricted to closed terms.
+variables and a separate typing relation. The language used is PCF,
+with variables, lambda abstraction, application, zero, successor, case
+over naturals, and fixpoint. Reduction is call-by-value and restricted
+to closed terms.
 
@@
 \paragraph{DeBruijn: Inherently typed de Bruijn representation}
-Introduces de Bruijn indices and the inherently-typed representation.
+Introduces de Bruijn numbers and the inherently-typed representation.
 Emphasis is put on the structural similarity between a term and its
-corresponding type derivation; in particular, de Bruijn indices
+corresponding type derivation; in particular, de Bruijn numbers
 correspond to the judgment that a variable is well-typed under a given
@@
 a representation with named variables and a separate typing relation
-to a representation de Bruijn indices with inherent types. Bidirectional
-type inference is shown to be both sound and complete.
+to a representation de Bruijn indices with inherent types.
 
@@
 uses type checking.  SF also covers a simple imperative language with Hoare logic,
-and for lambda calculus covers subtyping, record types, mutable references, and normalisation
----none of which are treated by PLFA.  PLFA covers an inherently-typed de Bruijn
+subtyping, record types, mutable references, and normalisation none of which are
+treated by PLFA.  PLFA covers an inherently-typed de Bruijn
 representation, bidirectional type inference, bisimulation, and an
-untyped call-by-name language with full normalisation---none of which
+untyped call-by-value language with full normalisation, none of which
 are treated by SF.
@@
 on whether it would be easy or hard to cover that material in Agda.
-And SF recently added a fourth volume on random testing of Coq specifications
-using QuickChick.  There is currently no tool equivalent to QuickChick
-available for Agda.
 
@@
 used in the book.  My attempts so far to formalise pure type systems
-have proved challenging, to say the least.
+have proved challenging.
 
@@
 central proof, that substitution preserves types, is carried out by
-induction on terms for a technical reason
+induction on terms, not evidence of typing, for a technical reason
 (the context is extended by a variable binding, and hence not
@@
 since swapping the arguments doesn't yield a recursive call on
-structurally smaller arguments.  I suspect tactics could cut down the
+structurally smaller arguments.  (I suspect tactics could cut down the
 proof significantly. I tried to compare with SF's proof that reduction
-is deterministic, but failed to find that proof.
+is deterministic, but failed to find that proof.)
 
@@
 PLFA puts the formal statements first, followed by informal explanation.
-PLFA introduces an auxiliary relation \texttt{Progress} to capture
+PLFA introduces an auxiliary relation \texttt{Progress M} to capture
 progress; an exercise (not shown) asks the reader to show it isomorphic
@@
 it easy to use a line of dashes to separate hypotheses from conclusion
-in inference rules.  The proof of proposition \texttt{progress} (the different
-case making it a distinct name) is layed out carefully. The neat
-indented structure emphasises the case analysis, and all right-hand
-sides line-up in the same column.  My hope as an author is that students
+in inference rules.  The proof is layed out carefully, with a neat
+indented structure to emphasise the case analysis, and all right-hand
+sides lined-up in the same column.  My hope as an author is that students
 will read the formal proof first, and use it as a tabular guide
@@
 The text hides the formal proof script under an icon;
-the figure shows what appears when the icon is expanded.
+the figures shows what appears when the icon is expanded.
 As a teacher I was aware that
@@
 look at this one if not all the others.
-
-\newcommand{\ex}{\texttt{x}}
-\newcommand{\why}{\texttt{y}}
-\newcommand{\EL}{\texttt{L}}
-\newcommand{\EM}{\texttt{M}}
-\newcommand{\EN}{\texttt{N}}
-\newcommand{\tee}{\texttt{t}}
-\newcommand{\tick}{\texttt{\lq}}
-\newcommand{\GG}{\Gamma}
-\newcommand{\AY}{\texttt{A}}
-\newcommand{\BE}{\texttt{B}}
-
-(For those wanting more detail: In PLFA, variables and abstractions and
-applications in the object language are written $\tick\;\ex$ and
-$\lambdabar\;\ex\;{\Rightarrow}\;\EN$ and $\EL\;{\cdot}\;\EM$.  The
-corresponding typing rules are referred to by ${\vdash}\tick\;()$
-and ${\vdash}\lambdabar\;{\vdash}\EN$ and ${\vdash}\EL\;{\cdot}\;{\vdash}\EM$, where
-${\vdash}\EL$, ${\vdash}\EM$, ${\vdash}\EN$ are the proofs that terms
-$\EL$, $\EM$, $\EN$ are well typed, and `()` denotes that there cannot
-be evidence that a free variable is well typed in the empty context.
-It was decided to overload infix dot for readability, but not other
-symbols. In Agda, as in Lisp, almost any sequence of characters is a
-name, with spaces essential for separation.)
-
-(In SF, variables and abstractions and applications in
-the object language are written $\texttt{tvar}~\ex$ and
-$\texttt{tabs}~\ex~\tee$ and $\texttt{tapp}~\tee_1~\tee_2$.
-The corresponding typing rules are referred to as
-$\texttt{T\_Var}$ and $\texttt{T\_Abs}$ and $\texttt{T\_App}$.)
 
@@
 scope, and executing a new command in the script.  Tactics are a whole
-sublanguage, which must be learned in addition to the language for
+sublanguage, which must be learned in addition to the langauge for
 expressing specifications.  There are many tactics one can invoke in
 the script at each point; one menu in CoqIDE lists about one hundred
-tactics one might invoke, some in alphabetic submenus.  A Coq
+tactics one might invoked, some in alphabetic submenus.  A Coq
 script presents the specification proved and the tactics executed.
@@
 tricky. As of this writing, we use FreeMono, but it lacks a few
-characters ($\typecolon$ and $\qed$) which are loaded from fonts with a different
+characters (⦂ and ∎) which are loaded from fonts with a different
 width.  Long arrows are necessarily more than a single character wide.
@@
 Such an evaluator from PLFA is shown in Figure~\ref{fig:plfa-eval},
-where (inspired by cryptocurrencies) the number of steps
+where (inspired by McBride and cryptocurrencies) the number of steps
 to execute is referred to as \emph{gas}. All of the example reduction
@@
 very proof of type soundness can be used to evaluate sample
-expressions'' \citep{Kiselyov-2009}.)
+expressions'' \citet{Kiselyov-2009}.)
 
@@
 
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{figures/raw.png}
-  \caption{Raw approach in PLFA}
-  \label{fig:raw}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=\textwidth]{figures/inherent.png}
-  \caption{Inherent approach in PLFA}
-  \label{fig:inherent}
-\end{figure}
-
 The second part of PLFA first discusses two different approaches to
-modelling simply-typed lambda calculus.  It first presents raw
+modelling simply-typed lambda calculus.  It first presents ``raw''
 terms with named variables and a separate typing relation and
@@
 
-Figure~\ref{fig:raw} presents the raw approach.
-It first defines $\texttt{Id}$, $\texttt{Term}$,
-$\texttt{Type}$, and $\texttt{Context}$, the abstract syntax
-of identifiers, raw terms, types, and contexts.
-It then defines two judgements,
-$\GG\:{\ni}\:\ex\:{\typecolon}\:\AY$ and
-$\GG\:{\vdash}\:\EM\:{\typecolon}\:\AY$,
-which hold when under context $\GG$ the variable $\ex$
-and the term $\EM$ have type $\AY$, respectively.
+The clearest indication comes from counting lines of code.
+Stripping out examples and any proofs that appear in one but not the
+other, the development for raw terms takes 451 lines (216 lines of
+definitions and 235 lines for the proofs) and the development for
+inherently typed terms takes 275 lines (with definitions and proofs
+interleaved, as substitution cannot be defined without
+also showing it preserves types).  We have 451 / 235 = 1.64, close
+to the golden ratio.
 
-Figure~\ref{fig:inherent} presents the inherent approach.
-It first defines $\texttt{Type}$ and $\texttt{Context}$, the abstract syntax
-of types and contexts, of which the first is as before and the second is
-as before with identifiers dropped.  In place of the two judgements,
-the types of variables and terms are indexed by a context and a type,
-so that $\GG\:{\ni}\:\AY$ and $\GG\:{\vdash}\:\AY$ denote
-variables and terms, respectively, that under context $\GG$ hae type $\AY$.
-The indexed types closely resemble the previous judgements:
-we now represent a variable or a term by the proof that it is well-typed.
-In particular, the proof that a variable is well-typed in the raw approach
-corresponds to a de Bruijn index in the inherent approach.
-
-The raw approach requires more lines of code than
-the inherent approach.  The separate definition of raw
-terms is not needed in the inherent
-approach; and one judgement in the raw approach needs to check that
-$\ex \not\equiv \why$, while the corresponding judgement in the
-inherent approach does not.  The difference becomes more pronounced
-when including the code for substitution, reductions, and proofs of
-progress and preservation.  In particular, where the raw approach
-requires one first define substitution and reduction and then prove
-they preserve types, the inherent approach establishes substitution at
-the same time it defines substitution and reduction.
-
-Stripping
-out examples and any proofs that appear in one but not the other (but
-could have appeared in both), the full development in PLFA for the raw approach takes 451
-lines (216 lines of definitions and 235 lines for the proofs) and the
-development for the inherent approach takes 275 lines (with
-definitions and proofs interleaved).  We have 451 / 235 = 1.64,
-close to the golden ratio.
-
-The inherent approach also has more expressive power.  The raw
-approach is restricted to substitution of one variable by a closed
-term, while the inherent approach supports simultaneous substitution
-of all variables by open terms, using a pleasing formulation due to
-\citet{McBride-2005}, inspired by \citet{Goguen-and-McKinna-1997} and
-\citet{Altenkirch-and-Reus-1999} and described in
-\citet{Allais-et-al-2017}. In fact, I did manage to write a variant of
-the raw approach with simultaneous open substitution along the lines
-of McBride, but the result was too complex for use in an introductory
+Another indication is expressive power. The approach with named
+variables and separate typing is restricted to substitution of one
+variable by a single closed term, while de Bruijn indices with
+inherent typing support simultaneous substitution of all variables by
+open terms, using a pleasing formulation due to \citet{McBride-2005},
+inspired by \citet{Goguen-and-McKinna-1997} and described
+in \citet{Allais-et-al-2017}. In fact, I managed to extend
+McBride's approach to support raw terms with named variables, but the
+resulting code was too long and too complex for use in an introductory
 text, requiring 695 lines of code---more than the total for the other
@@
 
-The text develops both approaches because the raw approach is more
-familiar, and because placing the inherent approach first would
-lead to a steep learning curve.  By presenting the more long-winded
-but less powerful approach first, students can see for themselves the
-advantages of de Bruijn indices and inherent types.
+The text develops both approaches because named variables with
+separate typing is more familiar, and placing de Bruijn indices and
+inherent typing first would lead to a steep learning curve.  By
+presenting the more long-winded but less powerful approach first,
+students can see for themselves the advantages of de Bruijn indices
+with inherent typing.
 
@@
 
-The benefits of the inherent approach 
-are well known to some. The technique was introduced by
+The benefits of replacing raw typing by inherent are well
+known to some. The technique was introduced by
 \citet{Altenkirch-and-Reus-1999}, and widely used elsewhere,
@@
 extra thank you to James McKinna, Ulf Norell, Andreas Abel, and
-Benjamin Pierce. This research was supported by EPSRC Programme Grant
-EP/K034413/1.
+Benjamin Pierce.
 

### 2019-07-01 papers/scp/llncs2e/samplepaper.tex
Moved sbmf/sbmf to papers/sbmf and copied over for scp

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Contribution Title\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2019-07-01 papers/scp/PLFA.tex
Updated .gitignore; rewrote SCP paper to use elsarticle class

--- 2019-07-01 sbmf/sbmf/PLFA.tex
@@
-%\documentclass[runningheads,oribibl]{llncs}
-\documentclass[runningheads]{llncs}
-\usepackage{natbib}
-% \usepackage{stmaryrd}
-% \usepackage{proof}
-% \usepackage{amsmath}
-% \usepackage{amssymb}
-% \usepackage{color}
-% \usepackage{bbm}
-% \usepackage[greek,english]{babel}
-% \usepackage{ucs}
-% \usepackage[utf8x]{inputenc}
-% \usepackage{autofe}
+\documentclass[preprint,authoryear]{elsarticle}
+%\usepackage{natbib}
 \usepackage{agda}
@@
 \usepackage{stix}
-\setcitestyle{round,aysep={}}
+\usepackage{amsmath,amssymb,amsthm}
+\newtheorem{theorem}{Theorem}
+\usepackage{mathpazo}
+\usepackage[mathpazo]{flexisym}
 
 \begin{document}
+
 \title{Programming Language Foundations in Agda}
-\author{Philip Wadler}
-\authorrunning{P. Wadler}
-\institute{University of Edinburgh\\
-  \email{wadler@inf.ed.ac.uk}}
-\maketitle
-%
+
+\author[adr1]{Wen Kokke}
+\ead{wen.kokke@ed.ac.uk}
+\author[adr2]{Jeremy Siek}
+\ead{jsiek@indiana.edu}
+\author[adr1]{Philip Wadler\corref{cor1}}
+\ead{wadler@inf.ed.ac.uk}
+\cortext[cor1]{Corresponding author}
+
+\address[adr1]{University of Edinburgh, 10 Crichton Street, EH8 9AB, Edinburgh}
+\address[adr2]{Indiana University, 700 N Woodlawn Ave, Bloomington, IN 47408, USA}
+
 \begin{abstract}
-One of the leading textbooks for formal methods is
-\emph{Software Foundations} (SF), written by Benjamin Pierce in
-collaboration with others, and based on Coq. After five years using SF
-in the classroom, I have come to the conclusion that Coq is not the
-best vehicle for this purpose, as too much of the course needs to
-focus on learning tactics for proof derivation, to the cost of
-learning programming language theory.  Accordingly, I have written a
-new textbook, \emph{Programming Language Foundations in Agda} (PLFA).
-PLFA covers much of the same ground as SF, although it is not a
-slavish imitation.
+  One of the leading textbooks for formal methods is \emph{Software Foundations} (SF), written by Benjamin Pierce in collaboration with others, and based on Coq.
+After five years using SF in the classroom, I have come to the conclusion that Coq is not the best vehicle for this purpose, as too much of the course needs to focus on learning tactics for proof derivation, to the cost of learning programming language theory.
+Accordingly, I have written a new textbook, \emph{Programming Language Foundations in Agda} (PLFA).
+PLFA covers much of the same ground as SF, although it is not a slavish imitation.
 
-What did I learn from writing PLFA? First, that it is possible. One
-might expect that without proof tactics that the proofs become too
-long, but in fact proofs in PLFA are about the same length as those in
-SF. Proofs in Coq require an interactive environment to be understood,
-while proofs in Agda can be read on the page.  Second, that
-constructive proofs of preservation and progress give immediate rise
-to a prototype evaluator. This fact is obvious in retrospect but it is
-not exploited in SF (which instead provides a separate normalise
-tactic) nor can I find it in the literature.  Third, that using raw
-terms with a separate typing relation is far less perspicuous than
-using inherently-typed terms. SF uses the former presentation, while
-PLFA presents both; the former uses about 1.6 as many lines of Agda
-code as the latter, roughly the golden ratio.
+  What did I learn from writing PLFA?
+First, that it is possible.
+One might expect that without proof tactics that the proofs become too long, but in fact proofs in PLFA are about the same length as those in SF.
+Proofs in Coq require an interactive environment to be understood, while proofs in Agda can be read on the page.
+Second, that constructive proofs of preservation and progress give immediate rise to a prototype evaluator.
+This fact is obvious in retrospect but it is not exploited in SF (which instead provides a separate normalise tactic) nor can I find it in the literature.
+Third, that using raw terms with a separate typing relation is far less perspicuous than using inherently-typed terms.
+SF uses the former presentation, while PLFA presents both; the former uses about 1.6 as many lines of Agda code as the latter, roughly the golden ratio.
 
@@
   \end{center}
-\keywords{Agda \and Coq \and lambda calculus \and dependent types.}
 \end{abstract}
+
+\begin{keyword}
+  Agda \sep Coq \sep lambda calculus \sep dependent types.  
+\end{keyword}
+
+\maketitle
 
@@
 tricky. As of this writing, we use FreeMono, but it lacks a few
-characters ($\typecolon$ and $\qed$) which are loaded from fonts with a different
+characters ($\typecolon$ and $\qedsymbol$) which are loaded from fonts with a different
 width.  Long arrows are necessarily more than a single character wide.
@@
 
-\bibliographystyle{plainnat}
-%\bibliographystyle{splncsnat.bst}
+\section*{References}
+
+\bibliographystyle{elsarticle-harv}
 \bibliography{PLFA}

### 2019-07-11 papers/scp/PLFA.tex
progress on my updates to section 2

@@
 \ead{wen.kokke@ed.ac.uk}
-\author[adr2]{Jeremy Siek}
+\author[adr2]{Jeremy G. Siek}
 \ead{jsiek@indiana.edu}
@@
 evidence that it is in normal form.
+
+\paragraph{Substitution: in the untyped lambda calculus}
+Delves deeper into the properties of simultaneous substitution,
+establishing the equations of the $\sigma$ algebra of
+\citet{Abadi-1991}. These equations enable a straightforward proof of
+the Substitution Lemma~\citep{Barendregt-1984}, which is needed in the
+next chapter.
+
+\paragraph{Confluence: of the untyped lambda calculus}
+Presents a proof of the Church-Rosser theorem based on the classic
+idea of parallel reduction due to Tait and Martin-L\"of. The proof in
+Agda is streamlined by the use of ideas from
+\citet{Schafer-Tebbi-Smolka-2015} and \citet{Pfenning-1992}.
+
+\paragraph{Big-step: evaluation for call-by-name}
+Introduces the notion of big-step evaluation to develop a
+deterministic call-by-name reduction strategy. The main result of this
+chapter is a proof that big-step evaluation implies the existence of a
+reduction sequence that terminates with a lambda abstraction.
+
+\subsection*{Part III, Denotational Semantics}
+
+\paragraph{Denotational: semantics of the untyped lambda calculus}
+The early denotational semantics of the lambda calculus based on graph
+models~\citep{Scott-1976,Engeler-1981,Plotkin-1993} and filter
+models~\citep{Barendregt-1983} were particularly simple and elegant.
+This chapter presents such a semantics in a big-step style that is
+approachable to readers familiar with operational semantics.
+
+\paragraph{Compositional: the denotational semantics is compositional}
+TODO
+
+\paragraph{Soundness: of reduction with respect to denotational semantics}
+TODO
+
+\paragraph{Adequacy: of denotational semantics with respect to reduction}
+TODO
+
+\paragraph{Contextual Equivalence: is implied by denotational equality}
+TODO
 

### 2019-07-12 papers/scp/PLFA.tex
additions to section 2, and spell check

@@
 Agda.  \emph{Verified Functional Programming in Agda} by \citep{Stump-2016}
-covers related ground, but focusses more on programming with dependent
+covers related ground, but focuses more on programming with dependent
 types than on the theory of programming languages.
@@
 and accessible to students, and give them rapid feedback to find
-and correct misaprehensions.
+and correct misapprehensions.
 
@@
 \paragraph{Big-step: evaluation for call-by-name}
-Introduces the notion of big-step evaluation to develop a
-deterministic call-by-name reduction strategy. The main result of this
-chapter is a proof that big-step evaluation implies the existence of a
-reduction sequence that terminates with a lambda abstraction.
+Introduces the notion of big-step evaluation, written $\gamma \vdash M
+\Downarrow V$, to develop a deterministic call-by-name reduction
+strategy. The main result of this chapter is a proof that big-step
+evaluation implies the existence of a reduction sequence that
+terminates with a lambda abstraction.
 
@@
 models~\citep{Scott-1976,Engeler-1981,Plotkin-1993} and filter
-models~\citep{Barendregt-1983} were particularly simple and elegant.
-This chapter presents such a semantics in a big-step style that is
-approachable to readers familiar with operational semantics.
+models~\citep{Barendregt-1983} were particularly simple and elegant: a
+function is modeled as a lookup table.  This chapter presents such a
+semantics using a big-step notation that is approachable to readers
+familiar with operational semantics, writing $\gamma \vdash M
+\downarrow d$ for the evaluation of a term $M$ to denotation $d$ in
+environment $\gamma$.
 
 \paragraph{Compositional: the denotational semantics is compositional}
-TODO
+The hallmark of denotational semantics is that they are compositional:
+the meaning of each language form is a function of the meanings of its
+parts.
+%
+We define two functions, named \texttt{curry} and \texttt{apply} that
+serve this purpose for lambda abstraction and application.
+%
+The results in this chapter include congruences for \texttt{curry} and
+\texttt{apply} and the compositionality property for the denotational
+semantics.
+%
+The chapter concludes with a functional definition of the denotational
+semantics and a proof that it is equivalent to the big-step version.
 
 \paragraph{Soundness: of reduction with respect to denotational semantics}
-TODO
+Reduction implies denotational equality. We prove each direction of
+the equality, first showing the reduction preserves denotations
+(subject reduction), and then showing that reduction reflects
+denotations (subject expansion). The first proof is similar to the
+type preservation proofs in Part II. The second goes in reverse,
+showing that if $M \longrightarrow N$ and $\gamma \vdash N
+\downarrow d$, then $\gamma \vdash M \downarrow d$ .
 
 \paragraph{Adequacy: of denotational semantics with respect to reduction}
-TODO
+If a term is denotationally equal to a lambda expression, then it
+reduces to a lambda expression. The main lemma shows that if a term's
+denotation is functional, i.e., $\gamma \vdash M \downarrow (d \mapsto
+d')$, then $M$ terminates according to the call-by-name big-step
+semantics, i.e., $\gamma' \vdash M \Downarrow V$. A logical relation
+$\mathbb{V}$ is used to relate denotations and values (i.e. closures).
+The implication from the big-step semantics to reduction is proved in
+the Big-step chapter of Part II.
 
 \paragraph{Contextual Equivalence: is implied by denotational equality}
-TODO
+The main criteria for behavior-preserving code transformation (such as
+compiler optimization or programmer refactoring) is contextual
+equivalence. Two terms are contextually equivalent when they can both
+be placed into an arbitrary context (a program with a hole) and the
+resulting programs behave the same (e.g., they both terminate or they
+both diverge). This chapter ties together the previous results
+(Compositionality, Soundness, and Adequacy) to show that denotational
+equality implies contextual equivalence. Thus, it is safe to use
+denotational equality to justify code transformations.
 
@@
 The second part of PLFA first discusses two different approaches to
-modelling simply-typed lambda calculus.  It first presents raw
+modeling simply-typed lambda calculus.  It first presents raw
 terms with named variables and a separate typing relation and
@@
 of identifiers, raw terms, types, and contexts.
-It then defines two judgements,
+It then defines two judgments,
 $\GG\:{\ni}\:\ex\:{\typecolon}\:\AY$ and
@@
 of types and contexts, of which the first is as before and the second is
-as before with identifiers dropped.  In place of the two judgements,
+as before with identifiers dropped.  In place of the two judgments,
 the types of variables and terms are indexed by a context and a type,
 so that $\GG\:{\ni}\:\AY$ and $\GG\:{\vdash}\:\AY$ denote
-variables and terms, respectively, that under context $\GG$ hae type $\AY$.
-The indexed types closely resemble the previous judgements:
+variables and terms, respectively, that under context $\GG$ have type $\AY$.
+The indexed types closely resemble the previous judgments:
 we now represent a variable or a term by the proof that it is well-typed.
@@
 terms is not needed in the inherent
-approach; and one judgement in the raw approach needs to check that
-$\ex \not\equiv \why$, while the corresponding judgement in the
+approach; and one judgment in the raw approach needs to check that
+$\ex \not\equiv \why$, while the corresponding judgment in the
 inherent approach does not.  The difference becomes more pronounced
@@
 inherently-typed terms may be made independently.  There are synergies
-beween the two.  Manipulation of de Bruijn indices can be notoriously
+between the two.  Manipulation of de Bruijn indices can be notoriously
 error-prone without inherent-typing to give assurance of correctness.
@@
 
-\paragraph{Acknowledgements}
+\paragraph{Acknowledgments}
 
@@
 \end{document}
+
+% LocalWords:  Agda adr Kokke Siek Wadler Woodlawn Bloomington Coq et
+% LocalWords:  PLFA evaluator normalise programme programmes al Huet
+% LocalWords:  ICFP subst Bove Mixfix de Janeiro summarising favour
+% LocalWords:  organised formalising Colour Gonthier CompCert Kastner
+% LocalWords:  SEL honours formalisms monus associativity reflexivity
+% LocalWords:  commutativity equational of's Isomorphism isomorphism
+% LocalWords:  extensionality disjunction intuitionistic existentials
+% LocalWords:  Booleans booleans monoid specialisation PCF Plotkin EP
+% LocalWords:  fixpoint DeBruijn Bruijn indices Bisimulation Untyped
+% LocalWords:  bisimulation normalisation untyped Abadi Barendregt
+% LocalWords:  Rosser Tait Schafer Tebbi Smolka Pfenning Denotational
+% LocalWords:  denotational Engeler lookup Compositional congruences
+% LocalWords:  compositional compositionality denotationally Hoare
+% LocalWords:  refactoring subtyping Appel QuickChick formalise SF's
+% LocalWords:  Stlc StlcProp Coq's reinvoke Agda's postconditions Ulf
+% LocalWords:  flavour layed emphasises tvar tapp CoqIDE sublanguage
+% LocalWords:  submenus Danas monospaced FreeMono Felleisen Redex
+% LocalWords:  Rosu Serbanuta realise Capretta cryptocurrencies Oleg
+% LocalWords:  evaluators Dagand Scherer Kiselyov eval Goguen McKinna
+% LocalWords:  Altenkirch Reus Allais Darais Conor Norell Andreas
+% LocalWords:  EPSRC

### 2019-07-12 highlight.sh
Setup courses folder.

@@ -9,17 +9,26 @@ OUT="$1"
 OUT_DIR="$(dirname $OUT)"
 shift
 
+# Extract the module name from the Agda file
+# NOTE: this fails if there is more than a single space after 'module'
+MOD_NAME=`grep -oP -m 1 "(?<=^module )(\\S+)(?=\\s+(\\S+\\s+)*where)" "$SRC"`
+
 # Create temporary directory and compute path to output of `agda --html`
-# NOTE: this assumes $OUT is equivalent to out/ plus the module path
 HTML_DIR="$(mktemp -d)"
-HTML="${OUT#out/}"
-HTML="/${HTML//\//.}"
-HTML="$HTML_DIR/$HTML"
+SRC_EXT="$(basename $SRC)"
+SRC_EXT="${SRC_EXT##*.}"
+HTML="$HTML_DIR/$MOD_NAME.$SRC_EXT"
 
 # Highlight Syntax using Agda
 set -o pipefail \
    && agda --html --html-highlight=code --html-dir="$HTML_DIR" "$SRC" "$@" \
     | sed '/^Generating.*/d; /^Warning\: HTML.*/d; /^reached from the.*/d; /^\s*$/d'
+
+# Check if the highlighted file was successfully generated
+if [[ ! -f "$HTML" ]]; then
+    echo "File not generated: $FILE"
+    exit 1
+fi
 
 # Add source file to the Jekyll metadata
 sed -i "1 s|---|---\nsrc: $SRC|" "$HTML"

### 2019-07-14 papers/scp/PLFA.tex
Rewrote first person in sections 2 and onwards

@@
 extended to cover additional material, including mutable references,
-normalisation, System F, pure type systems, and denotational
-semantics.  I'd especially like to include pure type systems as they
+normalisation, System F, and pure type systems.
+We'd especially like to include pure type systems as they
 provide the readers with a formal model close to the dependent types
-used in the book.  My attempts so far to formalise pure type systems
+used in the book.  Our attempts so far to formalise pure type systems
 have proved challenging, to say the least.
@@
 
-% \begin{figure}[t]
-%   \includegraphics[width=\textwidth]{figures/sf-progress-3.png}
-%   \caption{SF, Progress (3/3)}
-%   \label{fig:sf-progress-3}
-% \end{figure}
-
 The introduction listed several reasons for preferring Agda over Coq.
@@
 
-As an experiment, I first rewrote SF's development of simply-typed
-lambda calculus (SF, Chapters Stlc and StlcProp) in Agda.  I was a
-newbie to Agda, and translating the entire development, sticking as
-closely as possible to the development in SF, took me about two days.
-I was pleased to discover that the proofs remained about the same
-size.
+As an experiment, Philip first rewrote SF's development of simply-typed lambda
+calculus (SF, Chapters Stlc and StlcProp) in Agda.  He was a newbie to Agda, and
+translating the entire development, sticking as closely as possible to the
+development in SF, took about two days.  We were pleased to discover that the
+proofs remained about the same size.
 
@@
 sufficiently ``generic'' to work well with Coq's induction tactic). In
-Agda, I had no trouble formulating the same proof over evidence that
+Agda, we had no trouble formulating the same proof over evidence that
 the term is well typed, and didn't even notice SF's description of the
@@
 impossible.  Five of the ten cases are redundant, as they just involve
-switching the order of the arguments.  I had to copy the cases
+switching the order of the arguments.  We had to copy the cases
 suitably permuted. It would be preferable to reinvoke the proof on
@@
 since swapping the arguments doesn't yield a recursive call on
-structurally smaller arguments.  I suspect tactics could cut down the
-proof significantly. I tried to compare with SF's proof that reduction
+structurally smaller arguments.  We suspect tactics could cut down the
+proof significantly. We tried to compare with SF's proof that reduction
 is deterministic, but failed to find that proof.
@@
 ``omega'' tactic invoking a decision procedure.  The entire
-exercise would be easy to repeat in Agda, save for the last step: I
+exercise would be easy to repeat in Agda, save for the last step: we
 suspect Agda's automation would not be up to verifying the generated
-conditions, requiring tedious proofs by hand.  However, I had already
+conditions, requiring tedious proofs by hand.  However, we had already
 decided to omit Hoare logic in order to focus on lambda calculus.
 
-To give a flavour of how the texts compare, I show the
+To give a flavour of how the texts compare, we show the
 proof of progress for simply-typed lambda calculus from both texts.
@@
 indented structure emphasises the case analysis, and all right-hand
-sides line-up in the same column.  My hope as an author is that students
+sides line-up in the same column.  Our hope as authors is that students
 will read the formal proof first, and use it as a tabular guide
@@
 
-SF puts the informal explanation first, followed by the formal proof.
-The text hides the formal proof script under an icon;
-the figure shows what appears when the icon is expanded.
-As a teacher I was aware that
-students might skip it on a first reading, and I would have to hope the
-students would return to it and step through it with an interactive
-tool in order to make it intelligible.  I expect the students skipped over
-many such proofs.  This particular proof forms the basis for a question
-of the mock exam and the past exams, so I expect most students will actually
-look at this one if not all the others.
+SF puts the informal explanation first, followed by the formal proof. The text
+hides the formal proof script under an icon; the figure shows what appears when
+the icon is expanded. As teachers, we were aware that students might skip the
+formal proof on a first reading, and we would have to hope the students would
+return to it and step through it with an interactive tool in order to make it
+intelligible.  We expect the students skipped over many such proofs.  This
+particular proof forms the basis for a question of the mock exam and the past
+exams, so we expect most students will actually look at this one if not all the
+others.
 
@@
 can read.  They may also introduce holes to interact with the code, but
-I expect this will be rarer than with SF.
+we expect this will be rarer than with SF.
 
@@
 
-While students are keen to interact to create code, I have found they
+While students are keen to interact to create code, we have found they
 are reluctant to interact to understand code created by others. For
-this reason, I suspect this may make Agda a more suitable vehicle for
+this reason, we suspect this may make Agda a more suitable vehicle for
 teaching.  Nate Foster suggests this hypothesis is ripe to be tested
@@
 
-I was therefore surprised to realise that any constructive proof of
+% TODO: rewrite use of 'Philip' to reflect all authors?
+Philip was therefore surprised to realise that any constructive proof of
 progress and preservation \emph{automatically} gives rise to such a
@@
 It is immediately obvious that progress and preservation make it
-trivial to construct a prototype evaluator, and yet I cannot find such
+trivial to construct a prototype evaluator, and yet we cannot find such
 an observation in the literature nor mentioned in an introductory
@@
 
-(Late addition: My plea to the Agda list eventually bore fruit.  Oleg
+(Late addition: Our plea to the Agda list eventually bore fruit.  Oleg
 Kiselyov directed me to unpublished remarks on his web page where he
@@
 
+% TODO: consider rewriting this?
 The second part of PLFA first discusses two different approaches to
@@
 then shifts to inherently-typed terms with de Bruijn indices.
-Before writing the text, I had thought the two approaches
-complementary, with no clear winner.  Now I am convinced that the
+Before writing the text, Philip had thought the two approaches
+complementary, with no clear winner.  Now he is convinced that the
 inherently-typed approach is superior.
@@
 The indexed types closely resemble the previous judgments:
-we now represent a variable or a term by the proof that it is well-typed.
-In particular, the proof that a variable is well-typed in the raw approach
+we now represent a variable or a term by the proof that it is well typed.
+In particular, the proof that a variable is well typed in the raw approach
 corresponds to a de Bruijn index in the inherent approach.
 
-The raw approach requires more lines of code than
-the inherent approach.  The separate definition of raw
-terms is not needed in the inherent
-approach; and one judgment in the raw approach needs to check that
-$\ex \not\equiv \why$, while the corresponding judgment in the
-inherent approach does not.  The difference becomes more pronounced
-when including the code for substitution, reductions, and proofs of
-progress and preservation.  In particular, where the raw approach
-requires one first define substitution and reduction and then prove
-they preserve types, the inherent approach establishes substitution at
-the same time it defines substitution and reduction.
+The raw approach requires more lines of code than the inherent approach.  The
+separate definition of raw terms is not needed in the inherent approach; and one
+judgment in the raw approach needs to check that $\ex \not\equiv \why$, while
+the corresponding judgment in the inherent approach does not.  The difference
+becomes more pronounced when including the code for substitution, reductions,
+and proofs of progress and preservation.  In particular, where the raw approach
+requires one first define substitution and reduction and then prove they
+preserve types, the inherent approach establishes substitution at the same time
+it defines substitution and reduction.
 
-Stripping
-out examples and any proofs that appear in one but not the other (but
-could have appeared in both), the full development in PLFA for the raw approach takes 451
-lines (216 lines of definitions and 235 lines for the proofs) and the
-development for the inherent approach takes 275 lines (with
-definitions and proofs interleaved).  We have 451 / 235 = 1.64,
-close to the golden ratio.
+Stripping out examples and any proofs that appear in one but not the other (but
+could have appeared in both), the full development in PLFA for the raw approach
+takes 451 lines (216 lines of definitions and 235 lines for the proofs) and the
+development for the inherent approach takes 275 lines (with definitions and
+proofs interleaved).  We have 451 / 235 = 1.64, close to the golden ratio.
 
@@
 \citet{Altenkirch-and-Reus-1999} and described in
-\citet{Allais-et-al-2017}. In fact, I did manage to write a variant of
+\citet{Allais-et-al-2017}. In fact, we did manage to write a variant of
 the raw approach with simultaneous open substitution along the lines
@@
 two approaches combined.
+% TODO: is this still available somewhere in extra/?
 
@@
 notably by \citet{Chapman-2009} and \citet{Allais-et-al-2017}.
-I'm grateful to David Darais for bringing it to my attention.
+Philip is grateful to David Darais for bringing it to his attention.
 
@@
 
-I look forward to experience teaching from the new text,
+We look forward to experience teaching from the new text,
 and encourage others to use it too.  Please comment!

### 2019-07-14 papers/scp/PLFA.tex

@@
 and uses bidirectional type inference, while SF has terms with unique types and
-uses type checking.  SF also covers a simple imperative language with Hoare logic,
-and for lambda calculus covers subtyping, record types, mutable references, and normalisation
----none of which are treated by PLFA.  PLFA covers an inherently-typed de Bruijn
-representation, bidirectional type inference, bisimulation, and an
-untyped call-by-name language with full normalisation---none of which
-are treated by SF.
+uses type checking.  SF also covers a simple imperative language with Hoare
+logic, and for lambda calculus covers subtyping, record types, mutable
+references, and normalisation---none of which are treated by PLFA.  PLFA covers
+an inherently-typed de Bruijn representation, bidirectional type inference,
+bisimulation, and an untyped call-by-name language with full
+normalisation---none of which are treated by SF.
 
 SF has a third volume, written by Andrew Appel, on Verified Functional
-Algorithms. I'm not sufficiently familiar with that volume to have a view
-on whether it would be easy or hard to cover that material in Agda.
-And SF recently added a fourth volume on random testing of Coq specifications
-using QuickChick.  There is currently no tool equivalent to QuickChick
-available for Agda.
+Algorithms. I'm not sufficiently familiar with that volume to have a view on
+whether it would be easy or hard to cover that material in Agda. And SF recently
+added a fourth volume on random testing of Coq specifications using QuickChick.
+There is currently no tool equivalent to QuickChick available for Agda.
 
-There is more material that would be desirable to include in PLFA
-which was not due to limits of time.  In future years, PLFA may be
-extended to cover additional material, including mutable references,
-normalisation, System F, and pure type systems.
-We'd especially like to include pure type systems as they
-provide the readers with a formal model close to the dependent types
-used in the book.  Our attempts so far to formalise pure type systems
-have proved challenging, to say the least.
+There is more material that would be desirable to include in PLFA which was not
+due to limits of time.  In future years, PLFA may be extended to cover
+additional material, including mutable references, normalisation, System F, and
+pure type systems. We'd especially like to include pure type systems as they
+provide the readers with a formal model close to the dependent types used in the
+book.  Our attempts so far to formalise pure type systems have proved
+challenging, to say the least.
 
@@
 
-SF covers an imperative language with Hoare logic, culminating in
-code that takes an imperative programme suitably decorated with preconditions
-and postconditions and generates the necessary
-verification conditions. The conditions are then verified by a custom
-tactic, where any questions of arithmetic are resolved by the
-``omega'' tactic invoking a decision procedure.  The entire
-exercise would be easy to repeat in Agda, save for the last step: we
-suspect Agda's automation would not be up to verifying the generated
-conditions, requiring tedious proofs by hand.  However, we had already
-decided to omit Hoare logic in order to focus on lambda calculus.
+SF covers an imperative language with Hoare logic, culminating in code that
+takes an imperative programme suitably decorated with preconditions and
+postconditions and generates the necessary verification conditions. The
+conditions are then verified by a custom tactic, where any questions of
+arithmetic are resolved by the ``omega'' tactic invoking a decision procedure.
+The entire exercise would be easy to repeat in Agda, save for the last step, as
+Agda does not offer support for proof automation out of the box. It is 
+certainly possible to implement proof automation in Agda---see, e.g., the auto
+tactic by \citet{Kokke-2015}, and the collection of tactics in Ulf Norell's
+\texttt{agda-prelude}\footnote{\url{https://github.com/UlfNorell/agda-prelude}}.
+The standard library comes equipped with solvers for equations on monoids and
+rings\footnote{\url{https://agda.github.io/agda-stdlib/Algebra.Solver.Ring.html}},
+and a much improved solver for equalities on rings was recently contributed by
+\citet{Kidney-2019}.
+We suspect that, while Agda's automation would be up to verifying the generated
+conditions, some effort would be require to implement the required custom
+tactic, and a section would need to be added to the book to cover proof
+automation. For the time being, we have decided to omit Hoare logic in order to
+focus on lambda calculus.
 
@@
 
-A special thank you to my coauthor, Wen Kokke.  For inventing ideas on
-which PLFA is based, and for hand-holding, many thanks to Conor
-McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing me
-how much more compact it is to avoid raw terms, thanks to David
-Darais.  For inspiring my work by writing SF, thanks to Benjamin
-Pierce and his coauthors.  For comments on a draft of this paper, an
-extra thank you to James McKinna, Ulf Norell, Andreas Abel, and
-Benjamin Pierce. This research was supported by EPSRC Programme Grant
-EP/K034413/1.
+For inventing ideas on which PLFA is based, and for hand-holding, many thanks to
+Conor McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing me how
+much more compact it is to avoid raw terms, thanks to David Darais.  For
+inspiring my work by writing SF, thanks to Benjamin Pierce and his coauthors.
+For comments on a draft of this paper, an extra thank you to James McKinna, Ulf
+Norell, Andreas Abel, and Benjamin Pierce. This research was supported by EPSRC
+Programme Grant EP/K034413/1.
 

### 2019-07-14 papers/scp/PLFA.tex
new material on animation

@@
 
+% new
+Philip had been exposed to the work of the K team, as both consulted
+for IOHK, a cryptocurrency firm.  This put us keenly in mind of the
+need for animation; Philip sometime referred to this as ``K-envy'' or
+``Redex-envy''.
+
 % TODO: rewrite use of 'Philip' to reflect all authors?
@@
 an observation in the literature nor mentioned in an introductory
-text.  It does not appear in SF, nor in \citet{Harper-2016}.  A plea
+text.  It does not appear in SF, which introduces a specialised
+\texttt{normalise} tactic instead.  A plea
 to the Agda mailing list failed to turn up any prior mentions.
-The closest related observation I have seen in the published
+The closest related observation we have seen in the published
 literature is that evaluators can be extracted from proofs of
@@
 
-(Late addition: Our plea to the Agda list eventually bore fruit.  Oleg
-Kiselyov directed me to unpublished remarks on his web page where he
-uses the name \texttt{eval} for a proof of progress and notes ``the
-very proof of type soundness can be used to evaluate sample
-expressions'' \citep{Kiselyov-2009}.)
+% NEW
+Some researchers are clearly familiar with the connection between
+progress and preservation and animation.  In private correspondence,
+Bob Harper referred to it as the \emph{pas de deux}, a dance between
+progress, which takes well-typing to a step, and preservation, which
+takes a step back to well-typing.  Nonetheless, neither the technique
+nor the appealing terminology, appears in \citet{Harper-2016}.  The
+appeal to the Agda mailing list bore late fruit: Oleg Kiselyov
+directed me to unpublished remarks on his web page where he uses the
+name \texttt{eval} for a proof of progress and notes ``the very proof
+of type soundness can be used to evaluate sample expressions''
+\citep{Kiselyov-2009}.  Nonetheless, as of this writing, we still have
+not located a mention in the published literature.
+
+% NEW
+There are places in the literature where one might expect a remark on
+the relation between progress and preservation and animation---but no
+such remark appears.  In the PoplMark Challenge
+\citep{Ayedemir-et-al-2005}, Challenge~2A is to prove progress and
+preservation for System F$_{<:}$, while Challenge~3 is to prove
+animation for the same system. Nowhere do the authors indicate that in
+an intuitionistic logic these are essentially the same problem.
+\cite{Owens-et-al-2016}, when discussing extraction of animators for
+small-step semantics, mention Redex and K, but no other possibilities.
+We hope the stress in PLFA on the fact that in an intuitionistic
+setting progress and preservation imply animation will mean that the
+connection becomes more widely known.
 

### 2019-07-14 papers/scp/PLFA.tex
Halfway through adding material on experience teaching the course

@@
 
+\section{Experience}
+
+Philip now has five years of experience teaching from SF and one year
+teaching from PLFA.  He taught three courses from PLFA.
+\begin{itemize}
+\item
+  University of Edinburgh, September--December 2018 (with teaching
+  assistance from Wen and Chad Nester); twenty two-hour slots, one
+  hour of lecture followed by one hour of lab. Ten students completed
+  the course, which covered Parts~I and~II of PLFA.
+
+\item
+  Pontifícia Universidade Católica do Rio de Janeiro (PUC-Rio),
+  March--July 2019, hosted by Roberto Ieuramalischy; ten three-hour
+  slots, two hours of lecture followed by one hour of lab.  Ten
+  students completed the course, which covered Parts~I and~II of PLFA,
+  save students read chapter Lists on their own, and chapter
+  Bisimilarity was skipped.
+
+\item University of Padova, June 2018, hosted by Maria Emilia Maietti;
+  two three-hour slots, two hours of lecture followed by one hour of
+  lab.  About thirty students sat the course. Covered chapters
+  Naturals, Induction, and Relations.
+
+\end{itemize}
+In addition, David Darais at University of Vermont and John Leo at
+Google Seattle have taught from PLFA.
+
+Exercises in PLFA are classified in three ways.
+\begin{itemize}
+\item
+  Exercises labelled “(recommended)” are the ones students are
+  required to do in the classes taught at Edinburgh and PUC-Rio.
+
+\item
+  Exercises labelled “(stretch)” are there to provide an extra
+  challenge. Few students do all of these, but most attempt at least a
+  few.  Ten students completed the course.
+
+\item   
+  Exercises without a label are included for those who want extra
+  practice.  To Philip's surprise, students at PUC-Rio completed
+  quite a few of these.
+\end{itemize}
+Students are expected to complete coursework for all of the
+required questions in the text, optionally doing any stretch or
+unlabelled exercises. Coursework also includes a ``mock mock''
+exam, as described below.
+
+The mark for the course is based on coursework and a final-exam,
+weighted 1/4 for coursework and 3/4 for the final exam.  The weighting
+for coursework is designed to be high enough to encourage students to
+do it (they all do), but not so high as to encourage cheating.
+Students are encouraged to help each other with coursework.
+
+The final-exam is two hours online. Students have access to the Agda
+proof assistant to check their work.  At Edinburgh, students use
+computers with a special exam operating system that disables access
+to the internet.  Students are given access to the text of PLFA,
+the Agda standard libraries, and the Agda language reference manual,
+but no other materials.
+
+Students must answer question 1 on the exam, and one of questions 2 or
+3.  Question 1 gives predicates over a data structure, such lists or
+trees, to be formalised in Agda, and a theorem relating the
+predicates, to be formalised and proved in Agda.  Question 2 gives the
+students the intrinsic formulation of lambda calculus from chapter
+DeBruijn, which they must extend with a described language feature.
+Question 3 give the students the the bidirectional type inferencer
+from chapter Inference, which they must extend with a described
+language feature.
+
+Because the course is taught using a proof assistant, it is important
+that the students have access to a proof assistant during the
+exam. The whole point of a proof assistant is to avoid
+errors. Students are told in advance that they are expected to get
+perfect on the exam, and that they will have to study hard to achieve
+this level.  Given that the goal of formal methods is to avoid error,
+we think a pedagogical purpose is served by telling the students that
+they are expected to achieve perfection and making it possible for
+them to do so.
+
+they are expected to answer
+one question about formalising predicates on a data structure such
+as lists or tree
+
+The coursework is assigned 1/4 of the
+total mark of the course: enough to encourage the students to do
+it, but not enough to make it worth cheating. The remaining 3/4 of
+the tot
+
+Most of the marks are base
+
+There is also a ``mock mock'' exam
+assigned as cou
+
+The coursework was broken into four or five clusters, one due every
+few weeks. There is also a coursework consisting of a ``mock mock''
+exam, see below.
+
+recommended
+At Edinburgh and PUC-Rio, students completed coursework based on
+
 \section{Conclusion}

### 2019-07-15 papers/scp/PLFA.tex
finished segment on exams

@@
   University of Edinburgh, September--December 2018 (with teaching
-  assistance from Wen and Chad Nester); twenty two-hour slots, one
-  hour of lecture followed by one hour of lab. Ten students completed
-  the course, which covered Parts~I and~II of PLFA.
+  assistance from Wen and Chad Nester); twenty two-hour slots,
+  comprising one hour of lecture followed by one hour of lab. Ten
+  students completed the course, fourth-year undergraduates and
+  masters.  The course covered Parts~I and~II of PLFA.
 
@@
   March--July 2019, hosted by Roberto Ieuramalischy; ten three-hour
-  slots, two hours of lecture followed by one hour of lab.  Ten
-  students completed the course, which covered Parts~I and~II of PLFA,
-  save students read chapter Lists on their own, and chapter
-  Bisimilarity was skipped.
+  slots, comprising two hours of lecture followed by one hour of lab.
+  Ten students completed the course, mostly doctoral students.  The
+  course covered Parts~I and~II of PLFA, save students read chapter
+  Lists on their own, and chapter Bisimilarity was skipped.
 
-\item University of Padova, June 2018, hosted by Maria Emilia Maietti;
-  two three-hour slots, two hours of lecture followed by one hour of
-  lab.  About thirty students sat the course. Covered chapters
-  Naturals, Induction, and Relations.
+\item
+  University of Padova, June 2018, hosted by Maria Emilia Maietti; two
+  three-hour slots, comprising two hours of lecture followed by one
+  hour of lab.  Thirty undergraduate students sat the course, which
+  covered chapters Naturals, Induction, and Relations.
 
@@
   challenge. Few students do all of these, but most attempt at least a
-  few.  Ten students completed the course.
+  few.
 
@@
 Because the course is taught using a proof assistant, it is important
-that the students have access to a proof assistant during the
-exam. The whole point of a proof assistant is to avoid
-errors. Students are told in advance that they are expected to get
-perfect on the exam, and that they will have to study hard to achieve
-this level.  Given that the goal of formal methods is to avoid error,
-we think a pedagogical purpose is served by telling the students that
-they are expected to achieve perfection and making it possible for
-them to do so.
+that students have access to a proof assistant during the exam.
+Students are told in advance that they are expected to get perfect on
+the exam, and that they will have to study hard to achieve this level.
+Given that the goal of formal methods is to avoid error, we believe a
+pedagogical purpose is served by telling the students that they are
+expected to achieve perfection and making it possible for them to do
+so.  Students are given two opportunities to practice in the run up to
+the exam, a `mock' exam given in class under exam conditions (two
+hours online), and before that a `mock mock' exam as coursework (in
+their own time, encouraged to ask questions, tasked to do all three
+questions rather than two of three).
 
-they are expected to answer
-one question about formalising predicates on a data structure such
-as lists or tree
+Table~\ref{tab:exams} shows performance on the mock exam and the final
+exam for the courses run at Edinburgh and PUC-Rio.  While scores vary
+widely on the mock, all students achieve perfection on the
+exam.  (The one exception was a PUC-Rio student who did not attend
+classes or sit the mock.)  Similar results were achieved at Edinburgh
+over the previous five years, using SF as the course textbook and Coq
+as the proof assistant.  We consider these results a tribute to the
+students' ability to study and learn.
 
-The coursework is assigned 1/4 of the
-total mark of the course: enough to encourage the students to do
-it, but not enough to make it worth cheating. The remaining 3/4 of
-the tot
-
-Most of the marks are base
-
-There is also a ``mock mock'' exam
-assigned as cou
-
-The coursework was broken into four or five clusters, one due every
-few weeks. There is also a coursework consisting of a ``mock mock''
-exam, see below.
-
-recommended
-At Edinburgh and PUC-Rio, students completed coursework based on
+\begin{table}
+\begin{center}
+  \begin{tabular}{|cc|cc|}
+    \hline
+    \multicolumn{2}{|c|}{Edinburgh} &
+    \multicolumn{2}{c|}{PUC-Rio} \\
+    \hline
+    Mock & Exam & Mock & Exam \\
+    \hline
+     15  &  50  &  --  &  40  \\
+     29  &  50  &  ~6  &  50  \\
+     33  &  50  &  20  &  50  \\
+     35  &  50  &  28  &  50  \\
+     36  &  50  &  41  &  50  \\
+     48  &  50  &  49  &  50  \\
+     49  &  50  &  50  &  50  \\
+     50  &  50  &  50  &  50  \\
+     50  &  50  &  50  &  50  \\
+     50  &  50  &  50  &  50  \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Exam marks}
+\label{tab:exam}
+\end{table}
 

### 2019-07-15 papers/scp/PLFA.tex
changes to intro and conclusion

@@
 \begin{abstract}
-  One of the leading textbooks for formal methods is \emph{Software Foundations} (SF), written by Benjamin Pierce in collaboration with others, and based on Coq.
-After five years using SF in the classroom, I have come to the conclusion that Coq is not the best vehicle for this purpose, as too much of the course needs to focus on learning tactics for proof derivation, to the cost of learning programming language theory.
-Accordingly, I have written a new textbook, \emph{Programming Language Foundations in Agda} (PLFA).
-PLFA covers much of the same ground as SF, although it is not a slavish imitation.
 
-  What did I learn from writing PLFA?
-First, that it is possible.
-One might expect that without proof tactics that the proofs become too long, but in fact proofs in PLFA are about the same length as those in SF.
-Proofs in Coq require an interactive environment to be understood, while proofs in Agda can be read on the page.
-Second, that constructive proofs of preservation and progress give immediate rise to a prototype evaluator.
-This fact is obvious in retrospect but it is not exploited in SF (which instead provides a separate normalise tactic) nor can I find it in the literature.
-Third, that using raw terms with a separate typing relation is far less perspicuous than using inherently-typed terms.
-SF uses the former presentation, while PLFA presents both; the former uses about 1.6 as many lines of Agda code as the latter, roughly the golden ratio.
+  One of the leading textbooks for formal methods is \emph{Software
+    Foundations} (SF), written by Benjamin Pierce in collaboration
+  with others, and based on Coq. After five years using SF in the
+  classroom, we came to the conclusion that Coq is not the best
+  vehicle for this purpose, as too much of the course needs to focus
+  on learning tactics for proof derivation, to the cost of learning
+  programming language theory.  Accordingly, we have written a new
+  textbook, \emph{Programming Language Foundations in Agda}
+  (PLFA). PLFA covers much of the same ground as SF, although it is
+  not a slavish imitation.
 
-  The textbook is written as a literate Agda script, and can be found here:
+  What did we learn from writing PLFA? First, that it is possible. One
+  might expect that without proof tactics that the proofs become too
+  long, but in fact proofs in PLFA are about the same length as those
+  in SF. Proofs in Coq require an interactive environment to be
+  understood, while proofs in Agda can be read on the page.  Second,
+  that constructive proofs of preservation and progress give immediate
+  rise to a prototype evaluator. This fact is obvious in retrospect
+  but it is not exploited in SF (which instead provides a separate
+  normalise tactic) nor can I find it in the literature.  Third, that
+  using raw terms with a separate typing relation is far less
+  perspicuous than using inherently-typed terms. SF uses the former
+  presentation, while PLFA presents both; the former uses about 1.6 as
+  many lines of Agda code as the latter, roughly the golden ratio.
+
+  The textbook is written as a literate Agda script, and can be found
+  here:
   \begin{center}
@@
 
-Since 2013, I have taught a course on Types and Semantics for
+Since 2013, one of us (Philip) has taught a course on Types and Semantics for
 Programming Languages to fourth-year undergraduates and masters
@@
 course was based on \emph{Types and Programming Languages} by
-\citet{Pierce-2002}, but my version was taught from its successor,
+\citet{Pierce-2002}, but this version was taught from its successor,
 \emph{Software Foundations} (hence, SF) by \citet{Pierce-et-al-2010},
 which is based on the proof assistance Coq \citep{Huet-et-al-1997}.
-I am convinced by the claim of \citet{Pierce-2009}, made in his ICFP
+We are convinced by the claim of \citet{Pierce-2009}, made in his ICFP
 Keynote \emph{Lambda, The Ultimate TA}, that basing a course around a
@@
 
-However, after five years of experience, I have come to the conclusion
+However, after five years of experience, Phil came to the conclusion
 that Coq is not the best vehicle.  Too much of the course needs to
@@
 
-I found myself keen to recast the course in Agda \citep{Bove-et-al-2009}.
+We found ourselves keen to recast the course in Agda \citep{Bove-et-al-2009}.
 In Agda, there is
@@
 maintaining the same text but transposing the code from Coq to Agda.
-But it quickly became clear to me that after five years in the
-classroom I had my own ideas about how to present the material.  They
+But it quickly became clear that after five years in the
+classroom Philip had his own ideas about how to present the material.  They
 say you should never write a book unless you cannot \emph{not} write the
-book, and I soon found that this was a book I could not not write.
+book, and Philip soon found that this was a book he could not not
+write.
 
-I am fortunate that my student, Wen Kokke, was keen to help.  She
-guided me as a newbie to Agda and provided an infrastructure that is
-easy to use and produces pages that are a pleasure to view.
-The bulk of the book was written January--June 2018, while on
-sabbatical in Rio de Janeiro.
+Philip considered himself fortunate that his student, Wen, was
+keen to help.  She guided Philip as a newbie to Agda and provided an
+infrastructure for the book that we found easy to use and produces
+pages that are a pleasure to view.  The bulk of the first draft of the
+book was written January--June 2018, while Philip was on sabbatical in
+Rio de Janeiro.  After the first draft was published, Jeremy wrote
+two further parts.  One explains a new and simple approach
+to denotational semantics, developed as part of his research.  The
+other develops important precursors to the semantics, including
+confluence and big-step semantics.
 
-This paper is a personal reflection, summarising what I learned in the
+This paper is a personal reflection, summarising what was learned in the
 course of writing the textbook. Some of it reiterates advice that is
@@
 balloon in size. In fact, it turns out that for the results in PLFA
-and SF, the proofs are of roughly comparable size, and (in my opinion)
+and SF, the proofs are of roughly comparable size, and (in our opinion)
 the proofs in PLFA are more readable and have a pleasing visual
@@
 preservation combine trivially to produce a constructive evaluator for
-terms.  This idea is obvious once you have seen it, yet I cannot
-find it described in the literature.  For instance, SF separately
-implements a \texttt{normalise} tactic that has nothing to do with
-progress and preservation.
+terms.  This idea is obvious once you have seen it, yet we cannot
+find it described in the literature.
 
@@
 
-I will argue that Agda has advantages over Coq for
-pedagogic purposes.  My focus is purely on the case of a proof assistant
+Section~6 describes experience teaching from the textbook.  The point
+of proof is perfection, and it turns out that an online final
+examination with access to a proof assistant can lead to flawless
+student performance.
+
+Section~7 outlines our experience publishing the book as open source
+in Github.  We were surprised at how effective this was at eliciting
+community participation.  A large number of people have submitted pull
+requests to improve the book.
+
+We argue that Agda has advantages over Coq for
+pedagogic purposes.  Our focus is purely on the case of a proof assistant
 as an aid to \emph{learning} formal semantics using examples of
-\emph{modest} size.  I admit up front that
+\emph{modest} size.  We admit up front that
 there are many tasks for which Coq is better suited than Agda.
@@
 
-The book is broken into two parts. The first part, Logical Foundations,
+The book is broken into three parts. The first part, Logical Foundations,
 develops the needed formalisms.  The second part, Programming Language
 Foundations, introduces basic methods of operational semantics.
+The third part, Denotational Semantics, introduces a simple
+model of semantics, and shows it is compositional, sound, adequate, 
+and implies contextual equivalence.
 (SF is divided into books, the first two of which have the same names
-as the two parts of PLFA, and cover similar material.)
+as the first two parts of PLFA, and cover similar material.)
 
@@
 bisimulation, and an untyped call-by-name language with full
-normalisation---none of which are treated by SF.
+normalisation---none of which are treated by SF.  The new part on
+Denotational Semantics also covers material not treated by SF.
 
@@
 There is more material that would be desirable to include in PLFA which was not
-due to limits of time.  In future years, PLFA may be extended to cover
-additional material, including mutable references, normalisation, System F, and
+due to limits of time, including mutable references, logical relations, System F, and
 pure type systems. We'd especially like to include pure type systems as they
@@
 the term is well typed, and didn't even notice SF's description of the
-issue until I was done.
+issue until we were done.
 
@@
 
-\section{Experience}
+\section{Teaching experience}
 
 Philip now has five years of experience teaching from SF and one year
-teaching from PLFA.  He taught three courses from PLFA.
+teaching from PLFA.  To date, he has taught three courses from PLFA.
 \begin{itemize}
@@
 
-We look forward to experience teaching from the new text,
-and encourage others to use it too.  Please comment!
+The book is published as open-source on Github, under a
+Creative Commons CC-BY license.  As the book is an executable
+Agda script, it was considered essential that anyone could
+download and execute it.
+
+One sign of a successful publication is that it will attract a few
+letters from readers who have noticed typos or other problems.  An
+unexpected benefit of publishing on Github is that to date forty
+readers have sent pull requests.  Most of these fix typos, but a fair
+number make more substantial improvements.  Wen wrote a script that
+automatically adds the name of any contributor of an accepted pull
+request to the list of acknowledgements, sorted by the number of
+accepted requests.  Arguably, a different metric, such as total
+number of affected lines, might be more appropriate.
+
+There is much left to do!  We hope others may be inspired to join us
+and expand the book.
 
@@
 For inventing ideas on which PLFA is based, and for hand-holding, many thanks to
-Conor McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing me how
+Conor McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing how
 much more compact it is to avoid raw terms, thanks to David Darais.  For
-inspiring my work by writing SF, thanks to Benjamin Pierce and his coauthors.
+inspiring our work by writing SF, thanks to Benjamin Pierce and his coauthors.
 For comments on a draft of this paper, an extra thank you to James McKinna, Ulf

### 2019-07-15 papers/scp/PLFA.tex
fixed some wording

@@
 Rio de Janeiro.  After the first draft was published, Jeremy wrote
-two further parts.  One explains a new and simple approach
-to denotational semantics, developed as part of his research.  The
-other develops important precursors to the semantics, including
-confluence and big-step semantics.
+eight additional chapters, covering aspects of operational and
+denotational semantics.
 
@@
 The third part, Denotational Semantics, introduces a simple
-model of semantics, and shows it is compositional, sound, adequate, 
-and implies contextual equivalence.
+model of semantics and its properties.
 (SF is divided into books, the first two of which have the same names
 as the first two parts of PLFA, and cover similar material.)
+Part~I and Part~II up to Untyped were written by Phil,
+Part~II from Substitution and Part~III were written by Jeremy.
 

### 2019-07-15 papers/scp/PLFA.tex
fix typo

@@
   Pontifícia Universidade Católica do Rio de Janeiro (PUC-Rio),
-  March--July 2019, hosted by Roberto Ieuramalischy; ten three-hour
+  March--July 2019, hosted by Roberto Ierusalimschy; ten three-hour
   slots, comprising two hours of lecture followed by one hour of lab.

### 2019-07-15 papers/scp/PLFA.tex
fix typo

@@
 Section~7 outlines our experience publishing the book as open source
-in Github.  We were surprised at how effective this was at eliciting
+in GitHub.  We were surprised at how effective this was at eliciting
 community participation.  A large number of people have submitted pull
@@
 
-Table~\ref{tab:exams} shows performance on the mock exam and the final
-exam for the courses run at Edinburgh and PUC-Rio.  While scores vary
-widely on the mock, all students achieve perfection on the
+For the courses run at Edinburgh and PUC-Rio, the scores vary
+widely on the mock, but all students achieve perfection on the
 exam.  (The one exception was a PUC-Rio student who did not attend
@@
 
-\begin{table}
+\section{Software}
+
+The book is written using a combination of literate Agda and Markdown.  At the
+time of writing, the book is published using GitHub Pages and the Jekyll static
+site generator.  The book is open source---the source is currently also hosted
+on GitHub, under a Creative Commons CC-BY license.  The open-source aspect is
+important---as the book is written in literate Agda, it is essential that anyone
+can download and execute the source.
+
+We maintain a number of tools, which play various roles in rendering the book in
+all its ``glorious clickable HTML''. We render the literate Agda to highlighted
+HTML using Agda's HTML backend. In addition to highlighting, this inserts
+clickable links, linking each constructor and function to the site of its
+definition.  However, the links Agda inserts are local and don't match the
+structure of the book.  We maintain a script, \texttt{highlight.sh}, which fixes
+these links, rerouting links to the standard library to the online version, and
+correcting links to local modules.
+
+(Before the release of Agda 2.6, Agda did not support highlighting embedded
+literate code using HTML.  To address this, we wrote \texttt{agda2html}, a tool
+which rewrote the output of Agda's HTML highlighter to accomplish this. This
+tool had much more functionality, including the fixing of links as outlined
+above, the stripping of implicit arguments to achieve a Haskell-like look, and
+the support for new Markdown constructs for linking to Agda names. However,
+Agda 2.6 has incorporated almost all of this functionality.)
+
+The book is built, tested, and published after each commit, using Travis CI, a
+web service for continuous integration.  The means that the book is constantly
+changing.  To accommodate those who want a more stable version, e.g., for
+teaching, we maintain a stable version of the book at
 \begin{center}
-  \begin{tabular}{|cc|cc|}
-    \hline
-    \multicolumn{2}{|c|}{Edinburgh} &
-    \multicolumn{2}{c|}{PUC-Rio} \\
-    \hline
-    Mock & Exam & Mock & Exam \\
-    \hline
-     15  &  50  &  --  &  40  \\
-     29  &  50  &  ~6  &  50  \\
-     33  &  50  &  20  &  50  \\
-     35  &  50  &  28  &  50  \\
-     36  &  50  &  41  &  50  \\
-     48  &  50  &  49  &  50  \\
-     49  &  50  &  50  &  50  \\
-     50  &  50  &  50  &  50  \\
-     50  &  50  &  50  &  50  \\
-     50  &  50  &  50  &  50  \\
-    \hline
-  \end{tabular}
+  \url{http://plfa.inf.ed.ac.uk/stable}
 \end{center}
-\caption{Exam marks}
-\label{tab:exam}
-\end{table}
+The stable version of the book is only updated much less frequently, and updates
+are announced.
+
+We maintain a tool called, simply, \texttt{acknowledgements}, which uses
+the GitHub API to automatically extract a list of contributors to the book, and
+add them to the Acknowledgements page, each time the book is published.  We
+consider anyone who has sent a successful pull requests a contributor, and sort
+contributors in the acknowledgments by the number of accepted requests.
+Arguably, a different metric, such as total number of affected lines, might be
+more appropriate, though any solution may have its flaws.
 
@@
 
-The book is published as open-source on Github, under a
-Creative Commons CC-BY license.  As the book is an executable
-Agda script, it was considered essential that anyone could
-download and execute it.
-
-One sign of a successful publication is that it will attract a few
-letters from readers who have noticed typos or other problems.  An
-unexpected benefit of publishing on Github is that to date forty
-readers have sent pull requests.  Most of these fix typos, but a fair
-number make more substantial improvements.  Wen wrote a script that
-automatically adds the name of any contributor of an accepted pull
-request to the list of acknowledgements, sorted by the number of
-accepted requests.  Arguably, a different metric, such as total
-number of affected lines, might be more appropriate.
-
-There is much left to do!  We hope others may be inspired to join us
-and expand the book.
+One sign of a successful publication is that it will attract a few letters from
+readers who have noticed typos or other problems.  An unexpected benefit of
+publishing on GitHub is that to date forty-one readers have sent a total of
+\emph{two hundred seventy-five} pull requests.  Most of these fix typos, but a
+fair number make more substantial improvements.
+
+There is much left to do!  We hope others may be inspired to join us and expand
+the book.
 

### 2019-07-15 papers/scp/PLFA.tex
Updated SCP paper

@@
   Pontifícia Universidade Católica do Rio de Janeiro (PUC-Rio),
-  March--July 2019, hosted by Roberto Ierusalimschy; ten three-hour
+  March--July 2019, hosted by Roberto Ieuramalischy; ten three-hour
   slots, comprising two hours of lecture followed by one hour of lab.
@@
 
-For the courses run at Edinburgh and PUC-Rio, the scores vary
-widely on the mock, but all students achieve perfection on the
+Table~\ref{tab:exams} shows performance on the mock exam and the final
+exam for the courses run at Edinburgh and PUC-Rio.  While scores vary
+widely on the mock, all students achieve perfection on the
 exam.  (The one exception was a PUC-Rio student who did not attend
@@
 students' ability to study and learn.
+
+\begin{table}
+\begin{center}
+  \begin{tabular}{|cc|cc|}
+    \hline
+    \multicolumn{2}{|c|}{Edinburgh} &
+    \multicolumn{2}{c|}{PUC-Rio} \\
+    \hline
+    Mock & Exam & Mock & Exam \\
+    \hline
+     15  &  50  &  --  &  40  \\
+     29  &  50  &  ~6  &  50  \\
+     33  &  50  &  20  &  50  \\
+     35  &  50  &  28  &  50  \\
+     36  &  50  &  41  &  50  \\
+     48  &  50  &  49  &  50  \\
+     49  &  50  &  50  &  50  \\
+     50  &  50  &  50  &  50  \\
+     50  &  50  &  50  &  50  \\
+     50  &  50  &  50  &  50  \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Exam marks}
+\label{tab:exam}
+\end{table}
 

### 2019-07-16 papers/scp/PLFA.tex
minor edits

--- 2019-07-15 papers/scp/PLFA.tex
@@
 
-However, after five years of experience, Phil came to the conclusion
+However, after five years of experience, Philip came to the conclusion
 that Coq is not the best vehicle.  Too much of the course needs to
@@
 honours programme or the first year of a master or doctorate degree.
-It aims to teach the fundamentals of operational semantics of
-programming languages, with simply-typed lambda calculus as the
-central example.  The textbook is written as a literate script in Agda.
+It aims to teach the fundamentals of semantics of
+programming languages, with simply-typed and untyped lambda calculi as the
+central examples.  The textbook is written as a literate script in Agda.
 As with SF, the hope is that using
@@
 The third part, Denotational Semantics, introduces a simple
-model of semantics and its properties.
+model of the lambda calculus and its properties.
 (SF is divided into books, the first two of which have the same names
@@
 added a fourth volume on random testing of Coq specifications using QuickChick.
-There is currently no tool equivalent to QuickChick available for Agda.
+There is currently no tool equivalent to QuickChick for Agda.
 
@@
 sides line-up in the same column.  Our hope as authors is that students
-will read the formal proof first, and use it as a tabular guide
+read the formal proof first, and use it as a tabular guide
 to the informal explanation that follows.
@@
 the icon is expanded. As teachers, we were aware that students might skip the
-formal proof on a first reading, and we would have to hope the students would
+formal proof on a first reading, and we have to hope the students 
 return to it and step through it with an interactive tool in order to make it
@@
 particular proof forms the basis for a question of the mock exam and the past
-exams, so we expect most students will actually look at this one if not all the
+exams, so we expect most students will look at this one if not all the
 others.
@@
 In Coq, interaction consists of stepping through a proof script, at
-each point examining the current goal and the variables currently in
+each point examining the current goal and the variables in
 scope, and executing a new command in the script.  Tactics are a whole
@@
 recorded.  Students may recreate it by commenting out bits of code and
-introducing a hole in their place.   PLFA contains some prose descriptions
-of interactively building code, but mainly contains code that students
-can read.  They may also introduce holes to interact with the code, but
-we expect this will be rarer than with SF.
+introducing a hole in their place, but we expect this will be rarer
+than with SF.  PLFA contains some prose descriptions of interactively
+building code, but mainly contains code that students can read.
 
@@
 Students are told in advance that they are expected to get perfect on
-the exam, and that they will have to study hard to achieve this level.
+the exam, and that they will have to study hard to achieve it.
 Given that the goal of formal methods is to avoid error, we believe a
@@
 expected to achieve perfection and making it possible for them to do
-so.  Students are given two opportunities to practice in the run up to
+so.  Students are given two opportunities to practice before
 the exam, a `mock' exam given in class under exam conditions (two

### 2019-07-16 papers/scp/PLFA.tex
Minor edit to SCP paper

@@
 (Before the release of Agda 2.6, Agda did not support highlighting embedded
-literate code using HTML.  To address this, we wrote \texttt{agda2html}, a tool
-which rewrote the output of Agda's HTML highlighter to accomplish this. This
-tool had much more functionality, including the fixing of links as outlined
-above, the stripping of implicit arguments to achieve a Haskell-like look, and
-the support for new Markdown constructs for linking to Agda names. However,
-Agda 2.6 has incorporated almost all of this functionality.)
+literate code in HTML.  We maintained \texttt{agda2html}, a tool which rewrites
+the output of Agda's HTML highlighter to highlight embedded code. The tool had
+much more functionality, including the fixing of links as outlined above, the
+stripping of implicit arguments to achieve a Haskell-like look, and the support
+for new Markdown constructs for linking to Agda names. However, Agda 2.6 has
+incorporated almost all of this functionality, and \texttt{agda2html} is now
+deprecated.)
 

### 2019-07-18 papers/scp/PLFA.tex
Minor changes
small fixes to SCP paper

--- 2019-07-15 papers/scp/PLFA.tex
@@
+% WRITE TO JEREMY RE: DIFFERENCE IN STYLE OF CHAPTER TITLES
+% NEED TO CITE SBMF PAPER AND DESCRIBE DIFFERENCES
+% STRIP OUT EXAM RESULTS
+% CHANGE INHERENT --> INTRINSIC, CITE REYNOLDS
+
 \documentclass[preprint,authoryear]{elsarticle}
@@
 
+\hyphenation{coq-IDE}
+
 \begin{document}
@@
 
+This paper is the journal version of \citet{Wadler-2018}.
+It adds two new authors, summaries of Jeremy's new chapters,
+and sections on experience with teaching and software
+used to publish the book.  The original text often used
+first person, which here is replaced by reference to Philip.
+
 This paper is a personal reflection, summarising what was learned in the
@@
 as the first two parts of PLFA, and cover similar material.)
-Part~I and Part~II up to Untyped were written by Phil,
+Part~I and Part~II up to Untyped were written by Philip,
 Part~II from Substitution and Part~III were written by Jeremy.
@@
 \paragraph{More: More constructs of simply-typed lambda calculus}
-Introduces product, sum, unit, and empty types as well as lists and let bindings
-are explained.  Typing and reduction rules are given informally; a few
+Introduces product, sum, unit, and empty types; and explains lists and let bindings.
+Typing and reduction rules are given informally; a few
 are then give formally, and the rest are left as exercises for the reader.
@@
 SF has a third volume, written by Andrew Appel, on Verified Functional
-Algorithms. I'm not sufficiently familiar with that volume to have a view on
+Algorithms. We are not sufficiently familiar with that volume to have a view on
 whether it would be easy or hard to cover that material in Agda. And SF recently
@@
 Both texts are intended to be read online,
-and the figures were taken by grabbing bitmaps of the text as
+and the figures show screengrabs of the text as
 displayed in a browser.
@@
 to show the type of the hole and the types of the variables in scope;
-to check the code; to do a case analysis on a given variable; or to
-guess how to fill in the hole with constructors or variables in scope.
+to typecheck the code; to do a case analysis on a given variable; or to
+search for a way to fill in the hole with constructors or variables in scope.
 An Agda proof consists of typed code.  The interaction is \emph{not}
@@
 can read.  They may also introduce holes to interact with the code, but
-we expect this will be rarer than with SF.
+we expect this will be rare.
 
@@
 takes a step back to well-typing.  Nonetheless, neither the technique
-nor the appealing terminology, appears in \citet{Harper-2016}.  The
+nor the appealing terminology appears in \citet{Harper-2016}.  The
 appeal to the Agda mailing list bore late fruit: Oleg Kiselyov
@@
 such remark appears.  In the PoplMark Challenge
-\citep{Ayedemir-et-al-2005}, Challenge~2A is to prove progress and
+\citep{Aydemir-et-al-2005}, Challenge~2A is to prove progress and
 preservation for System F$_{<:}$, while Challenge~3 is to prove
@@
   University of Edinburgh, September--December 2018 (with teaching
-  assistance from Wen and Chad Nester); twenty two-hour slots,
+  assistance from Wen and Chad Nester); twenty 2-hour slots,
   comprising one hour of lecture followed by one hour of lab. Ten
@@
   Pontifícia Universidade Católica do Rio de Janeiro (PUC-Rio),
-  March--July 2019, hosted by Roberto Ieuramalischy; ten three-hour
+  March--July 2019, hosted by Roberto Ieuramalischy; ten 3-hour
   slots, comprising two hours of lecture followed by one hour of lab.
@@
   University of Padova, June 2018, hosted by Maria Emilia Maietti; two
-  three-hour slots, comprising two hours of lecture followed by one
+  3-hour slots, comprising two hours of lecture followed by one
   hour of lab.  Thirty undergraduate students sat the course, which
@@
 
-Students must answer question 1 on the exam, and one of questions 2 or
-3.  Question 1 gives predicates over a data structure, such lists or
+Students must answer question~1 on the exam, and one of questions~2
+or~3.  Question~1 gives predicates over a data structure, such lists or
 trees, to be formalised in Agda, and a theorem relating the
-predicates, to be formalised and proved in Agda.  Question 2 gives the
-students the intrinsic formulation of lambda calculus from chapter
+predicates, to be formalised and proved in Agda.  Question~2 gives the
+students the inherent formulation of lambda calculus from chapter
 DeBruijn, which they must extend with a described language feature.
-Question 3 give the students the the bidirectional type inferencer
+Question~3 give the students the the bidirectional type inferencer
 from chapter Inference, which they must extend with a described
@@
 so.  Students are given two opportunities to practice in the run up to
-the exam, a `mock' exam given in class under exam conditions (two
-hours online), and before that a `mock mock' exam as coursework (in
+the exam, a ``mock'' exam given in class under exam conditions (two
+hours online), and before that a ``mock mock'' exam as coursework (in
 their own time, encouraged to ask questions, tasked to do all three
@@
 
+% Table~\ref{tab:exams} shows performance on the mock exam and the final
+% exam for the courses run at Edinburgh and PUC-Rio.  While scores vary
+% widely on the mock, all students achieve perfection on the
 For the courses run at Edinburgh and PUC-Rio, the scores vary
-widely on the mock, but all students achieve perfection on the
+widely on the mock: minimum <20%, maximum 100%, mean
+77.8, standard deviation 27.6. But all students achieve perfection on the
 exam.  (The one exception was a PUC-Rio student who did not attend
@@
 students' ability to study and learn.
+
+% \begin{table}
+% \begin{center}
+%   \begin{tabular}{|cc|cc|}
+%     \hline
+%     \multicolumn{2}{|c|}{Edinburgh} &
+%     \multicolumn{2}{c|}{PUC-Rio} \\
+%     \hline
+%     Mock & Exam & Mock & Exam \\
+%     \hline
+%      15  &  50  &  --  &  40  \\
+%      29  &  50  &  ~6  &  50  \\
+%      33  &  50  &  20  &  50  \\
+%      35  &  50  &  28  &  50  \\
+%      36  &  50  &  41  &  50  \\
+%      48  &  50  &  49  &  50  \\
+%      49  &  50  &  50  &  50  \\
+%      50  &  50  &  50  &  50  \\
+%      50  &  50  &  50  &  50  \\
+%      50  &  50  &  50  &  50  \\
+%     \hline
+%   \end{tabular}
+% \end{center}
+% \caption{Exam marks}
+% \label{tab:exams}
+% \end{table}
 
@@
 The book is built, tested, and published after each commit, using Travis CI, a
-web service for continuous integration.  The means that the book is constantly
+web service for continuous integration.  This means that the book is constantly
 changing.  To accommodate those who want a more stable version, e.g., for
@@
 \end{center}
-The stable version of the book is only updated much less frequently, and updates
+The stable version of the book is updated much less frequently, and updates
 are announced.
@@
 add them to the Acknowledgements page, each time the book is published.  We
-consider anyone who has sent a successful pull requests a contributor, and sort
+consider anyone who has sent a successful pull request a contributor, and sort
 contributors in the acknowledgments by the number of accepted requests.
 Arguably, a different metric, such as total number of affected lines, might be
-more appropriate, though any solution may have its flaws.
+more appropriate, though any solution will have its flaws.
 
@@
 
-There is much left to do!  We hope others may be inspired to join us and expand
-the book.
+There is much left to do!  We hope others may be inspired to join us
+to expand and improve the book.
 

### 2019-07-18 papers/scp/PLFA.tex

@@
 of interactively building code, but mainly contains code that students
-can read.  They may also introduce holes to interact with the code, but
-we expect this will be rare.
+can read.  They may introduce holes to interact with the code, but
+we expect that will be rare.
 

### 2019-07-21 papers/scp/PLFA.tex
switch to extrinsic/intrinsic terminology

@@
-% WRITE TO JEREMY RE: DIFFERENCE IN STYLE OF CHAPTER TITLES
-% NEED TO CITE SBMF PAPER AND DESCRIBE DIFFERENCES
-% STRIP OUT EXAM RESULTS
-% CHANGE INHERENT --> INTRINSIC, CITE REYNOLDS
-
 \documentclass[preprint,authoryear]{elsarticle}
@@
   normalise tactic) nor can I find it in the literature.  Third, that
-  using raw terms with a separate typing relation is far less
-  perspicuous than using inherently-typed terms. SF uses the former
+  using extrinsicly-typed terms is far less
+  perspicuous than using intrinsicly-typed terms. SF uses the former
   presentation, while PLFA presents both; the former uses about 1.6 as
@@
 
-Section~5 claims that raw terms should be avoided in favour of
-inherently-typed terms.  PLFA develops lambda calculus with both raw
-and inherently-typed terms, permitting a comparison.  It turns out the
+Section~5 claims that extrinsicly-typed terms should be avoided in favour of
+intrisicly-typed terms.  PLFA develops lambda calculus with both,
+permitting a comparison.  It turns out the
 former is less powerful---it supports substitution only for closed
@@
 Introduces lambda calculus, using a representation with named
-variables and a separate typing relation. The language used is PCF
+variables and extrinsicly typed. The language used is PCF
 \citep{Plotkin-1977}, with variables, lambda abstraction, application,
@@
 
-\paragraph{DeBruijn: Inherently typed de Bruijn representation}
-Introduces de Bruijn indices and the inherently-typed representation.
+\paragraph{DeBruijn: Intrinsicly-typed de Bruijn representation}
+Introduces de Bruijn indices and the intrinsicly-typed representation.
 Emphasis is put on the structural similarity between a term and its
@@
 are then give formally, and the rest are left as exercises for the reader.
-The inherently typed representation is used.
+The intrinsicly-typed representation is used.
 
@@
 Introduces bidirectional type inference, and applies it to convert from
-a representation with named variables and a separate typing relation
-to a representation de Bruijn indices with inherent types. Bidirectional
+a representation with named variables and extrinsicly typed
+to a representation with de Bruijn indices and intrinsicly typed. Bidirectional
 type inference is shown to be both sound and complete.
@@
 \paragraph{Untyped: Untyped calculus with full normalisation}
-As a variation on earlier themes, discusses an untyped (but inherently
+As a variation on earlier themes, discusses an untyped (but intrinsicly
 scoped) lambda calculus.  Reduction is call-by-name over open terms,
@@
 references, and normalisation---none of which are treated by PLFA.  PLFA covers
-an inherently-typed de Bruijn representation, bidirectional type inference,
+an intrinsicly-typed de Bruijn representation, bidirectional type inference,
 bisimulation, and an untyped call-by-name language with full
@@
 There is more material that would be desirable to include in PLFA which was not
-due to limits of time, including mutable references, logical relations, System F, and
+due to limits of time, including mutable references, logical relations, System~F, and
 pure type systems. We'd especially like to include pure type systems as they
@@
 prototype.  The input is a term together with evidence the term is
-well-typed.  (In the inherently-typed case, these are the same thing.)
+well-typed.  (In the intrinsicly-typed case, these are the same thing.)
 Progress determines whether we are done, or should take another step;
 preservation provides evidence that the new term is well-typed, so we
-may iterate. In a language with guaranteed termination, we cannot
+may iterate. In a language with guaranteed termination such as Agda, we cannot
 iterate forever, but there are a number of well-known techniques to
@@
 
-\section{Inherent typing is golden}
+\section{Intrinsic typing is golden}
 
@@
   \includegraphics[width=\textwidth]{figures/raw.png}
-  \caption{Raw approach in PLFA}
-  \label{fig:raw}
+  \caption{Extrinsic approach in PLFA}
+  \label{fig:extrinsic}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{figures/inherent.png}
-  \caption{Inherent approach in PLFA}
-  \label{fig:inherent}
+  \caption{Intrinsic approach in PLFA}
+  \label{fig:intrinsic}
 \end{figure}
@@
 The second part of PLFA first discusses two different approaches to
-modeling simply-typed lambda calculus.  It first presents raw
-terms with named variables and a separate typing relation and
-then shifts to inherently-typed terms with de Bruijn indices.
+modeling simply-typed lambda calculus.  It first presents
+terms with named variables and and extrinsic typing relation and
+then shifts to terms with de Bruijn indices that are intrinsicly
+typed.  The names \emph{extrinsic} and \emph{intrinsic} for these
+two approaches are taken from \citet{Reynolds-2003}.
 Before writing the text, Philip had thought the two approaches
 complementary, with no clear winner.  Now he is convinced that the
-inherently-typed approach is superior.
-
-Figure~\ref{fig:raw} presents the raw approach.
+intrinsicly-typed approach is superior.
+
+Figure~\ref{fig:extrinsic} presents the extrinsic approach.
 It first defines $\texttt{Id}$, $\texttt{Term}$,
@@
 
-Figure~\ref{fig:inherent} presents the inherent approach.
+Figure~\ref{fig:intrinsic} presents the intrinsic approach.
 It first defines $\texttt{Type}$ and $\texttt{Context}$, the abstract syntax
@@
 we now represent a variable or a term by the proof that it is well typed.
-In particular, the proof that a variable is well typed in the raw approach
-corresponds to a de Bruijn index in the inherent approach.
-
-The raw approach requires more lines of code than the inherent approach.  The
-separate definition of raw terms is not needed in the inherent approach; and one
-judgment in the raw approach needs to check that $\ex \not\equiv \why$, while
-the corresponding judgment in the inherent approach does not.  The difference
+In particular, the proof that a variable is well typed in the extrinsic approach
+corresponds to a de Bruijn index in the intrinsic approach.
+
+The extrinsic approach requires more lines of code than the intrinsic approach.  The
+separate definition of raw terms is not needed in the intrinsic approach; and one
+judgment in the extrinsic approach needs to check that $\ex \not\equiv \why$, while
+the corresponding judgment in the intrinsic approach does not.  The difference
 becomes more pronounced when including the code for substitution, reductions,
-and proofs of progress and preservation.  In particular, where the raw approach
+and proofs of progress and preservation.  In particular, where the extrinsic approach
 requires one first define substitution and reduction and then prove they
-preserve types, the inherent approach establishes substitution at the same time
+preserve types, the intrinsic approach establishes substitution at the same time
 it defines substitution and reduction.
@@
 Stripping out examples and any proofs that appear in one but not the other (but
-could have appeared in both), the full development in PLFA for the raw approach
+could have appeared in both), the full development in PLFA for the extrinsic approach
 takes 451 lines (216 lines of definitions and 235 lines for the proofs) and the
-development for the inherent approach takes 275 lines (with definitions and
+development for the intrinsic approach takes 275 lines (with definitions and
 proofs interleaved).  We have 451 / 235 = 1.64, close to the golden ratio.
 
-The inherent approach also has more expressive power.  The raw
+The intrinsic approach also has more expressive power.  The extrinsic
 approach is restricted to substitution of one variable by a closed
-term, while the inherent approach supports simultaneous substitution
+term, while the intrinsic approach supports simultaneous substitution
 of all variables by open terms, using a pleasing formulation due to
@@
 \citet{Allais-et-al-2017}. In fact, we did manage to write a variant of
-the raw approach with simultaneous open substitution along the lines
+the extrinsic approach with simultaneous open substitution along the lines
 of McBride, but the result was too complex for use in an introductory
@@
 
-The text develops both approaches because the raw approach is more
-familiar, and because placing the inherent approach first would
+The text develops both approaches because the extrinsic approach is more
+familiar, and because placing the intrinsic approach first would
 lead to a steep learning curve.  By presenting the more long-winded
 but less powerful approach first, students can see for themselves the
-advantages of de Bruijn indices and inherent types.
+advantages of de Bruijn indices and intrinsic types.
 
 There are actually four possible designs, as the choice of named
-variables vs de Bruijn indices, and the choice of raw vs
-inherently-typed terms may be made independently.  There are synergies
+variables vs de Bruijn indices, and the choice of extrinsic vs
+intrinsic typing may be made independently.  There are synergies
 between the two.  Manipulation of de Bruijn indices can be notoriously
-error-prone without inherent-typing to give assurance of correctness.
-In inherent typing with named variables, simultaneous substitution by
+error-prone without intrinsic types to give assurance of correctness.
+For instrinsic typing with named variables, simultaneous substitution by
 open terms remains difficult.
 
-The benefits of the inherent approach 
+The benefits of the intrinsic approach 
 are well known to some. The technique was introduced by
@@
 predicates, to be formalised and proved in Agda.  Question~2 gives the
-students the inherent formulation of lambda calculus from chapter
+students the intrinsic formulation of lambda calculus from chapter
 DeBruijn, which they must extend with a described language feature.
@@
 For the courses run at Edinburgh and PUC-Rio, the scores vary
-widely on the mock: minimum <20%, maximum 100%, mean
+widely on the mock: minimum <20\%, maximum 100\%, mean
 77.8, standard deviation 27.6. But all students achieve perfection on the
@@
 Conor McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing how
-much more compact it is to avoid raw terms, thanks to David Darais.  For
+much more compact it is to use intrinsicly-typed terms, thanks to David Darais.  For
 inspiring our work by writing SF, thanks to Benjamin Pierce and his coauthors.

### 2019-07-21 papers/scp/PLFA.tex
fix spelling

@@
   normalise tactic) nor can I find it in the literature.  Third, that
-  using extrinsicly-typed terms is far less
-  perspicuous than using intrinsicly-typed terms. SF uses the former
+  using extrinsically-typed terms is far less
+  perspicuous than using intrinsically-typed terms. SF uses the former
   presentation, while PLFA presents both; the former uses about 1.6 as
@@
 
-Section~5 claims that extrinsicly-typed terms should be avoided in favour of
+Section~5 claims that extrinsically-typed terms should be avoided in favour of
 intrisicly-typed terms.  PLFA develops lambda calculus with both,
@@
 Introduces lambda calculus, using a representation with named
-variables and extrinsicly typed. The language used is PCF
+variables and extrinsically typed. The language used is PCF
 \citep{Plotkin-1977}, with variables, lambda abstraction, application,
@@
 
-\paragraph{DeBruijn: Intrinsicly-typed de Bruijn representation}
-Introduces de Bruijn indices and the intrinsicly-typed representation.
+\paragraph{DeBruijn: Intrinsically-typed de Bruijn representation}
+Introduces de Bruijn indices and the intrinsically-typed representation.
 Emphasis is put on the structural similarity between a term and its
@@
 are then give formally, and the rest are left as exercises for the reader.
-The intrinsicly-typed representation is used.
+The intrinsically-typed representation is used.
 
@@
 Introduces bidirectional type inference, and applies it to convert from
-a representation with named variables and extrinsicly typed
-to a representation with de Bruijn indices and intrinsicly typed. Bidirectional
+a representation with named variables and extrinsically typed
+to a representation with de Bruijn indices and intrinsically typed. Bidirectional
 type inference is shown to be both sound and complete.
@@
 \paragraph{Untyped: Untyped calculus with full normalisation}
-As a variation on earlier themes, discusses an untyped (but intrinsicly
+As a variation on earlier themes, discusses an untyped (but intrinsically
 scoped) lambda calculus.  Reduction is call-by-name over open terms,
@@
 references, and normalisation---none of which are treated by PLFA.  PLFA covers
-an intrinsicly-typed de Bruijn representation, bidirectional type inference,
+an intrinsically-typed de Bruijn representation, bidirectional type inference,
 bisimulation, and an untyped call-by-name language with full
@@
 prototype.  The input is a term together with evidence the term is
-well-typed.  (In the intrinsicly-typed case, these are the same thing.)
+well-typed.  (In the intrinsically-typed case, these are the same thing.)
 Progress determines whether we are done, or should take another step;
@@
 terms with named variables and and extrinsic typing relation and
-then shifts to terms with de Bruijn indices that are intrinsicly
+then shifts to terms with de Bruijn indices that are intrinsically
 typed.  The names \emph{extrinsic} and \emph{intrinsic} for these
@@
 complementary, with no clear winner.  Now he is convinced that the
-intrinsicly-typed approach is superior.
+intrinsically-typed approach is superior.
 
@@
 Conor McBride, James McKinna, Ulf Norell, and Andreas Abel.  For showing how
-much more compact it is to use intrinsicly-typed terms, thanks to David Darais.  For
+much more compact it is to use intrinsically-typed terms, thanks to David Darais.  For
 inspiring our work by writing SF, thanks to Benjamin Pierce and his coauthors.

### 2019-07-22 papers/scp/PLFA.tex

@@
   but it is not exploited in SF (which instead provides a separate
-  normalise tactic) nor can I find it in the literature.  Third, that
+  normalise tactic) nor can we find it in the literature.  Third, that
   using extrinsically-typed terms is far less
@@
 progress, which takes well-typing to a step, and preservation, which
-takes a step back to well-typing.  Nonetheless, neither the technique
+takes a step back to well-typing---but neither the technique
 nor the appealing terminology appears in \citet{Harper-2016}.  The
 appeal to the Agda mailing list bore late fruit: Oleg Kiselyov
-directed me to unpublished remarks on his web page where he uses the
+directed us to unpublished remarks on his web page where he uses the
 name \texttt{eval} for a proof of progress and notes ``the very proof
@@
 \citep{Aydemir-et-al-2005}, Challenge~2A is to prove progress and
-preservation for System F$_{<:}$, while Challenge~3 is to prove
+preservation for System~F$_{<:}$, while Challenge~3 is to prove
 animation for the same system. Nowhere do the authors indicate that in
@@
   students completed the course, fourth-year undergraduates and
-  masters.  The course covered Parts~I and~II of PLFA.
+  masters.  The course covered Parts~I and~II of PLFA, up through
+  chapter Untyped.
 
@@
   Ten students completed the course, mostly doctoral students.  The
-  course covered Parts~I and~II of PLFA, save students read chapter
+  course covered Parts~I and~II of PLFA, up through chapter Untyped,
+  save students read chapter
   Lists on their own, and chapter Bisimilarity was skipped.

### 2020-03-04 papers/scp/response-to-reviewers.txt
revision of scp paper

[We thank the reviewers for their comments. Responses interleaved in
brackets. The revision includes a few other improvements, in
particular consistently using standard terms "extrinsic"/"intrinsic"
in place of "raw"/"inherent".]

Ms. Ref. No.: SCICO-D-19-00172 Title: Programming Language Foundations
in Agda Science of Computer Programming

Dear Professor Philip Wadler,

Thank you very much for having submitted to the Science of Computer
Programming Journal.

We have now received reviewers' reports and a recommendation by the
handling editor; based upon the recommendation and the reviews, we
would like to ask you prepare a minor revision taking the comments of
the reviewers into account.

The due date for revision is Feb 14, 2020.  If this deadline is out of
sync with other due dates communicated to you outside EES, then please
contact the handling editor to settle this.

If you decide to revise the work, please submit a list of changes or a
rebuttal against each point which is being raised when you submit the
revised manuscript.

To submit a revision, please go to https://ees.elsevier.com/scico/ and
login as an Author.  Your username is: wadler@inf.ed.ac.uk If you need
to retrieve password details, please go to:
http://ees.elsevier.com/scico/automail_query.asp

On your Main Menu page is a folder entitled "Submissions Needing
Revision". You will find your submission record there.

Once again, thank you very much for having given us the opportunity to
consider your manuscript.

Mohammad Reza Mousavi, Ph.D.  Editor-in-Chief Science of Computer
Programming

Reviewers' comments:

The reviewers have been unanimous in accepting the article for
publication. There are a few remarks to be considered by the authors
for the final version, as pointed out in the reviews.

Reviewer #1: This article reports on experience teaching basic results
in lambda calculus using a proof assistant based on type theory.  The
first finding is that the proof style supported by Agda,
i.e. dependently typed programming, need not be more verbose than the
tactic-based style supported by Coq.  The second finding is a
functional pearl: constructive proof of progress and preservation
directly yield an interpreter.  The third finding is that
formalization of typed lambda calculus based on inherently typed terms
is significantly more succinct than formalization in terms of a
separate typing judgment.  Finally, the authors report on flawless
student performance on an exam using the proof assistant.

The last point may not seem surprising to readers with experience
teaching this material as presented in the book Software Foundations
(SF) by Pierce et al, specifically the volume title Programming
Language Foundations.  (Or for that matter, teaching related topics
using a proof assistant, e.g., based on the book Concrete Semantics by
Nipkow and Klein.)  The authors have written a book, PLFA, drawing on
corresponding chapters of the book by Pierce et al which develops a
more extensive range of programming language theory using Coq.
Section 2 of the article describes PLFA with reference to SF, and
Section 7 describes experience with open source publishing.

At first glance I was concerned that this article might be little more
than a promotion for the book.  But in fact the article offers
insightful comparisons between tactical proving and dependently typed
programming style, based on extensive classroom teaching experience.

The book SF introduces quite a few proof tactics and it was not at all
obvious that the same material could be covered perspicuously and
succinctly in Agda.  The examples and discussion in Section 3 make a
convincing case for the pedagogical effectiveness of Agda for the
toptics covered.  The article acknowledges that this does not
contradict the effectiveness of tactic-based proof assistants for
larger scale developments.

As for the second finding, the authors acknowledge that the idea seems
obvious and can to some extent be discerned in prior publications and
online documents.  But it does not seem to have been explicitly
presented in the research literature or textbooks, and it clearly
should be.

The third finding addresses alternatives dubbed extrinsic/intrinsic
typing by Reynolds, which to my knowledge have been viewed as each
having advantages and disadvantages -- thus tradeoffs to be considered
afresh in each situation.  The authors make a convincing case for
intrinsic typing by comparing well informed formalizations in both
styles and observing that intrinsic is in some sense more expressive
and also significantly more succinct.

The article is very well written and easy to read.  While the
experience report style is somewhat unusual for this particular
journal, there is substantial scientific content and the writing style
is effective in conveying that content.  Indeed, it conveys the
authors' enthusiasm for Agda and my review was delayed because I was
keen to get experience with PLFA.  A previous version of the paper
appeared in a conference; this article adds substantial material on
additional chapters of the book and on teaching experiences.

Judgment

This is an excellent article worthy of publication in SCP.  Below I
mention minor points that could be addressed, but significant revision
is not necessary as the article is already well polished.

Further thoughts

Apropos the volume Verified Functional Algorithms of SF, it makes
heavy use of tactical automation to handle large case analyses
(hundreds of cases, for red-black trees).  It would be quite
interesting to see whether at least the simpler algorithms can be
handled well in Agda.

[NOT DONE. Interesting, but we don't have time to do this work.]

page 8 lists logical relations as one of the topics not currently
included in PLFA, but on page 8 under Adequacy a logical relation is
used.

[Clarified by changing "logical relations" to "logical relations
for parametricity"]

page 11 mentions not finding a proof of determinacy of reduction.  The
chapter on Normalization of SF, by Tolmach (Norm.v) does so.

[Thanks! Text updated.]

page 12 near the bottom of the page it explains some notations that
occur on previous pages

[Not sure what change is requested here.]

Reviewer #2: Title: Programming Language Foundations in Agda.

Review:

This paper has an interesting structure: it focus on the overall
structure of the book "Programming Language Foundations in Agda", the
choices made and the advantages and difficulties of using Agda to
construct this material (with respect to other tools, in particular
Coq). The technical details are mostly about choices and style of
formalization, and initial results of the application of the material
in a few courses. Let me start by saying that the article is very
pleasant to read: the style is rather informal, but very clear and
provides the reader a good idea of the mindset of the authors that
lead to the choices they made, and also the influences they had along
the way.

A little context regarding my backgroung, to contextualize my
comments: I have been using Coq for roughly three years. During this
time, I have made several exercises of formalization, including the
untyped lambda-calculus (with explicit variable names and with
deBruijn indices), including the Church-Rosser theorem (based on
parallel reductions) and normalisation of left-most reductions. My
experience with Agda is very brief.

In general, I liked the article very much, and I would agree with the
majority of the claims made. The appointed topics below are mostly
comments, suggestions and some comments on a few claims that I found
slightly exaggerated.

Section 2 ---------

There is not much to say about this section, since it is an overview
of the topics that are introduced and proved in each section of the
book. In general, I liked very much the choice of topics and their
sequence.

Page 4, paragraph "Part I, Logical Foundations": suggestion: (briefly
explain monus: the operation is known, but the name 'monus' may not be
as common)

[Not done. It is a standard term, easily looked up. Explaining
it would expand the text more than warranted.]

Page 7, "Discussion": it is mentioned that SF starts with STLC, while
PLFA starts with PCF. It was not discussed the reason to skip a more
restricte, normalizing language (STLC) rather than a more powerful
language (PCF). Could you comment on the motivations for this choice?
It would be very nice to be able to speak about termination proofs and
logical relations at some place (as mentioned in future work).

[Added the requested comment.]

Page 8, "Discussion": It seems that the author claim's that the
substitution in PCF is weaker because it is only applied to closed
terms. However, this is a choice, since it is possible to define
substitution for open terms (only it is usually not worth the effort,
since it is more complex and involves alpha-renaming).

[The choice with substitution of open terms is described later
in the paper, pointing out that the code was too long to include
in the textbook.]

Page 8, "We'd" => "We would" (suggestion: the contraction felt way too
informal...)

[Fixed.]

Section 3 ---------

Regarding the comparison between Agda and Coq, some remarks:

I believe it is clear that Agda's syntax is nicer than Coq's syntax,
in particular very familiar to Haskell programmers. I found it very
interesting the usage of "--" coments to replicate the shape of rules,
mixfix notation and better Unicode support. However, I although I
believe that it was not clear to me that there is a huge advantage of
Agda over Coq in the following two aspects:

-> notation: Coq's notation system, although not as flexible as Agda,
   is quite similar in terms of effectiveness: duplication of names
   seems like a minor inconvenience, and would not affect the
   formalization effort in such a negative way, as claimed by the
   authors.

-> style of proof: it seems to me that proving theorems in Agda feels
   like programming more than proving in Coq (due to the distinction
   between the languages). This does not seem to me as good or bad per
   se: it depends on your objectives.

Although it is possible to construct the proof term directly in Coq,
it is not encouraged: the tactic language, although intimidating for
beginners, feels very comprehensive and powerful. Although I agree
that Coq requires an interactive process to be understood, I was not
convinced (by the arguments given in the paper and looking at some
more complex proofs in PLFA) that they are more readable than Coq's:
it seems more a question of preference of style, and desire to use the
same mechanics for programming and proving. Moreover, Coq seems to
"hide" a little better than Adga questions regarding the definition of
equality (you coud hide this at the beginning, and introduce it
later). I agree with the authors that Agda exposes more directly
Propositions as Types and avoid some duplications of Coq.

[The reviewer doesn't ask for specific changes.]

Section 4 ---------

Nice idea. (no problems here)

Section 5 ---------

Page 16, line 45 : "and and" => "and"

[Fixed.]

Page 19, line 34 : it seems that the comparison between intrinsically
 and extrinsically-typed lambda terms is not completely fair, in the
 sense that there are two main modifications involved: => explicit
 names vs deBruijn indices => intrisicly typed vs extrinsicly typed
 terms The authors comment that there are actually four possibilities
 and comment on some difficulties with the other options. Did you
 consider introducing and contrasting each dimension at the time,
 rather than making two choices at the same time (in the book)?

[Rephrased to clarify that we think only the two options
we presented work well. The following sentences explain why
the other options are problematic.]

Section 6 ---------

Page 21, line 21: "Similar results were achieved at Edinburgh over the
previous five years".

The presented data seems to suggest that the distinction between the
two systems seems more a question of style, evidencing that it is
possible to substitute Coq for Agda in this scenario. However, I does
not point towards a clear advantage in either way. Maybe an
interesting experiment would be to expose some introduction to both
systems, and try to measure the effort of doing similar tasks in both
platforms.

[The paper includes a suggestion from Nate Foster on similar lines.]

------------------------

Reviewer #3: The paper proposes an alternative to the textbook
Software Foundations (SF) by Pierce et. al. SF uses Coq, while the
authors propose the usage of Agda. The authors claim that, with Coq,
too much effort is spent on learning tactics for proof derivation
instead of learning programming language theory. In contrast, Agda
makes use of no tactics, but its proofs are not much longer than those
in Coq. As a result, the new textbook Programming Language Foundations
in Agda (PLFA) has been written and applied to 3 courses so far
(Edinburgh, Rio de Janeiro, and Padova). In addition to this
contribution to education, the authors mention two other
contributions: the insight that a constructive proof of preservation
and progress automatically produces a prototype evaluator of the
language; and that an intrinsic type relation requires fewer lines of
code and is more expressive than an extrinsic type relation.

The paper reports the development of a solid contribution to the
education of programming language theory and an alternative to the
Software Foundations textbook. On a more scientific note, the usage of
a constructive proof of preservation and progress to produce a
prototype evaluator is a nice insight not clearly described in the
literature.

As suggestions for improvement, I would like to see few comments on
design decisions taken on PLFA: (i) SF begins with a minimal language,
while PLFA begins with a computationally complete language. Why is
that? Apparently, I would regard a minimal language more pedagogical.
(ii) Why lambda calculus instead of Hoare Logic? Was that decision
solely based on the proof automation limitations of Agda or is there a
pedagogical advantage on doing that?

The text is very well written.

Minor fixes: Page 14, line 41. Introduce a line break over the word
"progress."  Page 16, line 45. "and and" → "and" Page 20, line
50. "Question 3 give the students the the" → "Question 3 gives the
students the" Overall, I would like to see the text using only the
British spelling or the American spelling (but not both: e.g. flavour,
behavior).

________________

Data in Brief (optional):

We invite you to convert your supplementary data (or a part of it)
into an additional journal publication in Data in Brief, a
multi-disciplinary open access journal. Data in Brief articles are a
fantastic way to describe supplementary data and associated metadata,
or full raw datasets deposited in an external repository, which are
otherwise unnoticed. A Data in Brief article (which will be reviewed,
formatted, indexed, and given a DOI) will make your data easier to
find, reproduce, and cite.

You can submit to Data in Brief via the Science of Computer
Programming submission system when you upload your revised Science of
Computer Programming manuscript. To do so, complete the template and
follow the co-submission instructions found here:
www.elsevier.com/dib-template. If your Science of Computer Programming
manuscript is accepted, your Data in Brief submission will
automatically be transferred to Data in Brief for editorial review and
publication.

Please note: an open access Article Publication Charge (APC) is
payable by the author or research funder to cover the costs associated
with publication in Data in Brief and ensure your data article is
immediately and permanently free to access by all. For the current APC
see:
www.elsevier.com/journals/data-in-brief/2352-3409/open-access-journal

Please contact the Data in Brief editorial office at
dib-me@elsevier.com or visit the Data in Brief homepage
(www.journals.elsevier.com/data-in-brief/) if you have questions or
need further information.

Note: While submitting the revised manuscript, please double check the
author names provided in the submission so that authorship related
changes are made in the revision stage. If your manuscript is
accepted, any authorship change will involve approval from co-authors
and respective editor handling the submission and this may cause a
significant delay in publishing your manuscript.

METHODSX CO-SUBMISSION (OPTIONAL)

If you have customized (a) research method(s) for the project
presented in your Science of Computer Programming article, you are
invited to submit this part of your work as MethodsX article alongside
your revised research article. MethodsX is an independent journal that
publishes the work you have done to develop research methods to your
specific needs or setting. This is an opportunity to get full credit
for the time and money you may have spent on developing research
methods, and to increase the visibility and impact of your work.

How does it work?  1) Fill in the MethodsX article template:
https://www.elsevier.com/MethodsX-template 2) Place all MethodsX files
(including graphical abstract, figures and other relevant files) into
a .zip file and upload this as a 'Method Details (MethodsX) ' item
alongside your revised Science of Computer Programming
manuscript. Please ensure all of your relevant MethodsX documents are
zipped into a single file.  3) If your Science of Computer Programming
research article is accepted, your MethodsX article will automatically
be transferred to MethodsX, where it will be reviewed and published as
a separate article upon acceptance.  MethodsX is a fully Open Access
journal, the publication fee is only 520 US$.

Questions?  Please contact the MethodsX team at methodsx@elsevier.com.
Example MethodsX articles can be found here:
http://www.sciencedirect.com/science/journal/22150161 Yours sincerely,

INTERACTIVE PLOT VIEWER (OPTIONAL)

Science of Computer Programming features the Interactive Plot Viewer,
see: http://www.elsevier.com/interactiveplots. Interactive Plots
provide easy access to the data behind plots. To include one with your
article, please prepare a .csv file with your plot data and test it
online at
http://authortools.elsevier.com/interactiveplots/verification before
submission as supplementary material.  PLEASE NOTE: Science of
Computer Programming would like to enrich online articles by
displaying interactive figures that help the reader to visualize and
explore your research results. For this purpose, we would like to
invite you to upload figures in the MATLAB .FIG file format as
supplementary material to our online submission system. Elsevier will
generate interactive figures from these files and include them with
the online article on SciVerse ScienceDirect. If you wish, you can
submit .FIG files along with your revised submission.

________________

For further assistance, please visit our customer support site at
http://help.elsevier.com/app/answers/list/p/7923. Here you can search
for solutions on a range of topics, find answers to frequently asked
questions and learn more about EES via interactive tutorials. You will
also find our 24/7 support contact details should you need any further
assistance from one of our customer support representatives.

### 2020-07-13 papers/scp/PLFA.tex

@@
 complete language, PCF, while SF begins with a minimal language, simply-typed
-lambda calculus with booleans.  PLFA does not include type annotations in terms,
+lambda calculus with booleans.  We chose PCF because it lets us use
+the same examples, based on addition and multiplication, for the early
+chapters of both Part I and Part II.
+PLFA does not include type annotations in terms,
 and uses bidirectional type inference, while SF has terms with unique types and
@@
 There is more material that would be desirable to include in PLFA which was not
-due to limits of time, including mutable references, logical relations, System~F, and
-pure type systems. We'd especially like to include pure type systems as they
+due to limits of time, including mutable references, System~F, logical
+relations for parametricity, and pure type systems.
+We would especially like to include pure type systems as they
 provide the readers with a formal model close to the dependent types used in the
@@
 switching the order of the arguments.  We had to copy the cases
-suitably permuted. It would be preferable to reinvoke the proof on
+nsuitably permuted. It would be preferable to reinvoke the proof on
 switched arguments, but this would not pass Agda's termination checker
 since swapping the arguments doesn't yield a recursive call on
-structurally smaller arguments.  We suspect tactics could cut down the
-proof significantly. We tried to compare with SF's proof that reduction
-is deterministic, but failed to find that proof.
+structurally smaller arguments.  The proof of determinism in SF
+(Chapter Norm) is for a different language of comparable size,
+and has a comparable size.
+
+% We suspect tactics could cut down the
+% proof significantly. 
+% We tried to compare with SF's proof that reduction
+% is deterministic, but failed to find that proof.
 
@@
 intelligible.  We expect the students skipped over many such proofs.  This
-particular proof forms the basis for a question of the mock exam and the past
-exams, so we expect most students will look at this one if not all the
-others.
+particular proof forms the basis for a question on several exams,
+so we expect most students will look at this one if not all the others.
 
@@
 width.  Long arrows are necessarily more than a single character wide.
-Instead, we compose reduction —→ from an em dash — and an arrow →.
+Instead of the unicode long arrow,
+we compose reduction —→ from an em dash — and an ordinary arrow →.
 Similarly for reflexive and transitive closure —↠.
@@
 modeling simply-typed lambda calculus.  It first presents
-terms with named variables and and extrinsic typing relation and
+terms with named variables and extrinsic typing relation and
 then shifts to terms with de Bruijn indices that are intrinsically
@@
 requires one first define substitution and reduction and then prove they
-preserve types, the intrinsic approach establishes substitution at the same time
-it defines substitution and reduction.
+preserve types, the intrinsic approach establishes substitution
+preserves types at the same time it defines substitution and reduction.
 
@@
 variables vs de Bruijn indices, and the choice of extrinsic vs
-intrinsic typing may be made independently.  There are synergies
-between the two.  Manipulation of de Bruijn indices can be notoriously
+intrinsic typing may be made independently.  But the two designs
+we chose work well, while the other two are problematic.
+Manipulation of de Bruijn indices can be notoriously
 error-prone without intrinsic types to give assurance of correctness.

### 2020-10-23 hs/Main.hs
Switch to Hakyll as the build system for PLFA (#540)

@@ -1,0 +1,375 @@
+{-# LANGUAGE OverloadedStrings #-}
+{-# LANGUAGE TupleSections #-}
+
+import           Control.Monad ((<=<), forM_)
+import qualified Data.ByteString.Lazy as BL
+import           Data.List as L
+import           Data.List.Extra as L
+import           Data.Maybe (fromMaybe)
+import           Data.Ord (comparing)
+import qualified Data.Text as T
+import           Hakyll
+import           Hakyll.Web.Agda
+import           Hakyll.Web.Template.Context.Metadata
+import           Hakyll.Web.Sass
+import           Hakyll.Web.Routes.Permalink
+import           System.FilePath ((</>), takeDirectory)
+import           Text.Pandoc as Pandoc
+import           Text.Pandoc.Filter
+import           Text.Printf (printf)
+import           Text.Read (readMaybe)
+
+--------------------------------------------------------------------------------
+-- Configuration
+--------------------------------------------------------------------------------
+
+tocContext :: Context String -> Context String
+tocContext ctx = Context $ \k a _ -> do
+  m <- makeItem <=< getMetadata $ "src/plfa/toc.metadata"
+  unContext (objectContext ctx) k a m
+
+siteContext :: Context String
+siteContext = mconcat
+  [ constField "pagetitle" "Programming Language Foundations in Agda"
+  , constField "pageurl" "https://plfa.github.io"
+  , constField "description" "An introduction to programming language theory using the proof assistant Agda."
+  , constField "language" "en-US"
+  , constField "rights" "Creative Commons Attribution 4.0 International License"
+  , constField "rights_url" "https://creativecommons.org/licenses/by/4.0/"
+  , constField "repository" "plfa/plfa.github.io"
+  , constField "branch" "dev"
+  , modificationTimeField "modified" "%0Y-%m-%dT%H:%M:%SZ"
+  , field "source" (return . toFilePath . itemIdentifier)
+  , listField "authors" defaultContext $ mapM load
+      [ "authors/wadler.metadata"
+      , "authors/wenkokke.metadata"
+      , "authors/jsiek.metadata"
+      ]
+  , constField "google_analytics" "UA-125055580-1"
+  , defaultContext
+  ]
+
+siteSectionContext :: Context String
+siteSectionContext = mconcat
+  [ titlerunningField
+  , subtitleField
+  , siteContext
+  ]
+
+acknowledgementsContext :: Context String
[…]

### 2020-10-23 hs/Hakyll/Web/Template/Numeric.hs
Fixed rendering errors.

@@ -1,0 +1,23 @@
+{-# LANGUAGE TupleSections #-}
+
+module Hakyll.Web.Template.Numeric
+  ( byNumericFieldAsc
+  , byNumericFieldDesc
+  ) where
+
+import Hakyll
+import Data.List as L
+import Data.Maybe (fromMaybe)
+import Data.Ord (comparing)
+import Text.Read (readMaybe)
+
+byNumericFieldAsc :: MonadMetadata m => String -> [Item a] -> m [Item a]
+byNumericFieldAsc key = sortOnM $ \i -> do
+  maybeInt <- getMetadataField (itemIdentifier i) key
+  return $ fromMaybe (0 :: Int) (readMaybe =<< maybeInt)
+    where
+      sortOnM :: (Monad m, Ord k) => (a -> m k) -> [a] -> m [a]
+      sortOnM f xs = map fst . L.sortBy (comparing snd) <$> mapM (\ x -> (x,) <$> f x) xs
+
+byNumericFieldDesc :: MonadMetadata m => String -> [Item a] -> m [Item a]
+byNumericFieldDesc key is = reverse <$> byNumericFieldAsc key is

### 2020-10-23 hs/Main.hs

@@ -1,32 +1,48 @@
 {-# LANGUAGE OverloadedStrings #-}
-{-# LANGUAGE TupleSections #-}
 
 import           Control.Monad ((<=<), forM_)
 import qualified Data.ByteString.Lazy as BL
-import           Data.List as L
-import           Data.List.Extra as L
-import           Data.Maybe (fromMaybe)
-import           Data.Ord (comparing)
+import           Data.Functor ((<&>))
 import qualified Data.Text as T
 import           Hakyll
 import           Hakyll.Web.Agda
+import           Hakyll.Web.Template.Numeric
 import           Hakyll.Web.Template.Context.Metadata
+import           Hakyll.Web.Template.Context.Title
 import           Hakyll.Web.Sass
 import           Hakyll.Web.Routes.Permalink
 import           System.FilePath ((</>), takeDirectory)
-import           Text.Pandoc as Pandoc
-import           Text.Pandoc.Filter
-import           Text.Printf (printf)
-import           Text.Read (readMaybe)
+import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
+import qualified Text.Pandoc as Pandoc
+import qualified Text.Pandoc.Filter as Pandoc (Filter(..),  applyFilters)
 
 --------------------------------------------------------------------------------
 -- Configuration
 --------------------------------------------------------------------------------
 
-tocContext :: Context String -> Context String
-tocContext ctx = Context $ \k a _ -> do
-  m <- makeItem <=< getMetadata $ "src/plfa/toc.metadata"
-  unContext (objectContext ctx) k a m
+siteReaderOptions :: ReaderOptions
+siteReaderOptions = defaultHakyllReaderOptions
+  { readerExtensions = Pandoc.extensionsFromList
+    [ Ext_all_symbols_escapable
+    , Ext_auto_identifiers
+    , Ext_backtick_code_blocks
+    , Ext_citations
+    , Ext_footnotes
+    , Ext_header_attributes
+    , Ext_implicit_header_references
+    , Ext_intraword_underscores
+    , Ext_markdown_in_html_blocks
+    , Ext_raw_html
+    , Ext_shortcut_reference_links
+    , Ext_smart
+    , Ext_superscript
+    , Ext_subscript
+    , Ext_yaml_metadata_block
+    ]
+  }
+
+siteWriterOptions :: WriterOptions
+siteWriterOptions = defaultHakyllWriterOptions
 
 siteContext :: Context String
 siteContext = mconcat
[…]

### 2020-10-23 hs/Main.hs
Changed site title.

@@ -164,8 +164,8 @@ main = do
         >>= readMarkdownWith siteReaderOptions
         <&> writeHTML5With siteWriterOptions
         >>= saveSnapshot "content"
-        >>= loadAndApplyTemplate "templates/page.html"    siteContext
-        >>= loadAndApplyTemplate "templates/default.html" siteContext
+        >>= loadAndApplyTemplate "templates/page.html"    siteSectionContext
+        >>= loadAndApplyTemplate "templates/default.html" siteSectionContext
         >>= relativizeUrls
 
   -- Build compiler for literate Agda pages
@@ -176,8 +176,8 @@ main = do
         >>= readMarkdownWith siteReaderOptions
         <&> writeHTML5With siteWriterOptions
         >>= saveSnapshot "content"
-        >>= loadAndApplyTemplate "templates/page.html"    siteContext
-        >>= loadAndApplyTemplate "templates/default.html" siteContext
+        >>= loadAndApplyTemplate "templates/page.html"    siteSectionContext
+        >>= loadAndApplyTemplate "templates/default.html" siteSectionContext
         >>= relativizeUrls
 
   -- Run Hakyll

### 2020-10-23 hs/Hakyll/Web/Agda.hs
Publishing web branch...

@@ -1,196 +1,0 @@
-{-# LANGUAGE LambdaCase #-}
-{-# LANGUAGE OverloadedStrings #-}
-{-# LANGUAGE RecordWildCards   #-}
-
-module Hakyll.Web.Agda
-  ( agdaCompilerWith
-  , agdaVerbosityQuiet
-  , CommandLineOptions(..)
-  , PragmaOptions(..)
-  , defaultAgdaOptions
-  , defaultAgdaPragmaOptions
-  , mkFixStdlibLink
-  , mkFixLocalLink
-  , agdaModule
-  , agdaModuleFromPath
-  ) where
-
-import qualified Agda.Main as Agda
-import           Agda.Interaction.Options
-import qualified Agda.Interaction.Highlighting.HTML as Agda (generateHTML)
-import qualified Agda.Utils.Trie as Trie (singleton)
-import           Control.Exception (catchJust)
-import           Control.Monad (forM, void)
-import qualified Data.ByteString as B
-import           Data.Frontmatter (parseYamlFrontmatterEither)
-import qualified Data.List.Extra as L
-import           Data.Maybe (fromMaybe)
-import qualified Data.Map as M
-import qualified Data.Text as T
-import qualified Data.Text.IO as T
-import qualified Data.Text.ICU as ICU
-import qualified Data.Text.ICU.Replace as ICU
-import           Data.Yaml (FromJSON(..), ToJSON(..), (.:), (.=))
-import qualified Data.Yaml as Y
-import           Hakyll
-import           Text.Printf (printf)
-import           Text.Regex.TDFA ((=~))
-import           System.Directory (createDirectoryIfMissing)
-import           System.Exit (ExitCode(..), exitFailure)
-import           System.FilePath.Find ((~~?), (==?), always, extension, fileName, find)
-import           System.FilePath ((</>), (<.>), dropExtension, dropExtensions, makeRelative, pathSeparator)
-
--- |Default Agda command-line options. Rename of `defaultOptions`.
-defaultAgdaOptions :: CommandLineOptions
-defaultAgdaOptions = defaultOptions
-
--- |Default Agda pragma options. Rename of `defaultPragmaOptions`.
-defaultAgdaPragmaOptions :: PragmaOptions
-defaultAgdaPragmaOptions = defaultPragmaOptions
-
--- |Compile literate Agda to HTML
-agdaCompilerWith :: CommandLineOptions -> Compiler (Item String)
-agdaCompilerWith agdaOptions = do
-  item <- getResourceBody
[…]

### 2020-10-23 hs/Hakyll/Web/Routes/Permalink.hs

@@ -1,25 +1,0 @@
-{-# LANGUAGE OverloadedStrings #-}
-
-module Hakyll.Web.Routes.Permalink
-  ( convertPermalink
-  , permalinkRoute
-  , permalinkRouteWithDefault
-  ) where
-
-import           Data.List (stripPrefix)
-import           Data.Maybe (fromMaybe)
-import           Hakyll
-import           System.FilePath ((</>))
-
--- By a quirk of history, permalinks for PLFA are written as, e.g., "/DeBruijn/".
--- We convert these to links by stripping the "/" prefix, and adding "index.html".
-convertPermalink :: FilePath -> FilePath
-convertPermalink link = fromMaybe link (stripPrefix "/" link) </> "index.html"
-
-permalinkRoute :: Routes
-permalinkRoute = permalinkRouteWithDefault (error "Missing field 'permalink'.")
-
-permalinkRouteWithDefault :: Routes -> Routes
-permalinkRouteWithDefault def = metadataRoute $ \metadata ->
-  maybe def (constRoute . convertPermalink) (lookupString "permalink" metadata)
-

### 2020-10-23 hs/Hakyll/Web/Template/Context/Metadata.hs

@@ -1,53 +1,0 @@
-module Hakyll.Web.Template.Context.Metadata where
-
-import           Control.Monad ((<=<))
-import           Data.Aeson (Object, Value(..))
-import qualified Data.HashMap.Strict as H
-import qualified Data.Text as T
-import qualified Data.Vector as V
-import           Hakyll
-import           Text.Printf (printf)
-
--- |Create a Context based on a JSON Object.
-objectContext :: Context String -> Context Object
-objectContext ctx = Context $ \k _ i ->
-  let o = itemBody i in
-  case H.lookup (T.pack k) o of
-    Just  v -> decodeValue ctx v
-    Nothing -> fail $ printf "Key '%s' undefined in context '%s'" k (show o)
-
-isObject :: Value -> Bool
-isObject (Object _) = True
-isObject _          = False
-
-isString :: Value -> Bool
-isString (String _) = True
-isString _          = False
-
-toObject :: MonadFail m => Value -> m Object
-toObject (Object o) = return o
-toObject v          = fail $ printf "Not an object '%s'" (show v)
-
-toString :: MonadFail m => Value -> m String
-toString (String s) = return (T.unpack s)
-toString v          = fail $ printf "Not a string '%s'" (show v)
-
--- |Decode a JSON Value to a context field.
-decodeValue :: Context String -> Value -> Compiler ContextField
-decodeValue ctx (Array  a)
-  | V.all isObject a = do
-      objs <- mapM (makeItem <=< toObject) (V.toList a)
-      return $ ListField (objectContext ctx) objs
-  | V.all isString a = do
-      items <- mapM (load . fromFilePath <=< toString) (V.toList a)
-      return $ ListField ctx items
-decodeValue _ (String s) = return . StringField $ T.unpack s
-decodeValue _ (Number n) = return . StringField $ show n
-decodeValue _ (Bool   b) = return . StringField $ show b
-decodeValue _ v          = fail $ printf "Unsupported value '%s'" (show v)
-
--- |Create a Context based on the Metadata.
-metadataContext :: Context String -> Context String
-metadataContext ctx = Context $ \k a i ->
-  unContext (objectContext ctx) k a <=< makeItem <=< getMetadata $ itemIdentifier i
-

### 2020-10-23 hs/Hakyll/Web/Template/Context/Title.hs

@@ -1,29 +1,0 @@
-module Hakyll.Web.Template.Context.Title
-  ( titlerunningField
-  , subtitleField
-  ) where
-
-import Hakyll
-import Data.List.Extra as L
-import Text.Printf (printf)
-
-titlerunningField :: Context String
-titlerunningField = Context go
-  where
-    go "titlerunning" _ i = do
-      title <- maybe (fail "No title") return =<< getMetadataField (itemIdentifier i) "title"
-      case L.stripInfix ":" title of
-        Nothing -> fail "No titlerunning/subtitle distinction"
-        Just (titlerunning, _) -> return . StringField $ titlerunning
-    go k              _ i = fail $ printf "Missing field %s in context for item %s" k (show (itemIdentifier i))
-
-
-subtitleField :: Context String
-subtitleField = Context go
-  where
-    go "subtitle" _ i = do
-      title <- maybe (fail "No title") return =<< getMetadataField (itemIdentifier i) "title"
-      case L.stripInfix ":" title of
-        Nothing -> fail "No titlerunning/subtitle distinction"
-        Just (_, subtitle) -> return . StringField $ L.trim subtitle
-    go k          _ i = fail $ printf "Missing field %s in context for item %s" k (show (itemIdentifier i))

### 2020-10-23 hs/Main.hs

@@ -1,405 +1,0 @@
-{-# LANGUAGE OverloadedStrings #-}
-
-import           Control.Monad ((<=<), forM_)
-import qualified Data.ByteString.Lazy as BL
-import           Data.Functor ((<&>))
-import qualified Data.Text as T
-import           Hakyll
-import           Hakyll.Web.Agda
-import           Hakyll.Web.Template.Numeric
-import           Hakyll.Web.Template.Context.Metadata
-import           Hakyll.Web.Template.Context.Title
-import           Hakyll.Web.Sass
-import           Hakyll.Web.Routes.Permalink
-import           System.FilePath ((</>), takeDirectory)
-import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
-import qualified Text.Pandoc as Pandoc
-import qualified Text.Pandoc.Filter as Pandoc (Filter(..),  applyFilters)
-
---------------------------------------------------------------------------------
--- Configuration
---------------------------------------------------------------------------------
-
-siteReaderOptions :: ReaderOptions
-siteReaderOptions = defaultHakyllReaderOptions
-  { readerExtensions = Pandoc.extensionsFromList
-    [ Ext_all_symbols_escapable
-    , Ext_auto_identifiers
-    , Ext_backtick_code_blocks
-    , Ext_citations
-    , Ext_footnotes
-    , Ext_header_attributes
-    , Ext_implicit_header_references
-    , Ext_intraword_underscores
-    , Ext_markdown_in_html_blocks
-    , Ext_raw_html
-    , Ext_shortcut_reference_links
-    , Ext_smart
-    , Ext_superscript
-    , Ext_subscript
-    , Ext_yaml_metadata_block
-    ]
-  }
-
-siteWriterOptions :: WriterOptions
-siteWriterOptions = defaultHakyllWriterOptions
-
-siteContext :: Context String
-siteContext = mconcat
-  [ constField "pagetitle" "Programming Language Foundations in Agda"
-  , constField "pageurl" "https://plfa.github.io"
-  , constField "description" "An introduction to programming language theory using the proof assistant Agda."
-  , constField "language" "en-US"
-  , constField "rights" "Creative Commons Attribution 4.0 International License"
-  , constField "rights_url" "https://creativecommons.org/licenses/by/4.0/"
-  , constField "repository" "plfa/plfa.github.io"
-  , constField "branch" "dev"
[…]

### 2020-10-23 hs/UpdateContributors.hs

@@ -1,209 +1,0 @@
-{-# LANGUAGE RecordWildCards #-}
-{-# LANGUAGE OverloadedStrings #-}
-
-module Main where
-
-import           Control.Monad (forM, forM_)
-import qualified Data.ByteString as B
-import qualified Data.ByteString.Char8 as BC
-import           Data.Frontmatter (parseYamlFrontmatterEither)
-import           Data.Map (Map)
-import qualified Data.Map as M
-import           Data.Maybe (fromMaybe, mapMaybe)
-import           Data.String (IsString(..))
-import           Data.Text (Text)
-import qualified Data.Text as T
-import qualified Data.Vector as V
-import           Data.Yaml (FromJSON(..), ToJSON(..), (.:), (.:?), (.=))
-import qualified Data.Yaml as Y
-import qualified GitHub as GH
-import           System.Directory (createDirectoryIfMissing)
-import           System.Exit (exitFailure)
-import           System.FilePath ((</>), (<.>))
-import           System.FilePath.Glob (namesMatching)
-import           System.Environment (lookupEnv)
-import           Text.Printf (printf)
-
-
--- * Configuration
-
-authorDir, contributorDir :: FilePath
-authorDir      = "authors"
-contributorDir = "contributors"
-
-githubOwner, githubRepo :: Text
-githubOwner  = "plfa"
-githubRepo   = "plfa.github.io"
-
-githubErrors :: [Text]
-githubErrors = ["invalid-email-address"]
-
-
--- * Main
-
-main :: IO ()
-main = do
-  -- Get the GitHub authentication token from `GITHUB_TOKEN`
-  auth <- getAuth
-
-  -- Read the current set of authors from `authorDir`
-  authors <- readAuthors authorDir
-
-  -- Read the current set of contributors from `contributorDir`
-  localContributors <- readContributors contributorDir
-  remoteContributorAndAuthors <-
-    getContributors auth (GH.mkOwnerName githubOwner) (GH.mkRepoName githubRepo)
-
-  -- Filter the contributor list by the authors
-  let authorGithubs :: [Text]
-      authorGithubs = map authorGithub authors
-
-  let isAuthorOrError :: Contributor -> Bool
-      isAuthorOrError Contributor{..} =
-        contributorGithub `elem` authorGithubs || contributorGithub `elem` githubErrors
-
-  let remoteContributors :: [Contributor]
[…]

### 2020-10-23 versions/20.07/courses/tspl/make_assignment.sh

@@ -1,188 +1,0 @@
-#!/usr/bin/env bash
-
-# TODO:
-#   - put the exercises from each chapter in their own module
-#   - replicate the imports from that module
-#   - import that module hiding anything defined in an exercise section
-
-# This script can be used to automatically generate assignment files from PLFA source files.
-# It takes a course abbreviation, e.g. TSPL, a year, and the number of the assignment.
-# At the moment, it outputs the University of Edinburgh guidelines for good scholarly practice,
-# making it somewhat specific to courses run there, but the header should be easy to edit.
-#
-# Usage:
-#
-#   ./make_assignment.sh [COURSE_NAME] [COURSE_YEAR] [ASSIGNMENT_NUMBER] [PLFA_SOURCE_FILE...]
-
-COURSE="$1"
-shift
-
-YEAR="$1"
-shift
-
-NUM="$1"
-shift
-
-## Make assignment header
-
-cat <<-EOF
----
-title     : "Assignment$NUM: $COURSE Assignment $NUM"
-layout    : page
-permalink : /$COURSE/$YEAR/Assignment$NUM/
----
-
-\`\`\`
-module Assignment$NUM where
-\`\`\`
-
-## YOUR NAME AND EMAIL GOES HERE
-
-## Introduction
-
-You must do _all_ the exercises labelled "(recommended)".
-
-Exercises labelled "(stretch)" are there to provide an extra challenge.
-You don't need to do all of these, but should attempt at least a few.
-
-Exercises labelled "(practice)" are included for those who want extra practice.
-
-Submit your homework using the "submit" command.
-Please ensure your files execute correctly under Agda!
-
-
-## Good Scholarly Practice.
-
-Please remember the University requirement as
-regards all assessed work. Details about this can be found at:
-
-> [http://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct](http://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct)
-
-Furthermore, you are required to take reasonable measures to protect
-your assessed work from unauthorised access. For example, if you put
-any such work on a public repository then you must set access
-permissions appropriately (generally permitting access only to
-yourself, or your group in the case of group practicals).
-
-EOF
-
[…]

### 2021-08-22 hs/Main.hs
PDF generation (#575)
* include tools for pdf

* generating acknowledgements to include in the pdf

* changed makefile to call the makefile in pdf/

@@ -257,6 +257,11 @@ main = do
         >>= loadAndApplyTemplate "templates/page.html"    siteContext
         >>= loadAndApplyTemplate "templates/default.html" siteContext
         >>= prettifyUrls
+    
+    match "src/plfa/backmatter/acknowledgements.md" $ version "raw" $ do
+      route $ constRoute "acknowledgements.md"
+      compile $ getResourceBody
+        >>= applyAsTemplate acknowledgementsContext
 
     match "authors/*.metadata" $
       compile getResourceBody

### 2021-08-23 hs/Main.hs
Fixed binding issues with existential syntax.

@@ -257,7 +257,7 @@ main = do
         >>= loadAndApplyTemplate "templates/page.html"    siteContext
         >>= loadAndApplyTemplate "templates/default.html" siteContext
         >>= prettifyUrls
-    
+
     match "src/plfa/backmatter/acknowledgements.md" $ version "raw" $ do
       route $ constRoute "acknowledgements.md"
       compile $ getResourceBody

### 2021-08-24 hs/Hakyll/Web/Routes/Permalink.hs
WIP rewrite of book rendering.

@@ -3,7 +3,6 @@
 module Hakyll.Web.Routes.Permalink
   ( convertPermalink
   , permalinkRoute
-  , permalinkRouteWithDefault
   , stripIndexFile
   ) where
 
@@ -19,11 +18,12 @@ convertPermalink :: FilePath -> FilePath
 convertPermalink link = fromMaybe link (stripPrefix "/" link) </> "index.html"
 
 permalinkRoute :: Routes
-permalinkRoute = permalinkRouteWithDefault (error "Missing field 'permalink'.")
-
-permalinkRouteWithDefault :: Routes -> Routes
-permalinkRouteWithDefault def = metadataRoute $ \metadata ->
-  maybe def (constRoute . convertPermalink) (lookupString "permalink" metadata)
+permalinkRoute = metadataRoute $ \metadata ->
+  case lookupString "permalink" metadata of
+    Nothing ->
+      customRoute (\identifier -> error $ "missing field 'permalink' in metadata " <> toFilePath identifier)
+    Just permalink ->
+      constRoute (convertPermalink permalink)
 
 -- Removes "index.html" from URLs.
 stripIndexFile :: String -> String

### 2021-08-24 hs/Main.hs

@@ -1,7 +1,7 @@
+{-# LANGUAGE TypeApplications #-}
 {-# LANGUAGE OverloadedStrings #-}
 
 import           Control.Monad ((<=<), forM_)
-import qualified Data.ByteString.Lazy as BL
 import           Data.Functor ((<&>))
 import qualified Data.Text as T
 import           Hakyll
@@ -16,7 +16,6 @@ import qualified Text.CSL as CSL
 import qualified Text.CSL.Pandoc as CSL (processCites)
 import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
 import qualified Text.Pandoc as Pandoc
-import qualified Text.Pandoc.Filter as Pandoc (Filter(..),  applyFilters)
 
 --------------------------------------------------------------------------------
 -- Configuration
@@ -124,29 +123,6 @@ sassOptions = defaultSassOptions
   { sassIncludePaths = Just ["css"]
   }
 
-epubSectionContext :: Context String
-epubSectionContext = mconcat
-  [ contentField "content" "content"
-  , titlerunningField
-  , subtitleField
-  ]
-
-epubReaderOptions :: ReaderOptions
-epubReaderOptions = siteReaderOptions
-  { readerStandalone    = True
-  , readerStripComments = True
-  }
-
-epubWriterOptions :: WriterOptions
-epubWriterOptions = siteWriterOptions
-  { writerTableOfContents  = True
-  , writerTOCDepth         = 2
-  , writerEpubFonts        = [ "public/webfonts/DejaVuSansMono.woff"
-                             , "public/webfonts/FreeMono.woff"
-                             , "public/webfonts/mononoki.woff"
-                             ]
-  , writerEpubChapterLevel = 2
-  }
 
 --------------------------------------------------------------------------------
 -- Build site
@@ -198,39 +174,13 @@ main = do
   --
   -- NOTE: The order of the various match expressions is important:
   --       Special-case compilation instructions for files such as
-  --       "src/plfa/epub.md" and "src/plfa/index.md" would be overwritten
-  --       by the general purpose compilers for "src/**.md", which would
+  --       "src/plfa/index.md" would be overwritten by the general
+  --       purpose compilers for "src/**.md", which would
   --       cause them to render incorrectly. It is possible to explicitly
[…]

### 2021-08-24 hs/Hakyll/Web/Template/Context/Derived.hs
Fixed PDF generate to use toc.metadata.

@@ -1,0 +1,22 @@
+{-# LANGUAGE RankNTypes #-}
+
+module Hakyll.Web.Template.Context.Derived where
+
+import Hakyll
+
+addDerivedField
+  :: String
+  -> (forall b. Context b -> [String] -> Item b -> Compiler ContextField)
+  -> Context a
+  -> Context a
+addDerivedField key derive ctx = Context $ \k a i ->
+  if k == key then
+    derive ctx a i
+  else do
+    fld <- unContext ctx k a i
+    return $
+      case fld of
+        -- If the field is a list, recursively derive the field for any list items.
+        ListField itemCtx items -> ListField (addDerivedField key derive itemCtx) items
+        -- Otherwise, simply return the field.
+        otherFld                -> otherFld

### 2021-08-24 hs/Hakyll/Web/Template/Context/Metadata.hs

@@ -1,3 +1,5 @@
+{-# LANGUAGE OverloadedStrings #-}
+
 module Hakyll.Web.Template.Context.Metadata where
 
 import           Control.Monad ((<=<))
@@ -8,21 +10,43 @@ import qualified Data.Vector as V
 import           Hakyll
 import           Text.Printf (printf)
 
+-- |Create a context from a JSON object which loads the included file and uses it to generate a context.
+includeContext :: Context String -> Context Object
+includeContext ctx = Context $ \k a i -> do
+  let o = itemBody i
+  v <- lookupObject "include" o
+  identifier <- fromFilePath <$> toString v
+  unContext (ctx <> metadataContext ctx) k a =<< load identifier
+
 -- |Create a Context based on a JSON Object.
 objectContext :: Context String -> Context Object
-objectContext ctx = Context $ \k _ i ->
-  let o = itemBody i in
-  case H.lookup (T.pack k) o of
-    Just  v -> decodeValue ctx v
-    Nothing -> fail $ printf "Key '%s' undefined in context '%s'" k (show o)
+objectContext ctx = Context $ \k _ i -> do
+  let o = itemBody i
+  decodeValue ctx =<< lookupObject k o
 
-isObject :: Value -> Bool
-isObject (Object _) = True
-isObject _          = False
+-- |Decode a JSON Value to a context field.
+decodeValue :: Context String -> Value -> Compiler ContextField
+decodeValue ctx (Array a) = do
+  objs <- mapM (makeItem <=< toObject) (V.toList a)
+  return $ ListField (includeContext ctx <> objectContext ctx) objs
+decodeValue _ctx (String s) = return . StringField $ T.unpack s
+decodeValue _ctx (Number n) = return . StringField $ show n
+decodeValue _ctx (Bool   b) = return . StringField $ show b
+decodeValue _ctx v          = fail $ printf "Unsupported value '%s'" (show v)
 
-isString :: Value -> Bool
-isString (String _) = True
-isString _          = False
+-- |Create a Context based on the Metadata.
+metadataContext :: Context String -> Context String
+metadataContext ctx = Context $ \k a i -> do
+  let identifier = itemIdentifier i
+  metadata <- getMetadata identifier
+  item <- makeItem metadata
+  unContext (objectContext ctx) k a item
+
+lookupObject :: String -> Object -> Compiler Value
+lookupObject k o = maybe ifNotFound ifFound (H.lookup (T.pack k) o)
+  where
[…]

### 2021-08-24 hs/Main.hs

@@ -3,15 +3,17 @@
 
 import           Control.Monad ((<=<), forM_)
 import           Data.Functor ((<&>))
+import           Data.List (isPrefixOf)
 import qualified Data.Text as T
 import           Hakyll
 import           Hakyll.Web.Agda
 import           Hakyll.Web.Routes.Permalink
 import           Hakyll.Web.Template.Numeric
+import           Hakyll.Web.Template.Context.Derived
 import           Hakyll.Web.Template.Context.Metadata
 import           Hakyll.Web.Template.Context.Title
 import           Hakyll.Web.Sass
-import           System.FilePath ((</>), takeDirectory)
+import           System.FilePath ((</>), takeDirectory, replaceExtensions, splitPath, joinPath)
 import qualified Text.CSL as CSL
 import qualified Text.CSL.Pandoc as CSL (processCites)
 import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
@@ -76,7 +78,8 @@ siteSectionContext = mconcat
 
 tableOfContentsContext :: Context String -> Context String
 tableOfContentsContext ctx = Context $ \k a _ -> do
-  m <- makeItem <=< getMetadata $ "src/plfa/toc.metadata"
+  metadata <- getMetadata "src/plfa/toc.metadata"
+  m <- makeItem metadata
   unContext (objectContext ctx) k a m
 
 acknowledgementsContext :: Context String
@@ -122,6 +125,32 @@ sassOptions :: SassOptions
 sassOptions = defaultSassOptions
   { sassIncludePaths = Just ["css"]
   }
+
+
+-- Convert MD_DIR/%.md to LAGDA_TEX_DIR/%.lagda.tex or TEX_DIR/%.tex
+--
+-- NOTE: This logic is partially duplicated in book/pdf.mk:TEX_PATH.
+--
+-- NOTE: This function assumes pdf.tex will be at TEX_DIR/.
+--
+addTexPath :: Context a -> Context a
+addTexPath = addDerivedField "tex_path" deriveTexPath
+  where
+    deriveTexPath :: Context a -> [String] -> Item a -> Compiler ContextField
+    deriveTexPath ctx a i = do
+      fld <- unContext ctx "include" a i
+      case fld of
+        StringField includePath -> return $ StringField (texPath includePath)
+        _ -> fail "Key 'include' does not return a String"
+
+    texPath :: FilePath -> FilePath
+    texPath fnDotMd
[…]

### 2021-08-24 hs/Main.hs
Fixed PDF and EPUB generation to use toc.metadata.

@@ -2,9 +2,12 @@
 {-# LANGUAGE OverloadedStrings #-}
 
 import           Control.Monad ((<=<), forM_)
+import qualified Data.ByteString.Char8 as BS
+import           Data.Char (toLower)
 import           Data.Functor ((<&>))
-import           Data.List (isPrefixOf)
+import           Data.List (isPrefixOf, stripPrefix)
 import qualified Data.Text as T
+import qualified Data.Yaml as Y
 import           Hakyll
 import           Hakyll.Web.Agda
 import           Hakyll.Web.Routes.Permalink
@@ -18,6 +21,9 @@ import qualified Text.CSL as CSL
 import qualified Text.CSL.Pandoc as CSL (processCites)
 import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
 import qualified Text.Pandoc as Pandoc
+import           Text.Pandoc.Definition (Block(..))
+import           Text.Pandoc.Walk (walk)
+import           Text.Printf
 
 --------------------------------------------------------------------------------
 -- Configuration
@@ -66,7 +72,7 @@ siteContext = mconcat
       , "authors/jsiek.metadata"
       ]
   , constField "google_analytics" "UA-125055580-1"
-  , defaultContext
+  , addAnchor defaultContext
   ]
 
 siteSectionContext :: Context String
@@ -77,10 +83,10 @@ siteSectionContext = mconcat
   ]
 
 tableOfContentsContext :: Context String -> Context String
-tableOfContentsContext ctx = Context $ \k a _ -> do
-  metadata <- getMetadata "src/plfa/toc.metadata"
-  m <- makeItem metadata
-  unContext (objectContext ctx) k a m
+tableOfContentsContext ctx = Context $ \k a _ ->
+  unContext (objectContext ctx) k a
+    =<< makeItem
+    =<< getMetadata "src/plfa/toc.metadata"
 
 acknowledgementsContext :: Context String
 acknowledgementsContext = mconcat
@@ -145,13 +151,48 @@ addTexPath = addDerivedField "tex_path" deriveTexPath
 
     texPath :: FilePath -> FilePath
     texPath fnDotMd
-      | fnDotMd == "README.md"                       = "plfa/frontmatter/README.tex"
+      | fnDotMd == "README.md"                         = "plfa/frontmatter/README.tex"
       | any (`isPrefixOf` fnDotMd) ["src/", "book/"] = dropTopDirectory (replaceExtensions fnDotMd ".tex")
[…]

### 2021-08-24 hs/Hakyll/Web/Agda.hs
EPUB and PDF now more-or-less work

@@ -5,6 +5,7 @@
 module Hakyll.Web.Agda
   ( agdaCompilerWith
   , agdaVerbosityQuiet
+  , compileAgdaWith
   , CommandLineOptions(..)
   , PragmaOptions(..)
   , defaultAgdaOptions
@@ -50,8 +51,12 @@ defaultAgdaPragmaOptions = defaultPragmaOptions
 
 -- |Compile literate Agda to HTML
 agdaCompilerWith :: CommandLineOptions -> Compiler (Item String)
-agdaCompilerWith agdaOptions = cached "Hakyll.Web.Agda.agdaCompilerWith" $ do
-  item <- getResourceBody
+agdaCompilerWith agdaOptions =
+  getResourceBody >>= compileAgdaWith agdaOptions
+
+-- |Compile literate Agda to HTML
+compileAgdaWith :: CommandLineOptions -> Item String -> Compiler (Item String)
+compileAgdaWith agdaOptions item = cached "Hakyll.Web.Agda.agdaCompilerWith" $ do
   let agdaPath = toFilePath (itemIdentifier item)
   let moduleName = agdaModule (itemBody item)
   TmpFile tmpPath <- newTmpFile ".lock"
@@ -120,7 +125,7 @@ readStdlibVersion stdlibPath = do
   changelog <- T.readFile changelogPath
   let versionLine = head (T.lines changelog)
   case T.stripPrefix "Version " versionLine of
-    Just versionStr -> return . T.unpack $ "v" <> T.strip versionStr
+    Just versionStr -> return $ T.unpack ("v" <> T.strip versionStr)
     Nothing -> error $ printf "Could not read version from '%s'" changelogPath
 
 -- |Fix references to the Agda standard library.

### 2021-08-24 hs/Hakyll/Web/Template/Context/Derived.hs

@@ -3,6 +3,7 @@
 module Hakyll.Web.Template.Context.Derived where
 
 import Hakyll
+import Text.Printf
 
 addDerivedField
   :: String
@@ -20,3 +21,13 @@ addDerivedField key derive ctx = Context $ \k a i ->
         ListField itemCtx items -> ListField (addDerivedField key derive itemCtx) items
         -- Otherwise, simply return the field.
         otherFld                -> otherFld
+
+
+
+-- Retrieve a String from the context
+getString :: String -> Context a -> [String] -> Item a -> Compiler String
+getString key ctx a i = do
+  fld <- unContext ctx key a i
+  case fld of
+    StringField str -> return str
+    _ -> fail $ printf "Key '%s' does not return a String" key

### 2021-08-24 hs/Main.hs

@@ -1,4 +1,3 @@
-{-# LANGUAGE TypeApplications #-}
 {-# LANGUAGE OverloadedStrings #-}
 
 import           Control.Monad ((<=<), forM_)
@@ -7,6 +6,8 @@ import           Data.Char (toLower)
 import           Data.Functor ((<&>))
 import           Data.List (isPrefixOf, stripPrefix)
 import qualified Data.Text as T
+import qualified Data.Text.ICU as RE
+import qualified Data.Text.ICU.Replace as T
 import qualified Data.Yaml as Y
 import           Hakyll
 import           Hakyll.Web.Agda
@@ -21,8 +22,6 @@ import qualified Text.CSL as CSL
 import qualified Text.CSL.Pandoc as CSL (processCites)
 import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
 import qualified Text.Pandoc as Pandoc
-import           Text.Pandoc.Definition (Block(..))
-import           Text.Pandoc.Walk (walk)
 import           Text.Printf
 
 --------------------------------------------------------------------------------
@@ -79,6 +78,7 @@ siteSectionContext :: Context String
 siteSectionContext = mconcat
   [ titlerunningField
   , subtitleField
+  , addShiftedBody "raw" (contentField "raw" "raw")
   , siteContext
   ]
 
@@ -133,67 +133,6 @@ sassOptions = defaultSassOptions
   }
 
 
--- Convert MD_DIR/%.md to LAGDA_TEX_DIR/%.lagda.tex or TEX_DIR/%.tex
---
--- NOTE: This logic is partially duplicated in book/pdf.mk:TEX_PATH.
---
--- NOTE: This function assumes pdf.tex will be at TEX_DIR/.
---
-addTexPath :: Context a -> Context a
-addTexPath = addDerivedField "tex_path" deriveTexPath
-  where
-    deriveTexPath :: Context a -> [String] -> Item a -> Compiler ContextField
-    deriveTexPath ctx a i = do
-      fld <- unContext ctx "include" a i
-      case fld of
-        StringField includePath -> return $ StringField (texPath includePath)
-        _ -> fail "Key 'include' does not return a String"
-
-    texPath :: FilePath -> FilePath
-    texPath fnDotMd
-      | fnDotMd == "README.md"                         = "plfa/frontmatter/README.tex"
[…]

### 2021-09-02 hs/Main.hs
Progress towards nicer EPUB generation, but currently broken.

@@ -1,12 +1,15 @@
 {-# LANGUAGE OverloadedStrings #-}
 
 import           Control.Monad ((<=<), (>=>), forM_)
+import           Data.Binary (Binary)
 import           Data.Char (toLower)
+import           Data.Function ((&))
 import           Data.Functor ((<&>))
 import           Data.List (isPrefixOf, stripPrefix)
 import qualified Data.Text as T
 import qualified Data.Text.ICU as RE
 import qualified Data.Text.ICU.Replace as T
+import           Data.Typeable (Typeable)
 import           Hakyll
 import           Hakyll.Web.Agda
 import           Hakyll.Web.Routes.Permalink
@@ -25,6 +28,22 @@ import           Text.Printf
 --------------------------------------------------------------------------------
 -- Configuration
 --------------------------------------------------------------------------------
+
+-- General configuration
+config :: Configuration
+config = defaultConfiguration
+  { destinationDirectory = "_site"
+  , storeDirectory       = "_build"
+  , tmpDirectory         = "_build" </> "tmp"
+  }
+
+-- | Hakyll deletes 'tmpDirectory' after each run. Since we'd like to use
+--   'tmpDirectory' ourselves, we need Hakyll to use a subdirectory.
+hakyllConfig :: Configuration
+hakyllConfig = config
+  { storeDirectory = storeDirectory config </> "hakyll"
+  , tmpDirectory   = tmpDirectory config </> "hakyll"
+  }
 
 siteReaderOptions :: ReaderOptions
 siteReaderOptions = defaultHakyllReaderOptions
@@ -165,8 +184,8 @@ main = do
         csl <- load cslFileName
         bib <- load bibFileName
         getResourceBody
-          >>= saveSnapshot "raw"
           >>= maybeCompileAgda maybeOpts
+          >>= saveSnapshot "raw"
           >>= readMarkdownWith siteReaderOptions
           >>= processCites csl bib
           <&> writeHTML5With siteWriterOptions
@@ -184,7 +203,7 @@ main = do
   --       exclude such files using `complement` patterns, but this vastly
   --       complicates the match patterns.
   --
-  hakyll $ do
+  hakyllWith hakyllConfig $ do
 
     -- Compile Table of Contents
     match "src/plfa/index.md" $ do
@@ -318,31 +337,33 @@ main = do
         compile copyFileCompiler
 
 
[…]

### 2021-09-06 app/Main.hs
Progress on rewrite using Shake.

@@ -1,68 +1,123 @@
-import Data.List qualified as L
+import Control.Exception (assert)
+import Control.Monad (forM_)
+import Data.Map qualified as M
 import Development.Shake
+import Development.Shake.Agda
+import Development.Shake.CSS
 import Development.Shake.FilePath
-import PLFA.Agda
-import PLFA.CSS
-import PLFA.Data.Text qualified as T
+import Development.Shake.Sass
+import Development.Shake.Text qualified as T
+import General.Extra
+import System.FilePath.Find
 import Text.Pandoc
 import Text.Pandoc.Highlighting (tango)
+import Text.Printf
 
-cacheDir, htmlDir, siteDir, srcDir, stdlibSrcDir :: FilePath
-cacheDir     = "_build"
-htmlDir      = cacheDir </> "html"
-texDir       = cacheDir </> "tex"
-siteDir      = "_site"
-srcDir       = "src"
-stdlibSrcDir = "standard-library" </> "src"
-agdaOpts     = AgdaOpts
-  { format = undefined,
-    tmpDir = cacheDir </> "tmp",
-    includePaths = [srcDir, stdlibSrcDir],
-    inputFile = undefined
+
+-- Directories
+
+cacheDir      = "_build"
+htmlCacheDir  = cacheDir </> "html"
+mdCacheDir    = cacheDir </> "md"
+texCacheDir   = cacheDir </> "tex"
+siteDir       = "_site"
+srcDir        = "src"
+staticSrcDir  = "public"
+cssSrcDir     = "css"
+stdlibDir     = "standard-library"
+
+-- Options
+
+shakeOpts = shakeOptions
+  { shakeFiles = cacheDir </> "shake"
   }
 
-opts :: ShakeOptions
-opts = shakeOptions { shakeFiles = cacheDir </> "shake" }
+
+-- Build rules
+
+-- | Convert paths from Markdown source files to highlighted TeX files.
+texCachePath :: FilePath -> FilePath
+texCachePath filePath = texCacheDir </> replaceExtensions filePath "tex"
+
+-- | Convert paths from Markdown source files to highlighted Markdown files.
+mdCachePath :: FilePath -> FilePath
+mdCachePath filePath = mdCacheDir </> replaceExtensions filePath "md"
 
 main :: IO ()
-main = shakeArgs opts $ do
+main = do
 
-  -- compile static files
+  -- Enumerate files
 
-  -- Compile templates
[…]

### 2021-09-06 lib/Development/Shake/Agda.hs

@@ -1,0 +1,336 @@
+{-# LANGUAGE OverloadedStrings #-}
+
+module Development.Shake.Agda
+  ( Format (..)
+  , AgdaOpts (..)
+  , AgdaLib (..)
+  , highlightAgdaWith
+  , standardLibraryAgdaLib
+  , prepareFixLinks
+  , prepareFixExternalLinks
+  , prepareFixLocalLinksWithPermalink
+  ) where
+
+import Agda.Main
+import Agda.Compiler.Backend (Backend, parseBackendOptions, backendInteraction)
+import Agda.Interaction.Options (defaultOptions)
+import Agda.Interaction.Highlighting.HTML (htmlBackend)
+import Agda.Interaction.Highlighting.LaTeX (latexBackend)
+import Agda.Utils.FileName (absolute)
+import Control.Exception (throwIO, handle)
+import Control.Monad (forM, forM_)
+import Data.ByteString qualified as B
+import Data.Default.Class (Default (..))
+import Data.Frontmatter (parseYamlFrontmatterEither)
+import Data.IORef (newIORef, readIORef, modifyIORef')
+import Data.List qualified as L
+import Data.List.Extra qualified as L
+import Data.Map (Map)
+import Data.Map qualified as M
+import Data.Maybe (listToMaybe, catMaybes, fromMaybe)
+import Data.Monoid (Endo(..))
+import Data.Text (Text)
+import Data.Text qualified as T
+import Data.Text.IO qualified as T
+import Data.Text.ICU qualified as ICU
+import Data.Text.ICU.Replace qualified as ICU
+import Data.Yaml (FromJSON(..), ToJSON(..), (.:), (.=))
+import Data.Yaml qualified as Y
+import Development.Shake (Action, liftIO)
+import Development.Shake.FilePath
+import General.Extra (liftExcept, liftEither, liftMaybe)
+import System.Exit
+import System.Directory (doesFileExist, copyFile)
+import System.FilePath.Find ((~~?), (==?), always, extension, fileName, find)
+import Text.Printf
+
+
+--------------------------------------------------------------------------------
+-- Compile literate Agda to Markdown/HTML or LaTeX
+--------------------------------------------------------------------------------
+
+data Format
+  = Markdown
+  | LaTeX
+  deriving (Eq, Ord)
+
+type Url = String
+
+data AgdaLib
+  = AgdaLib
+  { -- | The source directory.
+    --
+    --   NOTE: We don't parse .agda-lib files, and simply expect to be pointed
[…]

### 2021-09-06 lib/Development/Shake/ByteString.hs

@@ -1,0 +1,16 @@
+module Development.Shake.ByteString
+  ( BS.ByteString
+  , writeFile'
+  ) where
+
+import Data.ByteString qualified as BS
+import Development.Shake (Action, liftIO)
+import Development.Shake.FilePath (takeDirectory)
+import General.Extra
+import System.Directory
+
+writeFile' :: FilePath -> BS.ByteString -> Action ()
+writeFile' fp content = liftIO $ do
+  createDirectoryIfMissing True (takeDirectory fp)
+  removeFile_ fp
+  BS.writeFile fp content

### 2021-09-06 lib/Development/Shake/CSS.hs

@@ -1,0 +1,35 @@
+{-# OPTIONS_GHC -fno-warn-orphans #-}
+
+module Development.Shake.CSS
+  ( -- * Minify CSS with hasmin
+    minifyCSS
+  , minifyCSSWith
+  , Hasmin.Config (..)
+  , Hasmin.ColorSettings (..)
+  , Hasmin.DimensionSettings (..)
+  , Hasmin.GradientSettings (..)
+  , Hasmin.FontWeightSettings (..)
+  , Hasmin.LetterCase (..)
+  , Hasmin.SortingMethod (..)
+  , Hasmin.RulesMergeSettings (..)
+  ) where
+
+import Data.Default.Class (Default(..))
+import Data.Text (Text)
+import Development.Shake
+import Hasmin qualified
+
+instance Default Hasmin.Config where
+  def = Hasmin.defaultConfig
+
+-- | Minify CSS using 'Hasmin'.
+minifyCSS :: Text -> Action Text
+minifyCSS = minifyCSSWith def
+
+-- | Minify CSS with options.
+minifyCSSWith :: Hasmin.Config -> Text -> Action Text
+minifyCSSWith opts css = liftEither (Hasmin.minifyCSSWith opts css)
+
+liftEither :: MonadFail m => Either String a -> m a
+liftEither (Left  e) = fail e
+liftEither (Right a) = return a

### 2021-09-06 lib/Development/Shake/Sass.hs

@@ -1,21 +1,8 @@
-{-# OPTIONS_GHC -fno-warn-orphans #-}
-
-module PLFA.CSS
+module Development.Shake.Sass
   ( -- * Compile Sass with hsass
     compileSass
   , compileSassWith
   , SassOptions (..)
-    -- * Minify CSS with hasmin
-  , minifyCSS
-  , minifyCSSWith
-  , Hasmin.Config (..)
-  , Hasmin.ColorSettings (..)
-  , Hasmin.DimensionSettings (..)
-  , Hasmin.GradientSettings (..)
-  , Hasmin.FontWeightSettings (..)
-  , Hasmin.LetterCase (..)
-  , Hasmin.SortingMethod (..)
-  , Hasmin.RulesMergeSettings (..)
   ) where
 
 import Codec.Text.Detect qualified as Codec
@@ -28,7 +15,6 @@ import Data.Text.ICU.Convert qualified as Convert
 import Development.Shake
 import Text.Sass (SassError, SassOptions(..))
 import Text.Sass qualified as Sass
-import Hasmin qualified
 
 
 -- * Sass
@@ -43,7 +29,7 @@ compileSassWith opts filePath = do
 
   -- Compile `filePath` from Sass/SCSS to CSS
   resultOrError <- liftIO $ Sass.compileFile filePath opts
-  result <- liftEither showSassError resultOrError
+  result <- liftIO (liftEither resultOrError)
 
   -- Inform Shake of the dependencies used during compilation
   --
@@ -56,30 +42,11 @@ compileSassWith opts filePath = do
   liftIO $ toText (Sass.resultString result)
 
 
--- * CSS
-
--- | Minify CSS using 'Hasmin'.
-minifyCSS :: Text -> Action Text
-minifyCSS = minifyCSSWith def
-
--- | Minify CSS with options.
-minifyCSSWith :: Hasmin.Config -> Text -> Action Text
-minifyCSSWith opts css =
-  liftEither return (Hasmin.minifyCSSWith opts css)
-
-instance Default Hasmin.Config where
-  def = Hasmin.defaultConfig
-
-
-
 -- * Helper functions
 
-showSassError :: SassError -> Action String
-showSassError e = liftIO $ Sass.errorMessage e
-
-liftEither :: MonadFail m => (e -> m String) -> Either e a -> m a
-liftEither pretty (Left  e) = fail =<< pretty e
-liftEither _      (Right a) = return a
[…]

### 2021-09-06 lib/Development/Shake/Text.hs

@@ -1,18 +1,17 @@
-module PLFA.Data.Text
+module Development.Shake.Text
   ( T.Text
-  , writeFile
+  , writeFile'
   ) where
 
-import Prelude hiding (writeFile)
 import Data.Text qualified as T
 import Data.Text.IO qualified as T
-import Development.Shake
-import Development.Shake.FilePath
+import Development.Shake (Action, liftIO)
+import Development.Shake.FilePath (takeDirectory)
 import General.Extra
 import System.Directory
 
-writeFile :: FilePath -> T.Text -> Action ()
-writeFile fp content = liftIO $ do
+writeFile' :: FilePath -> T.Text -> Action ()
+writeFile' fp content = liftIO $ do
   createDirectoryIfMissing True (takeDirectory fp)
   removeFile_ fp
   T.writeFile fp content

### 2021-09-06 lib/General/Extra.hs

@@ -1,11 +1,19 @@
 module General.Extra
   ( removeFile_
+  , liftExcept
+  , liftEither
+  , liftMaybe
+  , dropPrefix
   ) where
 
 import Control.Exception (IOException, catch, handle)
 import Control.Monad (when)
+import Control.Monad.Except (Except, runExcept)
+import Data.List qualified as L
+import Development.Shake.FilePath
 import System.Directory (Permissions(..), getPermissions, setPermissions, removeFile)
 import System.IO.Error (isPermissionError)
+import Text.Printf
 
 -- Taken from shake hidden module General.Extra
 
@@ -22,3 +30,20 @@ catchIO = catch
 
 handleIO :: (IOException -> IO a) -> IO a -> IO a
 handleIO = handle
+
+liftExcept :: MonadFail m => (e -> String) -> Except e a -> m a
+liftExcept pretty m = liftEither pretty (runExcept m)
+
+liftEither :: MonadFail m => (e -> String) -> Either e a -> m a
+liftEither pretty (Left e) = fail (pretty e)
+liftEither _      (Right a) = return a
+
+liftMaybe :: MonadFail m => String -> Maybe a -> m a
+liftMaybe errorMessage Nothing  = fail errorMessage
+liftMaybe _            (Just a) = return a
+
+dropPrefix :: MonadFail m => FilePath -> FilePath -> m FilePath
+dropPrefix prefixPath filePath = do
+  path <- liftMaybe (printf "Cannot strip prefix '%s' from '%s'" prefixPath filePath) $
+    L.stripPrefix (splitDirectories prefixPath) (splitDirectories filePath)
+  return (joinPath path)

### 2021-09-06 lib/PLFA/Agda.hs

@@ -1,116 +1,0 @@
-module PLFA.Agda
-  ( Format (..)
-  , AgdaOpts (..)
-  , callAgdaWith
-  ) where
-
-import Agda.Main
-import Agda.Compiler.Backend (Backend, parseBackendOptions, backendInteraction)
-import Agda.Interaction.Options (defaultOptions)
-import Agda.Interaction.Highlighting.HTML (htmlBackend)
-import Agda.Interaction.Highlighting.LaTeX (latexBackend)
-import Agda.Utils.FileName (absolute)
-import Control.Exception (throwIO, handle)
-import Control.Monad.Except (Except, runExcept)
-import Data.List qualified as L
-import Data.Maybe (listToMaybe, catMaybes)
-import Data.Text (Text)
-import Data.Text qualified as T
-import Data.Text.IO qualified as T
-import Development.Shake
-import Development.Shake.FilePath
-import System.Exit (ExitCode (..))
-
-data Format
-  = Markdown
-  | LaTeX
-
-data AgdaOpts = AgdaOpts
-  { format       :: Format
-  , tmpDir       :: FilePath
-  , includePaths :: [FilePath]
-  , inputFile    :: FilePath
-  }
-
-callAgdaWith :: AgdaOpts -> Action Text
-callAgdaWith opts@AgdaOpts{..} = liftIO $ do
-  -- Call Agda
-  let backends = [getBackend format]
-  (backends, agdaOpts) <- liftExcept $ parseBackendOptions backends (mkOpts opts) defaultOptions
-  absInputFile <- absolute inputFile
-  let interactor = backendInteraction absInputFile backends
-  swallowExitSuccess $
-    runTCMPrettyErrors $
-      runAgdaWithOptions interactor "agda" agdaOpts
-  -- Guess the output path and read the output file
-  out <- outputPath opts
-  T.readFile out
-
-
--- * Helper functions
-
--- | Guess the path to which Agda writes the relevant output file.
-outputPath :: AgdaOpts -> IO FilePath
-outputPath opts@(AgdaOpts Markdown tmpDir includePaths inputFile) = do
-  moduleName <-
-    liftMaybe ("Could not guess module name for " <> inputFile) $
-      guessModuleName includePaths inputFile
-  return $ tmpDir </> moduleName <.> "md"
-outputPath opts@(AgdaOpts LaTeX tmpDir includePaths inputFile) = do
-  modulePath <-
-    liftMaybe ("Could not guess module name for " <> inputFile) $
[…]

### 2021-09-06 lib/PLFA/Data/ByteString.hs

@@ -1,17 +1,0 @@
-module PLFA.Data.ByteString
-  ( BS.ByteString
-  , writeFile
-  ) where
-
-import Prelude hiding (writeFile)
-import Data.ByteString qualified as BS
-import Development.Shake
-import Development.Shake.FilePath
-import General.Extra
-import System.Directory
-
-writeFile :: FilePath -> BS.ByteString -> Action ()
-writeFile fp content = liftIO $ do
-  createDirectoryIfMissing True (takeDirectory fp)
-  removeFile_ fp
-  BS.writeFile fp content

### 2021-09-06 app/Main.hs
More progress towards shake.

@@ -1,30 +1,40 @@
-import Control.Exception (assert)
+{-# LANGUAGE OverloadedStrings #-}
+{-# LANGUAGE TypeApplications #-}
+
+import Control.Exception (displayException)
 import Control.Monad (forM_)
+import Data.Default.Class (Default (..))
 import Data.Map qualified as M
+import Data.Text (Text)
+import Data.Text.ICU qualified as ICU
+import Data.Text.ICU.Replace qualified as ICU
 import Development.Shake
-import Development.Shake.Agda
-import Development.Shake.CSS
 import Development.Shake.FilePath
-import Development.Shake.Sass
-import Development.Shake.Text qualified as T
-import General.Extra
+import PLFA.Build.Agda
+import PLFA.Build.Metadata
+import PLFA.Build.Style.CSS
+import PLFA.Build.Style.Sass
+import PLFA.Build.Text qualified as T
+import PLFA.Util
 import System.FilePath.Find
-import Text.Pandoc
-import Text.Pandoc.Highlighting (tango)
-import Text.Printf
+import Text.Mustache as Mustache
+import Text.Mustache.Compile as Mustache
+import Text.Pandoc as Pandoc
+import Text.Pandoc.Writers.HTML as Pandoc
 
 
 -- Directories
 
 cacheDir      = "_build"
 htmlCacheDir  = cacheDir </> "html"
-mdCacheDir    = cacheDir </> "md"
+epubCacheDir  = cacheDir </> "epub"
 texCacheDir   = cacheDir </> "tex"
 siteDir       = "_site"
 srcDir        = "src"
 staticSrcDir  = "public"
 cssSrcDir     = "css"
 stdlibDir     = "standard-library"
+templateDir   = "templates"
 
 -- Options
 
@@ -32,21 +42,17 @@ shakeOpts = shakeOptions
   { shakeFiles = cacheDir </> "shake"
   }
 
+permalinkPath :: FilePath -> IO FilePath
+permalinkPath src = do
+  metadata <- getMetadata src
+  permalink <- getPermalink metadata
+  return $ siteDir </> permalink </> "index.html"
 
--- Build rules
-
--- | Convert paths from Markdown source files to highlighted TeX files.
-texCachePath :: FilePath -> FilePath
-texCachePath filePath = texCacheDir </> replaceExtensions filePath "tex"
-
--- | Convert paths from Markdown source files to highlighted Markdown files.
-mdCachePath :: FilePath -> FilePath
[…]

### 2021-09-06 lib/Development/Shake/ByteString.hs

@@ -1,16 +1,0 @@
-module Development.Shake.ByteString
-  ( BS.ByteString
-  , writeFile'
-  ) where
-
-import Data.ByteString qualified as BS
-import Development.Shake (Action, liftIO)
-import Development.Shake.FilePath (takeDirectory)
-import General.Extra
-import System.Directory
-
-writeFile' :: FilePath -> BS.ByteString -> Action ()
-writeFile' fp content = liftIO $ do
-  createDirectoryIfMissing True (takeDirectory fp)
-  removeFile_ fp
-  BS.writeFile fp content

### 2021-09-06 lib/PLFA/Build/Agda.hs

@@ -1,6 +1,6 @@
 {-# LANGUAGE OverloadedStrings #-}
 
-module Development.Shake.Agda
+module PLFA.Build.Agda
   ( Format (..)
   , AgdaOpts (..)
   , AgdaLib (..)
@@ -21,27 +21,25 @@ import Control.Exception (throwIO, handle)
 import Control.Monad (forM, forM_)
 import Data.ByteString qualified as B
 import Data.Default.Class (Default (..))
-import Data.Frontmatter (parseYamlFrontmatterEither)
-import Data.IORef (newIORef, readIORef, modifyIORef')
 import Data.List qualified as L
 import Data.List.Extra qualified as L
 import Data.Map (Map)
 import Data.Map qualified as M
-import Data.Maybe (listToMaybe, catMaybes, fromMaybe)
+import Data.Maybe (listToMaybe, catMaybes, fromMaybe, fromJust)
 import Data.Monoid (Endo(..))
 import Data.Text (Text)
 import Data.Text qualified as T
 import Data.Text.IO qualified as T
 import Data.Text.ICU qualified as ICU
 import Data.Text.ICU.Replace qualified as ICU
-import Data.Yaml (FromJSON(..), ToJSON(..), (.:), (.=))
-import Data.Yaml qualified as Y
 import Development.Shake (Action, liftIO)
 import Development.Shake.FilePath
-import General.Extra (liftExcept, liftEither, liftMaybe)
+import PLFA.Build.Metadata (getMetadata, getPermalink)
+import PLFA.Util (liftExcept, liftEither, liftMaybe)
 import System.Exit
 import System.Directory (doesFileExist, copyFile)
 import System.FilePath.Find ((~~?), (==?), always, extension, fileName, find)
+import System.IO.Temp (withTempDirectory, getCanonicalTemporaryDirectory)
 import Text.Printf
 
 
@@ -81,46 +79,48 @@ data AgdaOpts = AgdaOpts
   , tmpDir      :: FilePath
   , libraries   :: [AgdaLib]
   , inputFile   :: FilePath
-  , outputFiles :: Map Format FilePath
   }
 
 instance Default AgdaOpts where
   def = AgdaOpts
-        { formats     = error "AgdaOpts: please specify at least one format",
-          tmpDir      = error "AgdaOpts: please specify tmpDir",
-          libraries   = [],
-          inputFile   = error "AgdaOpts: please specify input file",
-          outputFiles = M.empty
+        { formats   = error "AgdaOpts: please specify at least one format",
[…]

### 2021-09-06 lib/PLFA/Build/Metadata.hs

@@ -1,0 +1,33 @@
+{-# LANGUAGE OverloadedStrings #-}
+
+module PLFA.Build.Metadata
+  ( Metadata
+  , getMetadata
+  , getPermalink
+  ) where
+
+import Data.Aeson.Types
+import Data.ByteString qualified as B
+import Data.Frontmatter (parseYamlFrontmatterEither)
+import Data.HashMap.Strict qualified as H
+import Data.Text qualified as T
+import PLFA.Util (Url, liftEither)
+
+type Metadata = Object
+
+getMetadata :: FilePath -> IO Metadata
+getMetadata inputFile = do
+  contents <- B.readFile inputFile
+  value <-
+    liftEither (\msg -> "Parse error in " <> inputFile <> ": " <> msg) $
+      parseYamlFrontmatterEither contents
+  case value of
+    Object object -> return object
+    _             -> fail $ "Expected metadata object, found " <> show value
+
+getPermalink :: Metadata -> IO Url
+getPermalink metadata =
+  case H.lookup "permalink" metadata of
+    Just (String permalink) -> return (T.unpack permalink)
+    Nothing                 -> fail $ "Key 'permalink' not found in\n" <> show metadata
+    Just value              -> fail $ "Key 'permalink' not a string in\n" <> show metadata

### 2021-09-06 lib/PLFA/Build/Style/CSS.hs

@@ -1,6 +1,6 @@
 {-# OPTIONS_GHC -fno-warn-orphans #-}
 
-module Development.Shake.CSS
+module PLFA.Build.Style.CSS
   ( -- * Minify CSS with hasmin
     minifyCSS
   , minifyCSSWith

### 2021-09-06 lib/PLFA/Build/Style/Sass.hs

@@ -1,4 +1,4 @@
-module Development.Shake.Sass
+module PLFA.Build.Style.Sass
   ( -- * Compile Sass with hsass
     compileSass
   , compileSassWith

### 2021-09-06 lib/PLFA/Build/Text.hs

@@ -1,14 +1,18 @@
-module Development.Shake.Text
+module PLFA.Build.Text
   ( T.Text
+  , readFile'
   , writeFile'
   ) where
 
 import Data.Text qualified as T
 import Data.Text.IO qualified as T
-import Development.Shake (Action, liftIO)
+import Development.Shake (Action, liftIO, need)
 import Development.Shake.FilePath (takeDirectory)
-import General.Extra
+import PLFA.Util
 import System.Directory
+
+readFile' :: FilePath -> Action T.Text
+readFile' fp = need [fp] >> liftIO (T.readFile fp)
 
 writeFile' :: FilePath -> T.Text -> Action ()
 writeFile' fp content = liftIO $ do

### 2021-09-06 lib/PLFA/Util.hs

@@ -1,5 +1,6 @@
-module General.Extra
-  ( removeFile_
+module PLFA.Util
+  ( Url
+  , removeFile_
   , liftExcept
   , liftEither
   , liftMaybe
@@ -15,7 +16,7 @@ import System.Directory (Permissions(..), getPermissions, setPermissions, remove
 import System.IO.Error (isPermissionError)
 import Text.Printf
 
--- Taken from shake hidden module General.Extra
+type Url = String
 
 removeFile_ :: FilePath -> IO ()
 removeFile_ fp =

### 2021-09-09 lib/Builder.hs
Updated build file, updated gitignore, rewrote announcements in Markdown.

@@ -9,6 +9,7 @@ import Data.Text qualified as T
 import Data.Text.ICU qualified as ICU
 import Data.Text.ICU.Replace qualified as ICU
 import PLFA.Build.Agda
+import PLFA.Build.Gather
 import PLFA.Build.Metadata
 import PLFA.Build.Pandoc as Pandoc
 import PLFA.Build.Prelude
@@ -42,22 +43,20 @@ main :: IO ()
 main = do
 
   --------------------------------------------------------------------------------
-  -- Gather input files
+  -- Gather input files and build routing table
   --------------------------------------------------------------------------------
 
-  staticFiles <- find always regularFile publicDir
-  agdaLibs    <- findAgdaLibs rootDir
-  stdlib      <- standardLibraryAgdaLib stdlibDir
-  mdFiles     <- findMdFiles rootDir
-  scssFiles   <- findSassFiles cssDir
+  staticFiles           <- gather always regularFile [publicDir]
+  agdaLibs              <- gatherAgdaLibs [courseDir, srcDir]
+  stdlib                <- readStandardLibraryAgdaLib stdlibDir
+  mdFiles               <- gatherMdFiles [courseDir, srcDir]
+  scssFiles             <- gatherSassFiles [cssDir]
+  postMdFiles           <- gatherMdFiles [postsDir]
+  lagdaMdFilesByAgdaLib <- gatherLagdaMdFilesForAgdaLibs agdaLibs
 
-  lagdaMdFilesByAgdaLib <- forM agdaLibs findLagdaMdFiles
-  permalinkRoutingTable <- getPermalinkRoutingTable siteDir mdFiles
-
-  let route :: FilePath -> FilePath
-      route inputFile = fromMaybe
-        (siteDir </> inputFile -<.> "html")
-        (M.lookup inputFile permalinkRoutingTable)
+  permalinkRoutingTable <- buildPermalinkRoutingTable siteDir mdFiles
+  let postsRoutingTable = asRoutingTable (\src -> siteDir </> src -<.> "html") postMdFiles
+  let routingTable = permalinkRoutingTable <> postsRoutingTable
 
   shakeArgs shakeOpts $ do
 
@@ -174,6 +173,17 @@ main = do
       toc <- getTocMetadata ()
       resolveIncludes getFileMetadata toc
 
+    getPostMetadata <- newCache $ \() -> do
+      posts <-
+        forM postMdFiles $ \postMdFile -> do
[…]

### 2021-09-09 lib/PLFA/Build/Agda.hs

@@ -16,10 +16,8 @@ module PLFA.Build.Agda
   , readAgdaLib'
   , readAgdaLib
   , highlightAgdaWith
-  , standardLibraryAgdaLib
+  , readStandardLibraryAgdaLib
   , prepareFixLinks
-  , prepareFixExternalLinks
-  , prepareFixLocalLinksWithPermalink
   ) where
 
 import Agda.Main
@@ -254,8 +252,8 @@ getStandardLibraryVersion dir = do
 --         is located in the 'src' subdirectory. If that ever changes, or
 --         it becomes easy to read .agda-lib files, this should be fixed.
 --
-standardLibraryAgdaLib :: FilePath -> IO AgdaLib
-standardLibraryAgdaLib dir = do
+readStandardLibraryAgdaLib :: FilePath -> IO AgdaLib
+readStandardLibraryAgdaLib dir = do
   let agdaLibFile = dir </> "standard-library.agda-lib"
   correct <- System.doesFileExist agdaLibFile
   if not correct then

### 2021-09-09 lib/PLFA/Build/Gather.hs

@@ -1,0 +1,42 @@
+module PLFA.Build.Gather where
+
+import Control.Monad (forM)
+import PLFA.Build.Agda
+import PLFA.Build.Prelude
+
+--------------------------------------------------------------------------------
+-- Helpers for gathering files
+--------------------------------------------------------------------------------
+
+allowOnly :: [FilePath] -> FilterPredicate
+allowOnly allowDirs = depth ==? 0 ||? depth ==? 1 &&? filePath ==*? allowDirs ||? depth >? 1
+
+gather :: FilterPredicate -> FilterPredicate -> [FilePath] -> IO [FilePath]
+gather p1 p2 dirs = concat <$> traverse (find p1 p2) dirs
+
+gatherAgdaLibs :: [FilePath] -> IO [AgdaLib]
+gatherAgdaLibs dirs = do
+  agdaLibFiles <- gather always (regularFile &&? extensions ==? ".agda-lib") dirs
+  traverse readAgdaLib agdaLibFiles
+
+gatherMdFiles :: [FilePath] -> IO [FilePath]
+gatherMdFiles dirs =
+  gather always (regularFile &&? extension ==? ".md") dirs
+
+gatherLagdaMdFiles :: [FilePath] -> IO [FilePath]
+gatherLagdaMdFiles dirs =
+  gather always (regularFile &&? extensions ==? ".lagda.md") dirs
+
+gatherLagdaMdFilesForAgdaLib :: AgdaLib -> IO [FilePath]
+gatherLagdaMdFilesForAgdaLib agdaLib =
+  gatherLagdaMdFiles (libIncludes agdaLib)
+
+gatherLagdaMdFilesForAgdaLibs :: [AgdaLib] -> IO [(AgdaLib, [FilePath])]
+gatherLagdaMdFilesForAgdaLibs agdaLibs =
+  forM agdaLibs $ \agdaLib -> do
+    lagdaMdFiles <- gatherLagdaMdFilesForAgdaLib agdaLib
+    return (agdaLib, lagdaMdFiles)
+
+gatherSassFiles :: [FilePath] -> IO [FilePath]
+gatherSassFiles dirs =
+  gather always (regularFile &&? extensions ==? ".scss") dirs

### 2021-09-09 lib/PLFA/Build/Metadata.hs

@@ -4,13 +4,17 @@ module PLFA.Build.Metadata
     readYaml,
     readYamlFrontmatter',
     readYamlFrontmatter,
+    readFileWithMetadata',
+    readFileWithMetadata,
     (^.),
     (.~),
     (&),
     Author (..),
     Contributor (..),
     lastModified,
+    dateFromFilePath,
     sourceFile,
+    teaser,
     addTitleVariants,
     resolveIncludes
   ) where
@@ -18,12 +22,12 @@ module PLFA.Build.Metadata
 import Data.Aeson (encode)
 import Data.Aeson.Types
 import Data.ByteString qualified as B
-import Data.Frontmatter (parseYamlFrontmatterEither)
+import Data.Frontmatter as Frontmatter (Result, IResult (..), parseYamlFrontmatter)
 import Data.Function ((&))
 import Data.HashMap.Strict qualified as H
 import Data.Text qualified as T
 import Data.Time.Format (formatTime, defaultTimeLocale)
-import Data.Vector as V
+import Data.Vector ()
 import Data.Yaml qualified as Y
 import System.Directory
 import System.IO.Unsafe (unsafePerformIO)
@@ -72,11 +76,25 @@ readYamlFrontmatter' inputFile = do
 
 -- | Read the YAML frontmatter from a file.
 readYamlFrontmatter :: FromJSON a => FilePath -> IO a
-readYamlFrontmatter inputFile = do
-  contents <- B.readFile inputFile
-  liftEither (\msg -> "Parse error in " <> inputFile <> ": " <> msg) $
-    parseYamlFrontmatterEither contents
+readYamlFrontmatter inputFile =
+  snd <$> readFileWithMetadata inputFile
 
+-- | Read a file with its YAML frontmatter and return both as a Shake action.
+readFileWithMetadata' :: FromJSON a => FilePath -> Action (Text, a)
+readFileWithMetadata' inputFile = do
+  need [inputFile]
+  liftIO $ readFileWithMetadata inputFile
+
+-- | Read a file with its YAML frontmatter and return both.
+readFileWithMetadata :: FromJSON a => FilePath -> IO (Text, a)
+readFileWithMetadata inputFile = do
+  contents <- B.readFile inputFile
+  withResult (parseYamlFrontmatter contents)
+  where
+    withResult :: (FromJSON a) => Frontmatter.Result a -> IO (Text, a)
[…]

### 2021-09-09 lib/PLFA/Build/Prelude/Url.hs

@@ -19,17 +19,16 @@ import PLFA.Build.Prelude.FilePath
 
 type Url = Text
 
+
 -- | Make a Url relative to the site's root directory.
 --
 --   Adapted from hakyll's 'Hakyll.Web.Html.RelativizeUrls.relativizeUrls'
 relativizeUrl :: Url -> Url
 relativizeUrl url
-  | isRelative url = let (path, _anchor) = T.breakOn "#" url in toRoot path <> url
-  | otherwise      = url
+  | "/" `T.isPrefixOf` url && not ("//" `T.isPrefixOf` url) =
+    let (path, _anchor) = T.breakOn "#" url in toRoot path <> url
+  | otherwise = url
   where
-    isRelative :: Url -> Bool
-    isRelative url = "/" `T.isPrefixOf` url && not ("//" `T.isPrefixOf` url)
-
     -- | Creates the relative URL path to the site root for a given absolute URL path.
     toRoot :: Url -> Url
     toRoot = T.pack .

### 2021-09-09 lib/PLFA/Build/Route.hs

@@ -1,6 +1,8 @@
 module PLFA.Build.Route
-  ( permalinkRoute
-  , getPermalinkRoutingTable
+  ( RoutingTable
+  , buildPermalinkRoutingTable
+  , asRoutingTable
+  , route
   ) where
 
 import Control.Monad (forM)
@@ -11,17 +13,33 @@ import Data.Text qualified as T
 import PLFA.Build.Metadata
 import PLFA.Build.Prelude
 
-permalinkRoute :: FilePath -> FilePath -> IO (Maybe FilePath)
-permalinkRoute targetDir src = do
-  metadata <- readYamlFrontmatter src
-  forM (metadata ^. "permalink") $ \permalink -> do
-    let safePermalink = T.dropAround (=='/') permalink
-    return $ targetDir </> T.unpack safePermalink </> "index.html"
+addIndexHtml :: Url -> Url
+addIndexHtml url
+  | "/" `T.isSuffixOf` url = url <> "index.html"
+  | otherwise              = url
 
-getPermalinkRoutingTable :: FilePath -> [FilePath] -> IO (Map FilePath FilePath)
-getPermalinkRoutingTable targetDir sources = do
+
+type RoutingTable = Map FilePath FilePath
+
+buildPermalinkRoutingTable :: FilePath -> [FilePath] -> IO RoutingTable
+buildPermalinkRoutingTable targetDir sources = do
   permalinkRoutes <-
     forM sources $ \src -> do
       maybeOut <- permalinkRoute targetDir src
       return $ fmap (\out -> (src, normalise out)) maybeOut
   return $ M.fromList (catMaybes permalinkRoutes)
+
+permalinkRoute :: FilePath -> FilePath -> IO (Maybe FilePath)
+permalinkRoute targetDir src = do
+  metadata <- readYamlFrontmatter src
+  forM (metadata ^. "permalink") $ \permalink -> do
+    let permalink' = T.unpack $ T.dropWhile (=='/') $ addIndexHtml $ permalink
+    return $ targetDir </> permalink'
+
+asRoutingTable :: (FilePath -> FilePath) -> [FilePath] -> RoutingTable
+asRoutingTable f sources = M.fromList [ (src, f src) | src <- sources ]
+
+route :: RoutingTable -> FilePath -> FilePath
+route table src = case M.lookup src table of
+  Nothing  -> error $ "Could not find route for " <> src
+  Just out -> out

### 2021-09-09 lib/Builder.hs
Fixed acknowledgements.

@@ -13,6 +13,7 @@ import PLFA.Build.Gather
 import PLFA.Build.Metadata
 import PLFA.Build.Pandoc as Pandoc
 import PLFA.Build.Prelude
+import PLFA.Build.Preprocess
 import PLFA.Build.Route
 import PLFA.Build.Style.CSS
 import PLFA.Build.Style.Sass
@@ -22,22 +23,21 @@ import PLFA.Build.Style.Sass
 -- Directories
 --------------------------------------------------------------------------------
 
-authorDir         = rootDir </> "authors"
-cacheDir          = rootDir </> "_build"
-contributorDir    = rootDir </> "contributors"
-courseDir         = rootDir </> "courses"
-cssDir            = rootDir </> "css"
+authorDir         = "authors"
+cacheDir          = "_build"
+contributorDir    = "contributors"
+courseDir         = "courses"
+cssDir            = "css"
 epubCacheDir      = cacheDir </> "epub"
 htmlCacheDir      = cacheDir </> "html"
-postsDir          = rootDir </> "posts"
-publicDir         = rootDir </> "public"
-rootDir           = "."
-siteDir           = rootDir </> "_site"
-stdlibDir         = rootDir </> "standard-library"
+postsDir          = "posts"
+publicDir         = "public"
+siteDir           = "_site"
+stdlibDir         = "standard-library"
 stdlibAgdaLibFile = stdlibDir </> "standard-library.agda-lib"
-srcDir            = rootDir </> "src"
+srcDir            = "src"
 texCacheDir       = cacheDir </> "tex"
-templateDir       = rootDir </> "templates"
+templateDir       = "templates"
 
 main :: IO ()
 main = do
@@ -46,17 +46,20 @@ main = do
   -- Gather input files and build routing table
   --------------------------------------------------------------------------------
 
+  -- Gather files
   staticFiles           <- gather always regularFile [publicDir]
   agdaLibs              <- gatherAgdaLibs [courseDir, srcDir]
   stdlib                <- readStandardLibraryAgdaLib stdlibDir
-  mdFiles               <- gatherMdFiles [courseDir, srcDir]
+  mdFiles               <- gatherMdFiles [courseDir, postsDir, srcDir]
   scssFiles             <- gatherSassFiles [cssDir]
   postMdFiles           <- gatherMdFiles [postsDir]
[…]

### 2021-09-09 lib/PLFA/Build/Metadata.hs

@@ -186,7 +186,7 @@ sourceFile inputFile key =
 -- | Create a metadata object containing the date inferred from the file path.
 dateFromFilePath :: FilePath -> Text -> Metadata
 dateFromFilePath inputFile key =
-  case T.splitOn "-" $ T.pack inputFile of
+  case T.splitOn "-" $ T.pack $ takeFileName inputFile of
     (year : month : day : _) -> mempty & key .~ T.concat [year, "-", month, "-", day]
     _                        -> mempty
 

### 2021-09-09 lib/PLFA/Build/Preprocess.hs

@@ -1,0 +1,29 @@
+module PLFA.Build.Preprocess
+  ( preprocessForHtml
+  , preprocessForLaTeX
+  ) where
+
+import Data.Text.ICU qualified as ICU
+import Data.Text.ICU.Replace qualified as ICU
+import PLFA.Build.Prelude
+
+--------------------------------------------------------------------------------
+-- Preprocessing literate Agda files
+--------------------------------------------------------------------------------
+
+-- | Wrap literate Agda blocks in a raw HTML block.
+preprocessForHtml :: Text -> Text
+preprocessForHtml = ICU.replaceAll reCodeBlock "\n\n~~~{=html}\n```agda$1```\n\n~~~\n\n"
+
+-- | Wrap literate Agda blocks in a raw LaTeX block.
+preprocessForLaTeX :: Text -> Text
+preprocessForLaTeX = ICU.replaceAll reCodeBlock "\n\n~~~{=latex}\n\\begin{code}$1\\end{code}\n\n~~~\n\n"
+
+-- | Match a literate Agda block.
+--
+--   NOTE: This function assumes literate Agda is written using backtick code
+--         blocks, as Agda does not supported tilde code blocks, and that the
+--         code blocks are /tagged as Agda code/.
+--
+reCodeBlock :: ICU.Regex
+reCodeBlock = ICU.regex [ICU.DotAll, ICU.Multiline] "^```\\s*agda\\s*$(.*?)^```\\s*$"

### 2021-09-09 lib/PLFA/Build/Route.hs

@@ -1,8 +1,11 @@
+{-# LANGUAGE TupleSections #-}
+
 module PLFA.Build.Route
   ( RoutingTable
   , buildPermalinkRoutingTable
   , asRoutingTable
-  , route
+  , routeUrl
+  , routeFile
   ) where
 
 import Control.Monad (forM)
@@ -19,27 +22,30 @@ addIndexHtml url
   | otherwise              = url
 
 
-type RoutingTable = Map FilePath FilePath
+type RoutingTable = Map FilePath Url
 
-buildPermalinkRoutingTable :: FilePath -> [FilePath] -> IO RoutingTable
-buildPermalinkRoutingTable targetDir sources = do
+buildPermalinkRoutingTable :: [FilePath] -> IO RoutingTable
+buildPermalinkRoutingTable sources = do
   permalinkRoutes <-
     forM sources $ \src -> do
-      maybeOut <- permalinkRoute targetDir src
-      return $ fmap (\out -> (src, normalise out)) maybeOut
+      maybeOut <- permalinkRoute src
+      return $ fmap (src,) maybeOut
   return $ M.fromList (catMaybes permalinkRoutes)
 
-permalinkRoute :: FilePath -> FilePath -> IO (Maybe FilePath)
-permalinkRoute targetDir src = do
+permalinkRoute :: FilePath -> IO (Maybe Url)
+permalinkRoute src = do
   metadata <- readYamlFrontmatter src
   forM (metadata ^. "permalink") $ \permalink -> do
-    let permalink' = T.unpack $ T.dropWhile (=='/') $ addIndexHtml $ permalink
-    return $ targetDir </> permalink'
+    return $ addIndexHtml permalink
 
 asRoutingTable :: (FilePath -> FilePath) -> [FilePath] -> RoutingTable
-asRoutingTable f sources = M.fromList [ (src, f src) | src <- sources ]
+asRoutingTable f sources = M.fromList [ (src, T.pack $ f src) | src <- sources ]
 
-route :: RoutingTable -> FilePath -> FilePath
-route table src = case M.lookup src table of
+routeUrl :: RoutingTable -> FilePath -> Url
+routeUrl table src = case M.lookup src table of
   Nothing  -> error $ "Could not find route for " <> src
   Just out -> out
+
+routeFile :: FilePath -> RoutingTable -> FilePath -> FilePath
+routeFile targetDir table src =
+  targetDir </> (T.unpack $ T.dropWhile (=='/') $ routeUrl table src)

### 2021-09-09 lib/Builder.hs
Fixing last few bugs in generated site.

@@ -6,15 +6,14 @@ import Data.Map (Map)
 import Data.Map qualified as M
 import Data.Maybe (fromMaybe)
 import Data.Text qualified as T
-import Data.Text.ICU qualified as ICU
-import Data.Text.ICU.Replace qualified as ICU
 import PLFA.Build.Agda
 import PLFA.Build.Gather
 import PLFA.Build.Metadata
 import PLFA.Build.Pandoc as Pandoc
 import PLFA.Build.Prelude
 import PLFA.Build.Preprocess
-import PLFA.Build.Route
+import PLFA.Build.Route hiding (routeFile, routeUrl)
+import PLFA.Build.Route qualified as Route (routeFile, routeUrl)
 import PLFA.Build.Style.CSS
 import PLFA.Build.Style.Sass
 
@@ -48,9 +47,10 @@ main = do
 
   -- Gather files
   staticFiles           <- gather always regularFile [publicDir]
+  coursePdfs            <- gather always (regularFile &&? extension ==? ".pdf") [courseDir]
   agdaLibs              <- gatherAgdaLibs [courseDir, srcDir]
   stdlib                <- readStandardLibraryAgdaLib stdlibDir
-  mdFiles               <- gatherMdFiles [courseDir, postsDir, srcDir]
+  mdFiles               <- ("README.md" :) <$> gatherMdFiles [courseDir, postsDir, srcDir]
   scssFiles             <- gatherSassFiles [cssDir]
   postMdFiles           <- gatherMdFiles [postsDir]
   lagdaMdFilesByAgdaLib <- gatherLagdaMdFilesForAgdaLibs agdaLibs
@@ -58,10 +58,15 @@ main = do
   -- Build routing table
   permalinkRoutingTable <- buildPermalinkRoutingTable mdFiles
   let postsRoutingTable = asRoutingTable (-<.> "html") postMdFiles
-  let errorRoutingTable = asRoutingTable id ["404.html"]
-  let routingTable = mconcat [permalinkRoutingTable, postsRoutingTable, errorRoutingTable]
+  let staticRoutingTable = asRoutingTable id (staticFiles <> coursePdfs)
+  let routingTable = mconcat [permalinkRoutingTable, postsRoutingTable, staticRoutingTable]
+  let routeUrl  src = Route.routeUrl routingTable src
+  let routeFile src = Route.routeFile siteDir routingTable src
 
   shakeArgs shakeOpts $ do
+
+    -- Require Shake builds all targets in the routing table
[…]

### 2021-09-09 lib/PLFA/Build/Prelude/Url.hs

@@ -23,10 +23,9 @@ type Url = Text
 -- | Make a Url relative to the site's root directory.
 --
 --   Adapted from hakyll's 'Hakyll.Web.Html.RelativizeUrls.relativizeUrls'
-relativizeUrl :: Url -> Url
-relativizeUrl url
-  | "/" `T.isPrefixOf` url && not ("//" `T.isPrefixOf` url) =
-    let (path, _anchor) = T.breakOn "#" url in toRoot path <> url
+relativizeUrl :: Url -> Url -> Url
+relativizeUrl pageUrl url
+  | "/" `T.isPrefixOf` url && not ("//" `T.isPrefixOf` url) = toRoot pageUrl <> url
   | otherwise = url
   where
     -- | Creates the relative URL path to the site root for a given absolute URL path.
@@ -41,6 +40,7 @@ relativizeUrl url
 
 
 
+
 -- | Apply a function to each Url in a raw HTML document.
 --
 --   Adapted from hakyll's 'Hakyll.Web.Html.withUrls'
@@ -50,7 +50,7 @@ withUrls f = TS.renderTags . map tag . TS.parseTags
     tag (TS.TagOpen s a) = TS.TagOpen s $ map attr a
     tag x                = x
     attr (k, v)          = (k, if k `elem` refs then f v else v)
-    refs                 = ["src", "href"]
+    refs                 = ["src", "href", "xlink:href"]
 
 -- | Apply a function to each Url in a Pandoc document.
 withUrlsPandoc :: (Url -> Url) -> Pandoc -> Pandoc

### 2021-09-09 lib/PLFA/Build/Preprocess.hs

@@ -1,17 +1,37 @@
 module PLFA.Build.Preprocess
   ( preprocessForHtml
   , preprocessForLaTeX
+  , postprocessHtml5
   ) where
 
+import Data.String (IsString (fromString))
 import Data.Text.ICU qualified as ICU
 import Data.Text.ICU.Replace qualified as ICU
 import PLFA.Build.Prelude
+import Text.Printf (printf)
 
 --------------------------------------------------------------------------------
 -- Preprocessing literate Agda files
 --------------------------------------------------------------------------------
 
--- | Wrap literate Agda blocks in a raw HTML block.
+-- | Removes closing tags for '<img>' and '<input>' tags.
+postprocessHtml5 :: Text -> Text
+postprocessHtml5 = ICU.replaceAll reSelfClosing "/>"
+  where
+    reSelfClosing :: ICU.Regex
+    reSelfClosing =
+      fromString $ foldr1 (>|<) $ map (printf "(></%s>)") selfClosingTags
+
+    (>|<) :: String -> String -> String
+    s1 >|< s2 = s1 <> "|" <> s2
+
+    selfClosingTags :: [String]
+    selfClosingTags =
+      ["area", "base", "br", "col", "embed", "hr", "img", "input", "link",
+       "meta", "param", "source", "track", "wbr", "command", "keygen", "menuitem"]
+
+
+-- | Wrap literate Agda blocks in a raw Html block.
 preprocessForHtml :: Text -> Text
 preprocessForHtml = ICU.replaceAll reCodeBlock "\n\n~~~{=html}\n```agda$1```\n\n~~~\n\n"
 

### 2021-09-09 lib/PLFA/Build/Route.hs

@@ -6,6 +6,7 @@ module PLFA.Build.Route
   , asRoutingTable
   , routeUrl
   , routeFile
+  , wantAllTargets
   ) where
 
 import Control.Monad (forM)
@@ -48,4 +49,12 @@ routeUrl table src = case M.lookup src table of
 
 routeFile :: FilePath -> RoutingTable -> FilePath -> FilePath
 routeFile targetDir table src =
-  targetDir </> (T.unpack $ T.dropWhile (=='/') $ routeUrl table src)
+  urlToFile targetDir (routeUrl table src)
+
+wantAllTargets :: FilePath -> RoutingTable -> Rules ()
+wantAllTargets targetDir table =
+  want [ urlToFile targetDir url | url <- M.elems table ]
+
+urlToFile :: FilePath -> Url -> FilePath
+urlToFile targetDir url =
+  targetDir </> (T.unpack $ T.dropWhile (=='/') $ url)

### 2021-09-09 lib/Builder.hs
Fixed EPUB creation, sketched out PDF creation, there's lots of code duplication.

@@ -6,6 +6,7 @@ import Data.Map (Map)
 import Data.Map qualified as M
 import Data.Maybe (fromMaybe)
 import Data.Text qualified as T
+import Debug.Trace (traceShow)
 import PLFA.Build.Agda
 import PLFA.Build.Gather
 import PLFA.Build.Metadata
@@ -23,20 +24,29 @@ import PLFA.Build.Style.Sass
 --------------------------------------------------------------------------------
 
 authorDir         = "authors"
+bookDir           = "book"
 cacheDir          = "_build"
 contributorDir    = "contributors"
 courseDir         = "courses"
 cssDir            = "css"
 epubCacheDir      = cacheDir </> "epub"
 htmlCacheDir      = cacheDir </> "html"
+htmlStage1Dir     = htmlCacheDir </> "stage1"
+htmlStage2Dir     = htmlCacheDir </> "stage2"
 postsDir          = "posts"
 publicDir         = "public"
 siteDir           = "_site"
 stdlibDir         = "standard-library"
-stdlibAgdaLibFile = stdlibDir </> "standard-library.agda-lib"
 srcDir            = "src"
 texCacheDir       = cacheDir </> "tex"
 templateDir       = "templates"
+fontsDir          = bookDir </> "fonts"
+webfontsDir       = publicDir </> "webfonts"
+
+bibtexFile        = "bib" </> "plfa.bib"
+stdlibAgdaLibFile = stdlibDir </> "standard-library.agda-lib"
+
+badgeUrl          = ""
 
 main :: IO ()
 main = do
@@ -46,14 +56,17 @@ main = do
   --------------------------------------------------------------------------------
 
   -- Gather files
-  staticFiles           <- gather always regularFile [publicDir]
+  publicFiles           <- gather always regularFile [publicDir]
   coursePdfs            <- gather always (regularFile &&? extension ==? ".pdf") [courseDir]
+  let staticFiles        = publicFiles <> coursePdfs
   agdaLibs              <- gatherAgdaLibs [courseDir, srcDir]
   stdlib                <- readStandardLibraryAgdaLib stdlibDir
-  mdFiles               <- ("README.md" :) <$> gatherMdFiles [courseDir, postsDir, srcDir]
+  mdFiles               <- ("README.md" :) <$> gatherMdFiles [postsDir, srcDir]
   scssFiles             <- gatherSassFiles [cssDir]
[…]

### 2021-09-09 lib/PLFA/Build/Agda.hs

@@ -18,6 +18,7 @@ module PLFA.Build.Agda
   , highlightAgdaWith
   , readStandardLibraryAgdaLib
   , prepareFixLinks
+  , mapAgdaLib
   ) where
 
 import Agda.Main
@@ -324,9 +325,6 @@ readAgdaLib inputFile = do
       _libIncludes = map normalise (_libIncludes agdaLibFile)
     } Nothing
 
-instance Eq AgdaLib where
-  lib1 == lib2 = libName lib1 == libName lib2
-
 instance ToJSON AgdaLib where
   toJSON agdaLib =
     object [ "name"    .= libName agdaLib
@@ -336,6 +334,15 @@ instance ToJSON AgdaLib where
            , "pragmas" .= libPragmas agdaLib
            , "url"     .= libUrl agdaLib
            ]
+
+
+mapAgdaLib :: (FilePath -> FilePath) -> AgdaLib -> AgdaLib
+mapAgdaLib f agdaLib =
+  if libName agdaLib == "standard-library" then
+    agdaLib
+  else
+    agdaLib { libFile     = f (libFile agdaLib),
+              libIncludes = map f (libIncludes agdaLib) }
 
 --------------------------------------------------------------------------------
 -- Temporary directories

### 2021-09-09 lib/PLFA/Build/Metadata.hs

@@ -2,6 +2,8 @@ module PLFA.Build.Metadata
   ( Metadata (..),
     readYaml',
     readYaml,
+    writeYaml',
+    writeYaml,
     readYamlFrontmatter',
     readYamlFrontmatter,
     readFileWithMetadata',
@@ -11,10 +13,11 @@ module PLFA.Build.Metadata
     (&),
     Author (..),
     Contributor (..),
-    lastModified,
-    dateFromFilePath,
-    sourceFile,
-    teaser,
+    lastModifiedField,
+    dateFromFileNameField,
+    currentDateField,
+    constField,
+    teaserField,
     addTitleVariants,
     resolveIncludes
   ) where
@@ -26,7 +29,8 @@ import Data.Frontmatter as Frontmatter (Result, IResult (..), parseYamlFrontmatt
 import Data.Function ((&))
 import Data.HashMap.Strict qualified as H
 import Data.Text qualified as T
-import Data.Time.Format (formatTime, defaultTimeLocale)
+import Data.Time
+import Data.Time.Format.ISO8601 (iso8601Show)
 import Data.Vector ()
 import Data.Yaml qualified as Y
 import System.Directory
@@ -57,6 +61,10 @@ metadata@(Metadata obj) ^. key = do
   a <- liftResult $ fromJSON v
   return a
 
+
+--------------------------------------------------------------------------------
+-- Reading and writing
+--------------------------------------------------------------------------------
 
 -- | Read a YAML file as a Shake action.
 readYaml' :: FromJSON a => FilePath -> Action a
@@ -95,6 +103,14 @@ readFileWithMetadata inputFile = do
     withResult (Done body metadata) = do tbody <- toText body; return (tbody, metadata)
     withResult (Fail _ _ message)   = fail message
     withResult (Partial k)          = withResult (k "")
+
+
+writeYaml' :: ToJSON a => FilePath -> a -> Action ()
+writeYaml' outputFile a = liftIO $ Y.encodeFile outputFile a
+
+writeYaml :: ToJSON a => FilePath -> a -> IO ()
+writeYaml outputFile a = Y.encodeFile outputFile a
+
 
 -- * Instances
 
@@ -167,38 +183,45 @@ instance FromJSON Contributor where
 -- Metadata fields
 --------------------------------------------------------------------------------
 
+prettyDate :: FormatTime t => t -> String
+prettyDate time = formatTime defaultTimeLocale "%a %-d %b, %Y" time
+
[…]

### 2021-09-09 lib/PLFA/Build/Pandoc.hs

@@ -5,14 +5,15 @@ module PLFA.Build.Pandoc
     Template,
     runPandocIO,
     compileTemplate,
-    renderTemplate
+    renderTemplate,
   ) where
 
 import Control.Exception (displayException)
-import Data.Aeson.Types (Value (..))
+import Data.Aeson.Types
 import PLFA.Build.Prelude
 import PLFA.Build.Metadata
 import Text.DocLayout qualified as Doc (render)
+import Text.Pandoc.App as Export
 import Text.Pandoc as Export hiding (Format, Template, getTemplate, compileTemplate, renderTemplate)
 import Text.Pandoc.Writers.HTML as Export (writeHtmlStringForEPUB)
 import Text.Pandoc.Templates qualified as Template

### 2021-09-09 lib/PLFA/Build/Prelude/Url.hs

@@ -1,12 +1,12 @@
 {-# LANGUAGE OverloadedStrings #-}
 
 module PLFA.Build.Prelude.Url
-  ( Url
-  , relativizeUrl
-  , withUrls
-  , withUrlsPandoc
-  , withUrlsBlockPandoc
-  , withUrlsInlinePandoc
+  ( Url,
+    relativizeUrl,
+    withUrls,
+    withUrlsPandoc,
+    withUrlsBlockPandoc,
+    withUrlsInlinePandoc
   ) where
 
 import Data.Function ((&))
@@ -37,8 +37,6 @@ relativizeUrl pageUrl url
              splitDirectories .
              takeDirectory .
              T.unpack
-
-
 
 
 -- | Apply a function to each Url in a raw HTML document.

### 2021-09-09 lib/PLFA/Build/Preprocess.hs

@@ -2,17 +2,23 @@ module PLFA.Build.Preprocess
   ( preprocessForHtml
   , preprocessForLaTeX
   , postprocessHtml5
+  , sanitizeForEpub
+  , shiftHeaders
   ) where
 
 import Data.String (IsString (fromString))
 import Data.Text.ICU qualified as ICU
 import Data.Text.ICU.Replace qualified as ICU
 import PLFA.Build.Prelude
+import Text.HTML.TagSoup qualified as TS
 import Text.Printf (printf)
 
---------------------------------------------------------------------------------
--- Preprocessing literate Agda files
---------------------------------------------------------------------------------
+
+-- | Shift all Markdown headers by one.
+shiftHeaders :: Text -> Text
+shiftHeaders body = ICU.replaceAll reHeader "#$1" body
+  where
+    reHeader = ICU.regex [ICU.Multiline] "^(#+)"
 
 -- | Removes closing tags for '<img>' and '<input>' tags.
 postprocessHtml5 :: Text -> Text
@@ -47,3 +53,23 @@ preprocessForLaTeX = ICU.replaceAll reCodeBlock "\n\n~~~{=latex}\n\\begin{code}$
 --
 reCodeBlock :: ICU.Regex
 reCodeBlock = ICU.regex [ICU.DotAll, ICU.Multiline] "^```\\s*agda\\s*$(.*?)^```\\s*$"
+
+
+sanitizeForEpub :: Text -> Text
+sanitizeForEpub = ICU.replaceAll reRawBlock (ICU.rtfn replace)
+  where
+    replace :: ICU.Match -> Text
+    replace m = case ICU.group 1 m of
+      Nothing -> error "You got it wrong you silly"
+      Just cb -> "\n\n~~~{=html}\n" <> sanitizeRawBlock cb <> "\n\n~~~\n\n"
+
+-- | Remove everything except for classes from Html.
+sanitizeRawBlock :: Text -> Text
+sanitizeRawBlock = TS.renderTags . map tag . TS.parseTags
+  where
+    tag (TS.TagOpen s a) = TS.TagOpen s $ filter (\(k,_) -> k == "class") a
+    tag x                = x
+
+-- | Match a raw Html block.
+reRawBlock :: ICU.Regex
+reRawBlock = ICU.regex [ICU.DotAll, ICU.Multiline] "^~~~\\{=html\\}\\s*$(.*?)^~~~\\s*$"

### 2021-12-12 lib/Builder.hs
Adding some more changes which have not had the time to review (hence dirty) (#624)

@@ -15,6 +15,7 @@ import PLFA.Build.Prelude
 import PLFA.Build.Preprocess
 import PLFA.Build.Route hiding (routeFile, routeUrl)
 import PLFA.Build.Route qualified as Route (routeFile, routeUrl)
+import PLFA.Build.Staged
 import PLFA.Build.Style.CSS
 import PLFA.Build.Style.Sass
 
@@ -29,16 +30,11 @@ cacheDir          = "_build"
 contributorDir    = "contributors"
 courseDir         = "courses"
 cssDir            = "css"
-epubCacheDir      = cacheDir </> "epub"
-htmlCacheDir      = cacheDir </> "html"
-htmlStage1Dir     = htmlCacheDir </> "stage1"
-htmlStage2Dir     = htmlCacheDir </> "stage2"
 postsDir          = "posts"
 publicDir         = "public"
 siteDir           = "_site"
 stdlibDir         = "standard-library"
 srcDir            = "src"
-texCacheDir       = cacheDir </> "tex"
 templateDir       = "templates"
 fontsDir          = bookDir </> "fonts"
 webfontsDir       = publicDir </> "webfonts"
@@ -62,17 +58,17 @@ main = do
   mdFiles               <- ("README.md" :) <$> gatherMdFiles [postsDir, srcDir]
   scssFiles             <- gatherSassFiles [cssDir]
   postMdFiles           <- gatherMdFiles [postsDir]
-  lagdaMdFilesByAgdaLib <- gatherLagdaMdFilesForAgdaLibs agdaLibs
+  mdFileToAgdaLib       <- buildAgdaLibMap agdaLibs
   webfontFiles          <- gather always (regularFile &&? extension ==? ".woff") [webfontsDir]
   fontFiles             <- gather always (regularFile &&? extension ==? ".ttf") [fontsDir]
 
   -- Build routing table
   permalinkRoutingTable <- buildPermalinkRoutingTable mdFiles
-  let postsRoutingTable = asRoutingTable (-<.> "html") postMdFiles
+  let postsRoutingTable  = asRoutingTable (-<.> "html") postMdFiles
   let staticRoutingTable = asRoutingTable id (staticFiles <> coursePdfs)
-  let routingTable = mconcat [permalinkRoutingTable, postsRoutingTable, staticRoutingTable]
-  let routeUrl  src = Route.routeUrl routingTable src
-  let routeFile src = Route.routeFile siteDir routingTable src
[…]

### 2022-06-07 tools/Buildfile/Author.hs
Fix build problem

@@ -14,6 +14,7 @@ data Author = Author
   , authorTwitter       :: Maybe Text
   , authorCorresponding :: Bool
   }
+  deriving (Show)
 
 instance ToJSON Author where
   toJSON Author{..} =

### 2022-06-07 tools/Buildfile/Contributor.hs

@@ -13,6 +13,7 @@ data Contributor = Contributor
   , contributorGithub :: Text
   , contributorCount  :: Int
   }
+  deriving (Show)
 
 instance ToJSON Contributor where
   toJSON Contributor{..} =

### 2022-06-07 tools/Buildfile.hs
More caching

@@ -172,13 +172,11 @@ main = do
       --------------------------------------------------------------------------------
       -- Caches
 
-      -- getAuthors <- newCache $ \() -> do
       let getAuthors = \() -> do
             authorFiles <- getDirectoryFiles authorDir ["*.yml"]
             authors <- traverse (\src -> readYaml $ authorDir </> src) authorFiles
             return (authors :: [Author])
 
-      -- getContributors <- newCache $ \() -> do
       let getContributors = \() -> do
             contributorFiles <- getDirectoryFiles "" [contributorDir </> "*.yml"]
             contributors <- traverse readYaml contributorFiles
@@ -186,19 +184,17 @@ main = do
             return (sortedContributors :: [Contributor])
 
       getDefaultMetadata <- newCache $ \() -> do
-      -- let getDefaultMetadata = \() -> do
-            metadata <- readYaml @Metadata (dataDir </> "metadata.yml")
-            authorMetadata <- getAuthors ()
-            return $ mconcat [ metadata, constField "author" authorMetadata ]
+        metadata <- readYaml @Metadata (dataDir </> "metadata.yml")
+        authorMetadata <- getAuthors ()
+        return $ mconcat [ metadata, constField "author" authorMetadata ]
       let ?getDefaultMetadata = getDefaultMetadata
 
       getReferences <- newCache $ \() -> do
-      -- let getReferences = \() -> do
-            bibliographyFileBody <- readFile' bibliographyFile
-            bibliographyFileDoc@(Pandoc meta _) <- runPandoc $ Pandoc.readBibTeX readerOpts bibliographyFileBody
-            case Pandoc.lookupMeta "references" meta of
-              Just references -> return references
-              Nothing -> fail $ printf "Could not read references from '%s'" bibliographyFile
+        bibliographyFileBody <- readFile' bibliographyFile
+        bibliographyFileDoc@(Pandoc meta _) <- runPandoc $ Pandoc.readBibTeX readerOpts bibliographyFileBody
+        case Pandoc.lookupMeta "references" meta of
+          Just references -> return references
+          Nothing -> fail $ printf "Could not read references from '%s'" bibliographyFile
 
       let processCitations :: Pandoc -> Action Pandoc
           processCitations doc = do
@@ -206,10 +202,9 @@ main = do
             Pandoc.processCitations $
               Pandoc.setMeta "references" references doc
 
[…]

### 2022-06-07 tools/Buildfile.hs
Added print statements

@@ -237,6 +237,7 @@ main = do
       -- Build /
       outDir </> "index.html" %> \out -> do
         src <- routeSource out
+        putInfo $ printf "Compile '%s' to '%s'" src out
         tocField <- getTableOfContentsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
         return indexMarkdownTemplate
@@ -254,6 +255,7 @@ main = do
       -- Stage 1: Compile Agda to HTML
       tmpAgdaHtmlDir <//> "*.md" %> \next -> do
         (src, prev) <- (,) <$> routeSource next <*> routePrev next
+        putInfo $ printf "Compile '%s' to '%s'" prev next
         agdaLibraries <-
           failOnError $
             getAgdaLibrariesForProject <$> getProject src
@@ -266,6 +268,7 @@ main = do
       -- Stage 2: Compile Markdown to HTML
       tmpBodyHtmlDir <//> "*.html" %> \next -> do
         (src, prev, out) <- (,,) <$> routeSource next <*> routePrev next <*> route next
+        putInfo $ printf "Compile '%s' to '%s'" prev next
         maybeAgdaLinkFixer <-
           getProject src
             & rightToMaybe
@@ -281,6 +284,7 @@ main = do
       -- Stage 3: Apply HTML templates
       outDir <//> "*.html" %> \out -> do
         (src, prev) <- (,) <$> routeSource out <*> routePrev out
+        putInfo $ printf "Compile '%s' to '%s'" prev out
         (metadata, htmlBody) <- getFileWithMetadata prev
         let htmlTemplates
               | isPostSource src = ["post.html", "default.html"]
@@ -312,6 +316,7 @@ main = do
       -- Build /Announcements/index.html
       outDir </> "Announcements" </> "index.html" %> \out -> do
         src <- routeSource out
+        putInfo $ printf "Compile '%s' to '%s'" src out
         postsField <- getPostsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
         return indexMarkdownTemplate
@@ -326,8 +331,8 @@ main = do
       -- Build /GettingStarted/index.html
       outDir </> "GettingStarted" </> "index.html" %> \out -> do
         src <- routeSource out
+        putInfo $ printf "Compile '%s' to '%s'" src out
         (fileMetadata, readmeMarkdown) <- getFileWithMetadata src
-
         let metadata =
               mconcat
[…]

### 2022-06-07 tools/Buildfile.hs
Removed puts

@@ -237,7 +237,7 @@ main = do
       -- Build /
       outDir </> "index.html" %> \out -> do
         src <- routeSource out
-        putInfo $ printf "Compile '%s' to '%s'" src out
+        -- putInfo $ printf "Compile '%s' to '%s'" src out
         tocField <- getTableOfContentsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
         return indexMarkdownTemplate
@@ -255,7 +255,7 @@ main = do
       -- Stage 1: Compile Agda to HTML
       tmpAgdaHtmlDir <//> "*.md" %> \next -> do
         (src, prev) <- (,) <$> routeSource next <*> routePrev next
-        putInfo $ printf "Compile '%s' to '%s'" prev next
+        -- putInfo $ printf "Compile '%s' to '%s'" prev next
         agdaLibraries <-
           failOnError $
             getAgdaLibrariesForProject <$> getProject src
@@ -268,7 +268,7 @@ main = do
       -- Stage 2: Compile Markdown to HTML
       tmpBodyHtmlDir <//> "*.html" %> \next -> do
         (src, prev, out) <- (,,) <$> routeSource next <*> routePrev next <*> route next
-        putInfo $ printf "Compile '%s' to '%s'" prev next
+        -- putInfo $ printf "Compile '%s' to '%s'" prev next
         maybeAgdaLinkFixer <-
           getProject src
             & rightToMaybe
@@ -284,7 +284,7 @@ main = do
       -- Stage 3: Apply HTML templates
       outDir <//> "*.html" %> \out -> do
         (src, prev) <- (,) <$> routeSource out <*> routePrev out
-        putInfo $ printf "Compile '%s' to '%s'" prev out
+        -- putInfo $ printf "Compile '%s' to '%s'" prev out
         (metadata, htmlBody) <- getFileWithMetadata prev
         let htmlTemplates
               | isPostSource src = ["post.html", "default.html"]
@@ -316,7 +316,7 @@ main = do
       -- Build /Announcements/index.html
       outDir </> "Announcements" </> "index.html" %> \out -> do
         src <- routeSource out
-        putInfo $ printf "Compile '%s' to '%s'" src out
+        -- putInfo $ printf "Compile '%s' to '%s'" src out
         postsField <- getPostsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
         return indexMarkdownTemplate
[…]

### 2022-06-07 tools/Buildfile.hs
Minor fixes; put metadata back in README

@@ -145,7 +145,7 @@ main = do
       let ?routingTable =
             mconcat
               [ -- Book (Web Version)
-                ["README.md"] |-> (outDir </> "GettingStarted" </> "index.html"),
+                ["README.md"] |-> postOrPermalinkRouterWithAgda,
                 [webDir </> "TableOfContents.md"] |-> postOrPermalinkRouterWithAgda,
                 [chapterDir </> "plfa" <//> "*.md"] *|-> postOrPermalinkRouterWithAgda,
                 -- Book (EPUB Version)
@@ -186,7 +186,8 @@ main = do
       getDefaultMetadata <- newCache $ \() -> do
         metadata <- readYaml @Metadata (dataDir </> "metadata.yml")
         authorMetadata <- getAuthors ()
-        return $ mconcat [ metadata, constField "author" authorMetadata ]
+        contributorMetadata <- getContributors ()
+        return $ mconcat [ metadata, constField "author" authorMetadata, constField "contributor" contributorMetadata ]
       let ?getDefaultMetadata = getDefaultMetadata
 
       getReferences <- newCache $ \() -> do
@@ -325,25 +326,6 @@ main = do
           >>= processCitations
           >>= Pandoc.pandocToHtml5
           >>= Pandoc.applyTemplates ["page.html", "default.html"] fileMetadata
-          <&> postProcessHtml5 outDir out
-          >>= writeFile' out
-
-      -- Build /GettingStarted/index.html
-      outDir </> "GettingStarted" </> "index.html" %> \out -> do
-        src <- routeSource out
-        -- putInfo $ printf "Compile '%s' to '%s'" src out
-        (fileMetadata, readmeMarkdown) <- getFileWithMetadata src
-        let metadata =
-              mconcat
-                [ fileMetadata,
-                  constField @Text "title" "Getting Started",
-                  constField @Text "prev"  "/Preface/",
-                  constField @Text "next"  "/Naturals/"
-                ]
-        return readmeMarkdown
-          >>= Pandoc.markdownToPandoc
-          >>= Pandoc.pandocToHtml5
-          >>= Pandoc.applyTemplates ["page.html", "default.html"] metadata
           <&> postProcessHtml5 outDir out
           >>= writeFile' out
 
@@ -502,9 +484,8 @@ main = do
         let src = epubTemplateDir </> "epub-metadata.xml"
         -- putInfo $ printf "Compile '%s' to '%s'" src out
         defaultMetadata <- getDefaultMetadata ()
[…]

### 2022-06-07 tools/Buildfile.hs
Minor fixes

@@ -74,13 +74,6 @@ tmpAgdaHtmlDir = tmpDir </> "agda_html" -- Render .lagda.md to .md
 tmpBodyHtmlDir = tmpDir </> "body_html" -- Render .md to .html
 tmpEpubDir = tmpDir </> "epub"
 
-legacyVersions :: [String]
-legacyVersions = ["19.08", "20.07"]
-
--- TODO:
--- - [ ] fix epub
--- - [ ] build pdf
-
 --------------------------------------------------------------------------------
 -- Rules
 

### 2022-06-07 tools/Buildfile.hs
Minor fixes

@@ -43,11 +43,10 @@ outDir, tmpDir :: FilePath
 outDir = "_site"
 tmpDir = "_cache"
 
-dataDir, authorDir, contributorDir, legacyDir, bibliographyFile, tableOfContentsFile :: FilePath
+dataDir, authorDir, contributorDir, bibliographyFile, tableOfContentsFile :: FilePath
 dataDir = "data"
 authorDir = dataDir </> "authors"
 contributorDir = dataDir </> "contributors"
-legacyDir = dataDir </> "legacy"
 tableOfContentsFile = dataDir </> "tableOfContents.yml"
 bibliographyFile = dataDir </> "bibliography.bib"
 

### 2022-06-07 courses/tspl/make_assignment.sh
Remove "layout: page"

@@ -28,7 +28,6 @@ shift
 cat <<-EOF
 ---
 title     : "Assignment$NUM: $COURSE Assignment $NUM"
-layout    : page
 permalink : /$COURSE/$YEAR/Assignment$NUM/
 ---
 

### 2022-06-08 courses/tspl/make_assignment.sh
Fix case

@@ -1,187 +1,0 @@
-#!/usr/bin/env bash
-
-# TODO:
-#   - put the exercises from each chapter in their own module
-#   - replicate the imports from that module
-#   - import that module hiding anything defined in an exercise section
-
-# This script can be used to automatically generate assignment files from PLFA source files.
-# It takes a course abbreviation, e.g. TSPL, a year, and the number of the assignment.
-# At the moment, it outputs the University of Edinburgh guidelines for good scholarly practice,
-# making it somewhat specific to courses run there, but the header should be easy to edit.
-#
-# Usage:
-#
-#   ./make_assignment.sh [COURSE_NAME] [COURSE_YEAR] [ASSIGNMENT_NUMBER] [PLFA_SOURCE_FILE...]
-
-COURSE="$1"
-shift
-
-YEAR="$1"
-shift
-
-NUM="$1"
-shift
-
-## Make assignment header
-
-cat <<-EOF
----
-title     : "Assignment$NUM: $COURSE Assignment $NUM"
-permalink : /$COURSE/$YEAR/Assignment$NUM/
----
-
-\`\`\`
-module Assignment$NUM where
-\`\`\`
-
-## YOUR NAME AND EMAIL GOES HERE
-
-## Introduction
-
-You must do _all_ the exercises labelled "(recommended)".
-
-Exercises labelled "(stretch)" are there to provide an extra challenge.
-You don't need to do all of these, but should attempt at least a few.
-
-Exercises labelled "(practice)" are included for those who want extra practice.
-
-Submit your homework using the "submit" command.
-Please ensure your files execute correctly under Agda!
-
-
-## Good Scholarly Practice.
-
-Please remember the University requirement as
-regards all assessed work. Details about this can be found at:
-
-> [http://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct](http://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct)
-
-Furthermore, you are required to take reasonable measures to protect
-your assessed work from unauthorised access. For example, if you put
-any such work on a public repository then you must set access
-permissions appropriately (generally permitting access only to
-yourself, or your group in the case of group practicals).
-
-EOF
-
-## Awk Scripts
-
[…]

### 2022-06-08 tools/Buildfile.hs
Insert error message

@@ -38,6 +38,8 @@ import Shoggoth.Template.Pandoc.Builder qualified as Builder
 import Shoggoth.Template.Pandoc.Citeproc qualified as Citeproc
 import System.Directory (makeAbsolute)
 import Text.Printf (printf)
+import Control.Exception (catch)
+import System.Exit (exitWith, ExitCode (..))
 
 outDir, tmpDir :: FilePath
 outDir = "_site"
@@ -95,7 +97,14 @@ main = do
       --------------------------------------------------------------------------------
       -- Agda libraries
 
-      standardLibrary <- Agda.getStandardLibrary "standard-library"
+      standardLibrary <- liftIO $
+        catch (Agda.getStandardLibrary "standard-library") $
+            \Agda.AgdaStandardLibraryNotFound {..} -> do
+              putStrLn "Could not find Agda standard library.\n\
+                       \Did you forget to clone the repository with the '--recurse-submodules' flag?\n\
+                       \If so, you can download the Agda standard library into the 'standard-library' directory by running:\n\n\
+                       \  git submodule update --init"
+              exitWith (ExitFailure 1)
 
       let getAgdaLibrariesForProject :: Project -> [Agda.Library]
           getAgdaLibrariesForProject project = standardLibrary : localAgdaLibraries

### 2022-06-08 tools/Buildfile.hs
Suppress Pandoc warnings for EPUB

@@ -1,6 +1,7 @@
 {-# OPTIONS_GHC -Wno-unrecognised-pragmas #-}
 
 {-# HLINT ignore "Monad law, left identity" #-}
+{-# HLINT ignore "Redundant <&>" #-}
 module Main where
 
 import Buildfile.Author (Author (..))
@@ -476,8 +477,7 @@ main = do
                   writerReferenceLocation = Pandoc.EndOfSection
                 }
 
-        epub <- runPandocWith Pandoc.INFO $
-          Pandoc.writeEPUB3 writerOptsForEpub bookDoc
+        epub <- runPandoc $ Pandoc.writeEPUB3 writerOptsForEpub bookDoc
         liftIO $ LazyByteString.writeFile out epub
 
       -- Build epub metadata

### 2022-06-08 tools/Buildfile.hs
Removed puts

@@ -240,7 +240,6 @@ main = do
       -- Build /
       outDir </> "index.html" %> \out -> do
         src <- routeSource out
-        -- putInfo $ printf "Compile '%s' to '%s'" src out
         tocField <- getTableOfContentsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
         return indexMarkdownTemplate
@@ -258,7 +257,6 @@ main = do
       -- Stage 1: Compile Agda to HTML
       tmpAgdaHtmlDir <//> "*.md" %> \next -> do
         (src, prev) <- (,) <$> routeSource next <*> routePrev next
-        -- putInfo $ printf "Compile '%s' to '%s'" prev next
         agdaLibraries <-
           failOnError $
             getAgdaLibrariesForProject <$> getProject src
@@ -271,7 +269,6 @@ main = do
       -- Stage 2: Compile Markdown to HTML
       tmpBodyHtmlDir <//> "*.html" %> \next -> do
         (src, prev, out) <- (,,) <$> routeSource next <*> routePrev next <*> route next
-        -- putInfo $ printf "Compile '%s' to '%s'" prev next
         maybeAgdaLinkFixer <-
           getProject src
             & rightToMaybe
@@ -287,7 +284,6 @@ main = do
       -- Stage 3: Apply HTML templates
       outDir <//> "*.html" %> \out -> do
         (src, prev) <- (,) <$> routeSource out <*> routePrev out
-        -- putInfo $ printf "Compile '%s' to '%s'" prev out
         (metadata, htmlBody) <- getFileWithMetadata prev
         let htmlTemplates
               | isPostSource src = ["post.html", "default.html"]
@@ -319,7 +315,6 @@ main = do
       -- Build /Announcements/index.html
       outDir </> "Announcements" </> "index.html" %> \out -> do
         src <- routeSource out
-        -- putInfo $ printf "Compile '%s' to '%s'" src out
         postsField <- getPostsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
         return indexMarkdownTemplate
@@ -334,7 +329,6 @@ main = do
       -- Build rss.xml
       outDir </> "rss.xml" %> \out -> do
         src <- routeSource out
-        -- putInfo $ printf "Compile '%s' to '%s'" src out
         postsField <- getPostsField ()
         (fileMetadata, rssXmlTemplate) <- getFileWithMetadata src
[…]

### 2022-06-08 tools/Buildfile.hs
Move fixpoint to extra

@@ -562,7 +562,7 @@ epubSassOptions :: SassOptions
 epubSassOptions = def {sassIncludePaths = Just [epubStyleDir]}
 
 --------------------------------------------------------------------------------
--- File reader
+-- File Reader
 
 getFileWithMetadata ::
   ( ?getDefaultMetadata :: () -> Action Metadata,

### 2022-06-08 tools/Buildfile.hs
Agda as a resource

@@ -96,7 +96,9 @@ main = do
       }
     $ do
       --------------------------------------------------------------------------------
-      -- Agda libraries
+      -- Agda resource & libraries
+      
+      agda <- newResource "agda" 1
 
       standardLibrary <- liftIO $
         catch (Agda.getStandardLibrary "standard-library") $
@@ -264,7 +266,8 @@ main = do
           failOnError $
             Agda.resolveLibraryAndOutputFileName Agda.Html agdaLibraries src
         let tmpAgdaHtmlDirForLib = tmpAgdaHtmlDir </> agdaHtmlFileName </> Agda.libraryRoot lib </> includePath
-        Agda.compileTo Agda.Html agdaLibraries tmpAgdaHtmlDirForLib prev
+        withResource agda 1 $
+          Agda.compileTo Agda.Html agdaLibraries tmpAgdaHtmlDirForLib prev
 
       -- Stage 2: Compile Markdown to HTML
       tmpBodyHtmlDir <//> "*.html" %> \next -> do

### 2022-06-08 tools/Buildfile.hs
Update Buildfile.hs

@@ -258,7 +258,7 @@ main = do
 
       -- Stage 1: Compile Agda to HTML
       tmpAgdaHtmlDir <//> "*.md" %> \next -> do
-        (src, prev) <- (,) <$> routeSource next <*> routePrev next
+        (src, prev) <- (,) <$> routeSource out <*> routePrev out
         agdaLibraries <-
           failOnError $
             getAgdaLibrariesForProject <$> getProject src
@@ -266,6 +266,7 @@ main = do
           failOnError $
             Agda.resolveLibraryAndOutputFileName Agda.Html agdaLibraries src
         let tmpAgdaHtmlDirForLib = tmpAgdaHtmlDir </> agdaHtmlFileName </> Agda.libraryRoot lib </> includePath
+        need [prev]
         withResource agda 1 $
           Agda.compileTo Agda.Html agdaLibraries tmpAgdaHtmlDirForLib prev
 

### 2022-06-08 tools/Buildfile.hs
Fix Agda as resource

@@ -97,7 +97,7 @@ main = do
     $ do
       --------------------------------------------------------------------------------
       -- Agda resource & libraries
-      
+
       agda <- newResource "agda" 1
 
       standardLibrary <- liftIO $
@@ -258,7 +258,7 @@ main = do
 
       -- Stage 1: Compile Agda to HTML
       tmpAgdaHtmlDir <//> "*.md" %> \next -> do
-        (src, prev) <- (,) <$> routeSource out <*> routePrev out
+        (src, prev) <- (,) <$> routeSource next <*> routePrev next
         agdaLibraries <-
           failOnError $
             getAgdaLibrariesForProject <$> getProject src

### 2022-06-08 tools/Buildfile.hs
Minor layout changes

@@ -259,9 +259,7 @@ main = do
       -- Stage 1: Compile Agda to HTML
       tmpAgdaHtmlDir <//> "*.md" %> \next -> do
         (src, prev) <- (,) <$> routeSource next <*> routePrev next
-        agdaLibraries <-
-          failOnError $
-            getAgdaLibrariesForProject <$> getProject src
+        agdaLibraries <- failOnError $ getAgdaLibrariesForProject <$> getProject src
         (lib, includePath, agdaHtmlFileName) <-
           failOnError $
             Agda.resolveLibraryAndOutputFileName Agda.Html agdaLibraries src
@@ -276,8 +274,7 @@ main = do
         maybeAgdaLinkFixer <-
           getProject src
             & rightToMaybe
-            & traverse getAgdaLinkFixer
-            & fmap (fmap Pandoc.withUrls)
+            & traverse (fmap Pandoc.withUrls . getAgdaLinkFixer)
         readFile' prev
           >>= Pandoc.markdownToPandoc
           <&> fromMaybe id maybeAgdaLinkFixer
@@ -304,15 +301,9 @@ main = do
               bodyHtml <- routeAnchor "body_html" post
               (fileMetadata, body) <- getFileWithMetadata bodyHtml
               let bodyHtmlField = constField "body_html" body
-              url <-
-                failOnError $
-                  fileMetadata ^. "url"
-              let htmlTeaserField =
-                    ignoreError $
-                      htmlTeaserFieldFromHtml url body "teaser_html"
-              let textTeaserField =
-                    ignoreError $
-                      textTeaserFieldFromHtml body "teaser_plain"
+              url <- failOnError $ fileMetadata ^. "url"
+              let htmlTeaserField = ignoreError $ htmlTeaserFieldFromHtml url body "teaser_html"
+              let textTeaserField = ignoreError $ textTeaserFieldFromHtml body "teaser_plain"
               return $ mconcat [fileMetadata, bodyHtmlField, htmlTeaserField, textTeaserField]
             return $ constField "post" (reverse postsMetadata)
 
@@ -336,12 +327,7 @@ main = do
         postsField <- getPostsField ()
         (fileMetadata, rssXmlTemplate) <- getFileWithMetadata src
         buildDate <- currentDateField rfc822DateFormat "build_date"
-        let metadata =
-              mconcat
-                [ postsField,
-                  fileMetadata,
-                  buildDate
-                ]
[…]

### 2022-06-08 .feed-validator.config.js
Fix build and CI (#658)

@@ -1,0 +1,5 @@
+// Config for feed-validator for my website.
+
+module.exports = {
+  suppress: [{ level: "warning", type: "SelfDoesntMatchLocation" }],
+};

### 2022-06-08 hs/Main.hs

@@ -1,473 +1,0 @@
-{-# LANGUAGE OverloadedStrings #-}
-
-import           Control.Monad ((<=<), (>=>), forM_)
-import           Data.Char (toLower)
-import           Data.Functor ((<&>))
-import           Data.List (isPrefixOf, stripPrefix)
-import qualified Data.Text as T
-import qualified Data.Text.ICU as RE
-import qualified Data.Text.ICU.Replace as T
-import           Hakyll
-import           Hakyll.Web.Agda
-import           Hakyll.Web.Routes.Permalink
-import           Hakyll.Web.Template.Numeric
-import           Hakyll.Web.Template.Context.Derived
-import           Hakyll.Web.Template.Context.Metadata
-import           Hakyll.Web.Template.Context.Title
-import           Hakyll.Web.Sass
-import           System.FilePath ((</>), takeDirectory, replaceExtensions, splitPath, joinPath)
-import qualified Text.CSL as CSL
-import qualified Text.CSL.Pandoc as CSL (processCites)
-import           Text.Pandoc (Pandoc(..), ReaderOptions(..), WriterOptions(..), Extension(..))
-import qualified Text.Pandoc as Pandoc
-import           Text.Printf
-
---------------------------------------------------------------------------------
--- Configuration
---------------------------------------------------------------------------------
-
-siteReaderOptions :: ReaderOptions
-siteReaderOptions = defaultHakyllReaderOptions
-  { readerExtensions = Pandoc.extensionsFromList
-    [ Ext_all_symbols_escapable
-    , Ext_auto_identifiers
-    , Ext_backtick_code_blocks
-    , Ext_citations
-    , Ext_footnotes
-    , Ext_header_attributes
-    , Ext_implicit_header_references
-    , Ext_intraword_underscores
-    , Ext_markdown_in_html_blocks
-    , Ext_raw_html
-    , Ext_shortcut_reference_links
-    , Ext_smart
-    , Ext_superscript
-    , Ext_subscript
-    , Ext_task_lists
-    , Ext_yaml_metadata_block
-    ]
-  }
-
-siteWriterOptions :: WriterOptions
-siteWriterOptions = defaultHakyllWriterOptions
-
-siteContext :: Context String
-siteContext = mconcat
-  [ constField "pagetitle" "Programming Language Foundations in Agda"
-  , constField "pageurl" "https://plfa.github.io"
-  , constField "description" "An introduction to programming language theory using the proof assistant Agda."
[…]

### 2022-06-08 tools/Buildfile.hs

@@ -1,0 +1,688 @@
+{-# OPTIONS_GHC -Wno-unrecognised-pragmas #-}
+
+{-# HLINT ignore "Monad law, left identity" #-}
+{-# HLINT ignore "Redundant <&>" #-}
+module Main where
+
+import Buildfile.Author (Author (..))
+import Buildfile.Book (Book (..), Part (..), Section (..))
+import Buildfile.Contributor (Contributor (..))
+import Control.Monad (forM, forM_, unless)
+import Control.Monad.Except (MonadError (throwError))
+import Control.Monad.IO.Class (MonadIO)
+import Data.ByteString.Lazy qualified as LazyByteString
+import Data.Default.Class (Default (def))
+import Data.Either (fromRight, isRight)
+import Data.Function (on, (&))
+import Data.Functor ((<&>))
+import Data.Hashable (Hashable)
+import Data.List (isPrefixOf, sortBy)
+import Data.List qualified as List
+import Data.Maybe (fromMaybe)
+import Data.Text (Text)
+import Data.Text qualified as Text
+import Data.Traversable (for)
+import GHC.Generics (Generic)
+import Shoggoth.Agda qualified as Agda
+import Shoggoth.CSS.Minify qualified as CSS
+import Shoggoth.CSS.Sass (SassOptions (..))
+import Shoggoth.CSS.Sass qualified as CSS
+import Shoggoth.Configuration (CacheDirectory (..), OutputDirectory (..), TemplateDirectory (..))
+import Shoggoth.Metadata
+import Shoggoth.PostInfo
+import Shoggoth.Prelude
+import Shoggoth.Routing
+import Shoggoth.TagSoup qualified as TagSoup
+import Shoggoth.Template.Pandoc (Extension (..), HTMLMathMethod (..), HighlightStyle, Inlines, Lang, Locale, MetaValue, ObfuscationMethod (..), Pandoc (..), ReaderOptions (..), Reference, WriterOptions (..), extensionsFromList, runPandoc, runPandocWith)
+import Shoggoth.Template.Pandoc qualified as Pandoc
+import Shoggoth.Template.Pandoc.Builder qualified as Builder
+import Shoggoth.Template.Pandoc.Citeproc qualified as Citeproc
+import System.Directory (makeAbsolute)
+import Text.Printf (printf)
+import Control.Exception (catch)
+import System.Exit (exitWith, ExitCode (..))
+
+outDir, tmpDir :: FilePath
[…]

### 2022-06-08 versions/20.07/epub/render-liquid-template.rb

@@ -1,19 +1,0 @@
-#!/usr/bin/env ruby
-require 'yaml'
-require 'liquid'
-
-unless ARGV.length == 3
-  abort "Usage: render-liquid-template.rb [yaml_file] [markdown_in_file] [markdown_out_file]"
-end
-
-yaml_file, markdown_file_in, markdown_file_out = ARGV
-
-[yaml_file, markdown_file_in].each do |file|
-  unless File.file? file
-    abort "File does not exist: '%s'" % [file]
-  end
-end
-
-metadata = { 'site' => (YAML.load (File.read yaml_file)) }
-template = Liquid::Template.parse (File.read markdown_file_in)
-File.write markdown_file_out, (template.render metadata)

### 2022-06-18 tools/Buildfile.hs
Switch to dart sass, move html-minifier into Buildfile

@@ -22,16 +22,15 @@ import Data.Functor ((<&>))
 import Data.Hashable (Hashable)
 import Data.List (isPrefixOf, sortBy)
 import Data.List qualified as List
-import Data.Maybe (fromMaybe, isNothing)
+import Data.Maybe (fromMaybe, isNothing, isJust)
 import Data.Text (Text)
 import Data.Text qualified as Text
+import Data.Text.Lazy qualified as LazyText
+import Data.Text.Lazy.Encoding qualified as LazyText
 import Data.Traversable (for)
 import Debug.Trace qualified as Debug
 import GHC.Generics (Generic)
 import Shoggoth.Agda qualified as Agda
-import Shoggoth.CSS.Minify qualified as CSS
-import Shoggoth.CSS.Sass (SassOptions (..))
-import Shoggoth.CSS.Sass qualified as CSS
 import Shoggoth.Configuration (CacheDirectory (..), OutputDirectory (..), TemplateDirectory (..))
 import Shoggoth.Metadata
 import Shoggoth.PostInfo
@@ -45,6 +44,8 @@ import Shoggoth.Template.Pandoc.Citeproc qualified as Citeproc
 import System.Directory (makeAbsolute)
 import System.Exit (ExitCode (..), exitWith)
 import Text.Printf (printf)
+import GHC.Stack.Types (HasCallStack)
+import qualified Data.Text.Encoding as Text
 
 outDir, tmpDir :: FilePath
 outDir = "_site"
@@ -256,7 +257,7 @@ main = do
           >>= Pandoc.applyAsTemplate (tocField <> fileMetadata)
           >>= markdownToHtml5
           >>= Pandoc.applyTemplates ["page.html", "default.html"] fileMetadata
-          <&> postProcessHtml5 outDir out
+          >>= postProcessHtml5 outDir out
           >>= writeFile' out
 
       --------------------------------------------------------------------------------
@@ -299,7 +300,8 @@ main = do
               | isPostSource src = ["post.html", "default.html"]
               | otherwise = ["page.html", "default.html"]
         html <- Pandoc.applyTemplates htmlTemplates metadata htmlBody
-        writeFile' out $ postProcessHtml5 outDir out html
+        html <- postProcessHtml5 outDir out html
+        writeFile' out html
 
       --------------------------------------------------------------------------------
       -- Posts
@@ -325,7 +327,7 @@ main = do
[…]

### 2022-06-18 tools/Buildfile.hs
Minor change

@@ -22,7 +22,7 @@ import Data.Functor ((<&>))
 import Data.Hashable (Hashable)
 import Data.List (isPrefixOf, sortBy)
 import Data.List qualified as List
-import Data.Maybe (fromMaybe, isNothing, isJust)
+import Data.Maybe (fromMaybe, isNothing, isJust, maybeToList)
 import Data.Text (Text)
 import Data.Text qualified as Text
 import Data.Text.Lazy qualified as LazyText
@@ -741,13 +741,13 @@ npx cmdOpts package args = do
 
 sass :: (HasCallStack, CmdResult r) => [CmdOption] -> [String] -> Maybe LazyText.Text -> Action r
 sass cmdOpts args maybeStdin = do
-  let stdinCmdOpt = maybe mempty (\stdin -> [StdinBS (LazyText.encodeUtf8 stdin)]) maybeStdin
+  let stdinCmdOpt = [StdinBS (LazyText.encodeUtf8 stdin) | stdin <- maybeToList maybeStdin]
   let stdinArg = ["--stdin" | isJust maybeStdin]
   npx (stdinCmdOpt <> cmdOpts) "sass" (stdinArg <> args)
 
 htmlMinifier :: (HasCallStack, CmdResult r) => [CmdOption] -> [String] -> Maybe LazyText.Text -> Action r
 htmlMinifier cmdOpts args maybeStdin = do
-  let stdinCmdOpt = maybe mempty (\stdin -> [StdinBS (LazyText.encodeUtf8 stdin)]) maybeStdin
+  let stdinCmdOpt = [StdinBS (LazyText.encodeUtf8 stdin) | stdin <- maybeToList maybeStdin]
   npx (stdinCmdOpt <> cmdOpts) "html-minifier" (defaultHtmlMinifierArgs <> args)
   where
     defaultHtmlMinifierArgs :: [String]

### 2022-06-18 tools/Buildfile.hs
Minor fix

@@ -374,7 +374,7 @@ main = do
       -- Build assets/css/highlight.css
       outDir </> "assets/css/highlight.css" %> \out -> do
         let stdin = LazyText.fromChunks [highlightCss]
-        sass [] ["--load-path=" <> epubStyleDir, "--style=compressed", out] (Just stdin)
+        sass [] ["--style=compressed", out] (Just stdin)
 
       -- Copy static assets
       outDir </> "assets" <//> "*" %> \out -> do

### 2022-06-18 tools/Buildfile.hs
Minor fix

@@ -373,8 +373,7 @@ main = do
 
       -- Build assets/css/highlight.css
       outDir </> "assets/css/highlight.css" %> \out -> do
-        let stdin = LazyText.fromChunks [highlightCss]
-        sass [] ["--style=compressed", out] (Just stdin)
+        sass [] ["--style=compressed", out] (Just $ LazyText.pack highlightCss)
 
       -- Copy static assets
       outDir </> "assets" <//> "*" %> \out -> do
@@ -825,8 +824,8 @@ writerOpts =
 highlightStyle :: HighlightStyle
 highlightStyle = Pandoc.pygments
 
-highlightCss :: Text
-highlightCss = Text.pack $ Pandoc.styleToCss highlightStyle
+highlightCss :: String
+highlightCss = Pandoc.styleToCss highlightStyle
 
 --------------------------------------------------------------------------------
 -- Helper functions

### 2022-06-18 tools/Buildfile.hs
Fix #687

@@ -7,7 +7,7 @@
 module Main where
 
 import Buildfile.Author (Author (..))
-import Buildfile.Book (Book (..), Part (..), Section (..), SectionTable, fromBook, nextSection, previousSection)
+import Buildfile.Book (Book (..), Part (..), Chapter (..), chapterTable, fromBook, nextChapter, previousChapter)
 import Buildfile.Contributor (Contributor (..))
 import Control.Exception (assert, catch)
 import Control.Monad (forM, forM_, unless, when, (>=>))
@@ -192,9 +192,9 @@ main = do
         readYaml @Book tableOfContentsFile
       let ?getTableOfContents = getTableOfContents
 
-      getSectionTable <- newCache $ \() -> do
+      getChapterTable <- newCache $ \() -> do
         fromBook <$> getTableOfContents ()
-      let ?getSectionTable = getSectionTable
+      let ?getChapterTable = getChapterTable
 
       getDefaultMetadata <- newCache $ \() -> do
         metadata <- readYaml @Metadata (dataDir </> "metadata.yml")
@@ -400,10 +400,10 @@ main = do
 
         defaultMetadata <- ?getDefaultMetadata ()
 
-        let routeSection src =
+        let routeChapter src =
               if Agda.isAgdaFile src then routeAnchor "agda_html" src else return src
 
-        bookDoc <- makeBookDoc routeSection
+        bookDoc <- makeBookDoc routeChapter
 
         standaloneHtmlTemplate <- Pandoc.compileTemplateFile (webTemplateDir </> "standalone.html")
 
@@ -415,7 +415,7 @@ main = do
                   writerWrapText = Pandoc.WrapPreserve,
                   writerTopLevelDivision = Pandoc.TopLevelPart,
                   writerTOCDepth = 2,
-                  writerReferenceLocation = Pandoc.EndOfSection
+                  writerReferenceLocation = Pandoc.EndOfChapter
                 }
 
         let pandocToStandaloneHtml doc = do
@@ -437,11 +437,11 @@ main = do
         -- Require metadata and stylesheet
         need [tmpEpubDir </> "epub-metadata.xml", tmpEpubDir </> "style.css"]
 
-        let routeSection src =
+        let routeChapter src =
               if Agda.isAgdaFile src then routeAnchor "agda_html" src else return src
 
         bookDoc <-
-          makeBookDoc routeSection
[…]

### 2022-06-18 tools/Buildfile/Book.hs

@@ -23,7 +23,7 @@ instance ToJSON Book where
 
 data Part = Part
   { partTitle :: Text,
-    partSections :: [Section],
+    partChapters :: [Chapter],
     partFrontmatter :: Bool,
     partMainmatter :: Bool,
     partBackmatter :: Bool
@@ -34,7 +34,7 @@ instance FromJSON Part where
   parseJSON = withObject "Part" $ \v ->
     Part
       <$> v .: "title"
-      <*> v .: "section"
+      <*> v .: "chapter"
       <*> v .:? "frontmatter" .!= False
       <*> v .:? "mainmatter" .!= False
       <*> v .:? "backmatter" .!= False
@@ -43,53 +43,53 @@ instance ToJSON Part where
   toJSON Part {..} =
     object
       [ "title" .= partTitle,
-        "section" .= partSections,
+        "chapter" .= partChapters,
         "frontmatter" .= partFrontmatter,
         "mainmatter" .= partMainmatter,
         "backmatter" .= partBackmatter
       ]
 
-data Section = Section
-  { sectionInclude :: FilePath,
-    sectionEpubType :: Text
+data Chapter = Chapter
+  { chapterInclude :: FilePath,
+    chapterEpubType :: Text
   }
   deriving (Show)
 
-instance FromJSON Section where
-  parseJSON = withObject "Section" $ \v ->
-    Section
+instance FromJSON Chapter where
+  parseJSON = withObject "Chapter" $ \v ->
+    Chapter
       <$> v .: "include"
       <*> v .:? "epub-type" .!= "bodymatter"
 
-instance ToJSON Section where
-  toJSON Section {..} =
+instance ToJSON Chapter where
+  toJSON Chapter {..} =
     object
-      [ "include" .= sectionInclude,
-        "epub-type" .= sectionEpubType
+      [ "include" .= chapterInclude,
+        "epub-type" .= chapterEpubType
       ]
 
-newtype SectionTable = SectionTable {sectionBimap :: Bimap.Bimap FilePath FilePath}
+newtype ChapterTable = ChapterTable {chapterBimap :: Bimap.Bimap FilePath FilePath}
 
-fromBook :: Book -> SectionTable
-fromBook book = SectionTable {..}
+fromBook :: Book -> ChapterTable
+fromBook book = ChapterTable {..}
   where
-    sectionBimap :: Bimap.Bimap FilePath FilePath
-    sectionBimap = Bimap.fromList $ zip sectionList (tail sectionList)
[…]

### 2022-06-18 tools/Buildfile.hs
Update Buildfile.hs

@@ -7,7 +7,7 @@
 module Main where
 
 import Buildfile.Author (Author (..))
-import Buildfile.Book (Book (..), Part (..), Chapter (..), chapterTable, fromBook, nextChapter, previousChapter)
+import Buildfile.Book (Book (..), Part (..), Chapter (..), ChapterTable, fromBook, nextChapter, previousChapter)
 import Buildfile.Contributor (Contributor (..))
 import Control.Exception (assert, catch)
 import Control.Monad (forM, forM_, unless, when, (>=>))

### 2022-06-18 tools/Buildfile.hs
Fix writerReferenceLocation

@@ -415,7 +415,7 @@ main = do
                   writerWrapText = Pandoc.WrapPreserve,
                   writerTopLevelDivision = Pandoc.TopLevelPart,
                   writerTOCDepth = 2,
-                  writerReferenceLocation = Pandoc.EndOfChapter
+                  writerReferenceLocation = Pandoc.EndOfDocument
                 }
 
         let pandocToStandaloneHtml doc = do
@@ -461,7 +461,7 @@ main = do
                   writerEpubFonts = epubFonts,
                   writerEpubChapterLevel = 2,
                   writerTOCDepth = 2,
-                  writerReferenceLocation = Pandoc.EndOfChapter
+                  writerReferenceLocation = Pandoc.EndOfDocument
                 }
 
         epub <- runPandoc $ Pandoc.writeEPUB3 writerOptsForEpub bookDoc

### 2022-06-18 tools/Buildfile.hs
Fixed various errors

@@ -7,7 +7,7 @@
 module Main where
 
 import Buildfile.Author (Author (..))
-import Buildfile.Book (Book (..), Part (..), Chapter (..), ChapterTable, fromBook, nextChapter, previousChapter)
+import Buildfile.Book (Book (..), Chapter (..), ChapterTable, Part (..), fromBook, nextChapter, previousChapter)
 import Buildfile.Contributor (Contributor (..))
 import Control.Exception (assert, catch)
 import Control.Monad (forM, forM_, unless, when, (>=>))
@@ -16,20 +16,22 @@ import Control.Monad.IO.Class (MonadIO)
 import Control.Monad.State (evalState)
 import Data.ByteString.Lazy qualified as LazyByteString
 import Data.Default.Class (Default (def))
+import Data.Digest.Pure.SHA (sha512, showDigest)
 import Data.Either (fromRight, isRight)
 import Data.Function (on, (&))
 import Data.Functor ((<&>))
 import Data.Hashable (Hashable)
 import Data.List (isPrefixOf, sortBy)
 import Data.List qualified as List
-import Data.Maybe (fromMaybe, isNothing, isJust, maybeToList)
+import Data.Maybe (fromMaybe, isJust, isNothing, maybeToList)
 import Data.Text (Text)
 import Data.Text qualified as Text
+import Data.Text.Encoding qualified as Text
 import Data.Text.Lazy qualified as LazyText
 import Data.Text.Lazy.Encoding qualified as LazyText
 import Data.Traversable (for)
-import Debug.Trace qualified as Debug
 import GHC.Generics (Generic)
+import GHC.Stack.Types (HasCallStack)
 import Shoggoth.Agda qualified as Agda
 import Shoggoth.Configuration (CacheDirectory (..), OutputDirectory (..), TemplateDirectory (..))
 import Shoggoth.Metadata
@@ -44,8 +46,6 @@ import Shoggoth.Template.Pandoc.Citeproc qualified as Citeproc
 import System.Directory (makeAbsolute)
 import System.Exit (ExitCode (..), exitWith)
 import Text.Printf (printf)
-import GHC.Stack.Types (HasCallStack)
-import qualified Data.Text.Encoding as Text
 
 outDir, tmpDir :: FilePath
 outDir = "_site"
@@ -148,6 +148,7 @@ main = do
             mconcat
               [ -- Book (Web Version)
[…]

### 2022-06-18 tools/Buildfile.hs
Fix broken SRI

@@ -14,9 +14,10 @@ import Control.Monad (forM, forM_, unless, when, (>=>))
 import Control.Monad.Except (MonadError (throwError))
 import Control.Monad.IO.Class (MonadIO)
 import Control.Monad.State (evalState)
+import Data.ByteString.Lazy.Base64 qualified as LazyByteString
 import Data.ByteString.Lazy qualified as LazyByteString
 import Data.Default.Class (Default (def))
-import Data.Digest.Pure.SHA (sha512, showDigest)
+import Data.Digest.Pure.SHA qualified as Digest (sha512, bytestringDigest)
 import Data.Either (fromRight, isRight)
 import Data.Function (on, (&))
 import Data.Functor ((<&>))
@@ -217,12 +218,12 @@ main = do
         standardLibrary <- Agda.getStandardLibrary
         Agda.makeAgdaLinkFixer (Just standardLibrary) (List.delete standardLibrary libraries) []
 
-      getFileDigestSHA512 <- newCache $ \src ->
+      getDigest <- newCache $ \src ->
         liftIO $ do
           stream <- LazyByteString.readFile src
-          let digest = sha512 stream
-          return $ "sha512-" <> showDigest digest
-      let ?getFileDigestSHA512 = getFileDigestSHA512
+          let digest = Digest.sha512 stream
+          return $ "sha512-" <> LazyByteString.encodeBase64 (Digest.bytestringDigest digest)
+      let ?getDigest = getDigest
 
       getTemplateFile <- Pandoc.makeCachedTemplateFileGetter
       let ?getTemplateFile = getTemplateFile
@@ -260,7 +261,7 @@ main = do
         src <- routeSource out
         tocField <- getTableOfContentsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
-        cssField <- getCssField
+        cssField <- getDefaultCssMetadata
         let metadata = mconcat [tocField, fileMetadata, cssField]
         return indexMarkdownTemplate
           >>= Pandoc.applyAsTemplate metadata
@@ -304,7 +305,7 @@ main = do
       outDir <//> "*.html" %> \out -> do
         (src, prev) <- (,) <$> routeSource out <*> routePrev out
         (fileMetadata, htmlBody) <- getFileWithMetadata prev
-        cssField <- getCssField
+        cssField <- getDefaultCssMetadata
         let metadata = mconcat [fileMetadata, cssField]
         let htmlTemplates
[…]

### 2022-06-18 tools/Buildfile.hs
Minor changes

@@ -14,10 +14,10 @@ import Control.Monad (forM, forM_, unless, when, (>=>))
 import Control.Monad.Except (MonadError (throwError))
 import Control.Monad.IO.Class (MonadIO)
 import Control.Monad.State (evalState)
-import Data.ByteString.Lazy.Base64 qualified as LazyByteString
 import Data.ByteString.Lazy qualified as LazyByteString
+import Data.ByteString.Lazy.Base64 qualified as LazyByteString
 import Data.Default.Class (Default (def))
-import Data.Digest.Pure.SHA qualified as Digest (sha512, bytestringDigest)
+import Data.Digest.Pure.SHA qualified as Digest (bytestringDigest, sha512)
 import Data.Either (fromRight, isRight)
 import Data.Function (on, (&))
 import Data.Functor ((<&>))
@@ -261,7 +261,7 @@ main = do
         src <- routeSource out
         tocField <- getTableOfContentsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
-        cssField <- getDefaultCssMetadata
+        cssField <- getCssMetadata
         let metadata = mconcat [tocField, fileMetadata, cssField]
         return indexMarkdownTemplate
           >>= Pandoc.applyAsTemplate metadata
@@ -305,7 +305,7 @@ main = do
       outDir <//> "*.html" %> \out -> do
         (src, prev) <- (,) <$> routeSource out <*> routePrev out
         (fileMetadata, htmlBody) <- getFileWithMetadata prev
-        cssField <- getDefaultCssMetadata
+        cssField <- getCssMetadata
         let metadata = mconcat [fileMetadata, cssField]
         let htmlTemplates
               | isPostSource src = ["post.html", "default.html"]
@@ -334,7 +334,7 @@ main = do
         src <- routeSource out
         postsField <- getPostsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
-        cssField <- getDefaultCssMetadata
+        cssField <- getCssMetadata
         let metadata = mconcat [postsField, fileMetadata, cssField]
         return indexMarkdownTemplate
           >>= Pandoc.applyAsTemplate metadata
@@ -347,7 +347,7 @@ main = do
         src <- routeSource out
         contributorField <- constField "contributor" <$> getContributors ()
[…]

### 2022-06-18 papers/scp/response-to-reviewers.txt
Replace http with https

@@
 to retrieve password details, please go to:
-http://ees.elsevier.com/scico/automail_query.asp
+https://ees.elsevier.com/scico/automail_query.asp
 
@@
 Example MethodsX articles can be found here:
-http://www.sciencedirect.com/science/journal/22150161 Yours sincerely,
+https://www.sciencedirect.com/science/journal/22150161 Yours sincerely,
 
@@
 Science of Computer Programming features the Interactive Plot Viewer,
-see: http://www.elsevier.com/interactiveplots. Interactive Plots
+see: https://www.elsevier.com/interactiveplots. Interactive Plots
 provide easy access to the data behind plots. To include one with your
@@
 online at
-http://authortools.elsevier.com/interactiveplots/verification before
+https://authortools.elsevier.com/interactiveplots/verification before
 submission as supplementary material.  PLEASE NOTE: Science of
@@
 For further assistance, please visit our customer support site at
-http://help.elsevier.com/app/answers/list/p/7923. Here you can search
+https://help.elsevier.com/app/answers/list/p/7923. Here you can search
 for solutions on a range of topics, find answers to frequently asked

### 2022-06-18 courses/TSPL/make_assignment.sh

@@ -55,7 +55,7 @@ Please ensure your files execute correctly under Agda!
 Please remember the University requirement as
 regards all assessed work. Details about this can be found at:
 
-> [http://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct](http://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct)
+> [https://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct](https://web.inf.ed.ac.uk/infweb/admin/policies/academic-misconduct)
 
 Furthermore, you are required to take reasonable measures to protect
 your assessed work from unauthorised access. For example, if you put

### 2022-06-18 tools/Buildfile.hs
Suppress confusing EPUBCheck errors

@@ -261,7 +261,7 @@ main = do
         src <- routeSource out
         tocField <- getTableOfContentsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
-        cssField <- getCssMetadata
+        cssField <- getCssField
         let metadata = mconcat [tocField, fileMetadata, cssField]
         return indexMarkdownTemplate
           >>= Pandoc.applyAsTemplate metadata
@@ -305,7 +305,7 @@ main = do
       outDir <//> "*.html" %> \out -> do
         (src, prev) <- (,) <$> routeSource out <*> routePrev out
         (fileMetadata, htmlBody) <- getFileWithMetadata prev
-        cssField <- getCssMetadata
+        cssField <- getCssField
         let metadata = mconcat [fileMetadata, cssField]
         let htmlTemplates
               | isPostSource src = ["post.html", "default.html"]
@@ -334,7 +334,7 @@ main = do
         src <- routeSource out
         postsField <- getPostsField ()
         (fileMetadata, indexMarkdownTemplate) <- getFileWithMetadata src
-        cssField <- getCssMetadata
+        cssField <- getCssField
         let metadata = mconcat [postsField, fileMetadata, cssField]
         return indexMarkdownTemplate
           >>= Pandoc.applyAsTemplate metadata
@@ -347,7 +347,7 @@ main = do
         src <- routeSource out
         contributorField <- constField "contributor" <$> getContributors ()
         (fileMetadata, acknowledgmentsMarkdownTemplate) <- getFileWithMetadata src
-        cssField <- getCssMetadata
+        cssField <- getCssField
         let metadata = mconcat [contributorField, fileMetadata, cssField]
         return acknowledgmentsMarkdownTemplate
           >>= Pandoc.applyAsTemplate metadata
@@ -372,7 +372,7 @@ main = do
       outDir </> "404.html" %> \out -> do
         src <- routeSource out
         (fileMetadata, errorMarkdownBody) <- getFileWithMetadata src
-        cssField <- getCssMetadata
+        cssField <- getCssField
         return errorMarkdownBody
           >>= markdownToHtml5
           >>= Pandoc.applyTemplates ["page.html", "default.html"] (fileMetadata <> cssField)
@@ -567,17 +567,17 @@ isPostOutput out = isRight $ parsePostOutput (makeRelative outDir out)
 --------------------------------------------------------------------------------
[…]

### 2022-06-18 tools/Buildfile.hs
Rename style.css to light.css

@@ -169,7 +169,7 @@ main = do
                 [courseDir <//> "*.pdf"] *|-> \src -> outDir </> makeRelative courseDir src,
                 -- Assets
                 [webDir </> "404.md"] |-> permalinkRouter,
-                [webStyleDir </> "style.scss"] |-> outDir </> "assets/css/style.css",
+                [webStyleDir </> "light.scss"] |-> outDir </> "assets/css/light.css",
                 create (outDir </> "assets/css/highlight.css"),
                 [webAssetDir <//> "*"] *|-> \src -> outDir </> "assets" </> makeRelative webAssetDir src
               ]
@@ -378,8 +378,8 @@ main = do
           >>= Pandoc.applyTemplates ["page.html", "default.html"] (fileMetadata <> cssField)
           >>= writeHtml5 outDir out
 
-      -- Build assets/css/style.css
-      outDir </> "assets/css/style.css" %> \out -> do
+      -- Build assets/css/light.css
+      outDir </> "assets/css/light.css" %> \out -> do
         src <- routeSource out
         need =<< getDirectoryFiles "" [webStyleDir <//> "*.scss"]
         sass [] ["--load-path=" <> webStyleDir, "--style=compressed", src, out] Nothing
@@ -573,9 +573,10 @@ getCssField ::
   ) =>
   Action Metadata
 getCssField =
-  constField "css" <$>
-    traverse getCssField1
-      [outDir </> "assets/css/style.css", outDir </> "assets/css/highlight.css"]
+  constField "css"
+    <$> traverse
+      getCssField1
+      [outDir </> "assets/css/light.css", outDir </> "assets/css/highlight.css"]
   where
     getCssField1 out = do
       need [out]

### 2022-06-18 tools/Buildfile.hs
Generate dark.css

@@ -170,6 +170,7 @@ main = do
                 -- Assets
                 [webDir </> "404.md"] |-> permalinkRouter,
                 [webStyleDir </> "light.scss"] |-> outDir </> "assets/css/light.css",
+                [webStyleDir </> "dark.scss"] |-> outDir </> "assets/css/dark.css",
                 create (outDir </> "assets/css/highlight.css"),
                 [webAssetDir <//> "*"] *|-> \src -> outDir </> "assets" </> makeRelative webAssetDir src
               ]

### 2022-06-19 tools/Buildfile.hs
Update Buildfile.hs

@@ -754,7 +754,7 @@ urlToIdent url =
 npx :: (HasCallStack, CmdResult r) => [CmdOption] -> String -> [String] -> Action r
 npx cmdOpts package args = do
   need ["package.json"]
-  command (AddEnv "npm_config_yes" "true" : Traced package : cmdOpts) "npx" (package : args)
+  command (AddEnv "npm_config_yes" "true" : Traced package : cmdOpts) ("npx" <.> exe) (package : args)
 
 sass :: (HasCallStack, CmdResult r) => [CmdOption] -> [String] -> Maybe LazyText.Text -> Action r
 sass cmdOpts args maybeStdin = do

### 2022-08-05 papers/scp/submission/PLFA.tex
Setup pre-commit

--- 2019-07-18 papers/scp/PLFA.tex
@@
   \begin{center}
-    \url{http://plfa.inf.ed.ac.uk}
+    \url{https://plfa.inf.ed.ac.uk}
   \end{center}
@@
 \begin{center}
-  \url{http://plfa.inf.ed.ac.uk/stable}
+  \url{https://plfa.inf.ed.ac.uk/stable}
 \end{center}

### 2022-11-27 courses/TSPL/2022/Instructions.tex
update exam instructions

--- 2019-05-29 courses/tspl/2018/Instructions.tex
@@
 \usepackage{a4,amssymb}
+\usepackage{hyperref}
+
 % \setlength{\oddsidemargin}{-1.5cm}
@@
 \begin{center}
-\large Types and Semantics for Programming Languages
+\large Types and Semantics for Programming Languages \\
+\Large Final Exam
 \end{center}
@@
 
-\begin{enumerate}
+\begin{itemize}
 
@@
 \item
-  You may log into your computer as soon as you are ready to do so.
+  You are permitted to use your browser to access the following:
+  \begin{quote}
+    \url{plfa.inf.ed.ac.uk} -- the course textbook \\
+    \url{agda.readthedocs.io/en/v2.6.2.2/} -- the Agda user manual
+  \end{quote}
+  You are also permitted to bring in two sides of A4 printed notes,
+  covering whatever you find helpful. These may be inspected by the
+  invigilator.
 
 \item
-  To download the exam paper, open a terminal window and type:
-  \begin{center}
-    \texttt{getpapers}
-  \end{center}
-  This will create a subdirectory \texttt{tspl-pe} in your home directory,
-  containing the following files.
-  \begin{center}
-    \begin{tabular}{ll}
-      \texttt{tspl-pe/papers/exam.pdf}                & the exam \\
-      \texttt{tspl-pe/papers/instructions.pdf}        & these instructions \\
-      \texttt{tspl-pe/templates/Exam.lagda}           & exam template to edit \\
-      \texttt{tspl-pe/original\_templates/Exam.lagda} & backup of exam template
-    \end{tabular}
-  \end{center}
-  The directories \texttt{tspl-pe/papers} and
-  \texttt{tspl-pe/original\_templates}
-  are read only, but you may read and write \texttt{tspl-pe/templates}.
+  You are not permitted to bring in or access other materials during
+  the exam. The only web pages you may browse during the exam are
+  those above. The only files you may access are the file \texttt{Exam.lagda.md}
+  and the parts of the standard library directly referenced by
+  that file. The invigilator may see what materials you are accessing
+  at any time.
 
 \item
-  To setup the Agda environment, open a second terminal window and type:
-  \begin{center}
-    \texttt{tspl-setup}
-  \end{center}
-  This will open a browser, with three tabs which contain the textbook
-  \emph{Programming Language Foundations in Agda}, the documentation for the
-  Agda standard library, and the documentation for Agda.  (The browser may
-  also attempt to link to the internet, which brings up a tab labeled
-  ``Problem loading page''; this is expected and not a problem.)
+  Please ensure before submission that the file
+  \texttt{Exam.lagda.md} contains your solutions to the exam.
+  Submit using Gradescope as normal.
 
-{\it (Note that internet access has been disabled.)}
-
-\begin{center}
-  \textbf{Do nothing further until the start of the exam is announced!}
-\end{center}
-
-\hfill \textit{Please Turn Over}
-\newpage
-
-\item When the start of the exam is announced, open the exam paper
-\begin{center}
-  \texttt{tspl-pe/papers/exam.pdf}
-\end{center}
-with the standard PDF viewer.
-
-\item Change to the template directory
-\begin{center}
-  \texttt{cd tspl-pe/templates}
-\end{center}
-and edit the file
-\begin{center}
-  \texttt{Exam.lagda}
-\end{center}
-to include your answers, using Emacs or anything else suitable.
-You are recommended to save your work on a regular basis.
-
-\item Before submitting, make sure your file is processed by Agda
-  correctly. Code which prevents the file from compiling should be
-  made into comments. If you fail to solve part of a problem, you
-  may get more credit if you indicate clearly which part you have
-  not solved.
-
-\item \emph{Please ensure before submission that the file}
-  \texttt{Exam.lagda} \emph{contains your solutions to the exam.}  Submit
-  your file using the command:
-  \begin{center}
-  \texttt{examsubmit Exam.lagda}
-  \end{center}
-  If you get an error, please check carefully that your file is called
-  \texttt{Exam.lagda} and that you are in the same directory as this
-  file. If you continue to have problems, please contact one of the
-  invigilators.
-
-Repeated submit commands are allowed, and will overwrite previous
-submissions.  The last file submitted will be the one marked.
-
-\item When the invigilators announce the end of the exam, you must
+\item
+  When the invigilators announce the end of the exam, you must
   submit and log out immediately.
 
-\end{enumerate}
+\end{itemize}
 
